---
title: "R Notebook"
output: html_notebook
---

```{r}
library(ggplot2)
library(gridExtra)
library("dplyr")
library(ComplexHeatmap)
library(tidyr)
library(paletteer)
library(vegan)
library(circlize)

source("D:/users/fperez/NKI_TMAs_AF/devNKI-scripts/05_Spatial-interactions/RCN_inspection_functions.R")
```

#Reading clinical info
```{r}
rcn.folder <- "Analysis_results/03_RCN/"
input.prefix <- "Cluster_ids_metaclustersmerged_"
project.folder <- "D:/users/fperez/NKI_TMAs_AF/"
output.napari <- "RCN_Napary2023_2"
out.folder.name <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/03_RCN_2024/"
method.rcn <- "radious46_k30"


cycif2samples <- read.table(file="D:/users/fperez/NKI_TMAs_AF/devNKI-scripts/03_Cell-typeClassification/tCicyf_patients-cores_pass.csv", sep=",", header = TRUE)
clinical.info <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Clinical_data/T-CycIF.updated.csv", sep=",",
                            header = TRUE)

clin.tcycif <- merge(cycif2samples, clinical.info, by.x="patient", by.y="TnumberAVL")

#Loading Molecular profiles
M.profiles <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Tables/Molecular_profiles_patients_20231208.csv",
             sep=",", header = TRUE)

#List of cores to get RNC data in Napari format
cores.interest <- c(24,28,31,109,110,112,113,114,116,118,120,122,10,11,111,113,115,117,119,121,141)
cores.interest <- sort(cores.interest)
cores.interest <- paste0("core",cores.interest)
```


#Merging M.profile info with tCycif core ID
```{r}
M.profiles.clin  <- merge(M.profiles, cycif2samples, by="patient", all.x=TRUE)
M.profiles.clin$imageid <- paste(M.profiles.clin$cycif.slide, M.profiles.clin$cycif.core.id, sep="_")
M.profiles.clin <- M.profiles.clin %>% select(patient,Molecular.profile2,therapy_sequence,imageid)
M.profiles.clin
```



```{r}
length(unique(M.profiles.clin$patient))
```



#Reading input RCN table file
```{r}
#Reading RCN id per cell
RCN.clusters <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Analysis_results/Spatial_analysis/Cluster_ids_radius46_k40_metaclusters_20240113.csv", header = TRUE, sep=",")

#Next lines to rename the RNC id according to the abundance. So, RCN=1 would be the most abundant 
RCN.order <- order(table(RCN.clusters[,2]), decreasing = TRUE)
RCN.order.values <- sapply(1:length(RCN.order), function(x){which(RCN.order == x)})
RCN.clusters[,2] <- RCN.order.values[c(RCN.clusters[,2] + 1)]

#Reading the merged RCNs ids
RCN.annotations <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Analysis_results/03_RCN_2024_metaclusters/Merged_RCNs.csv",
                              sep=",",header=TRUE)

#Adding an RCNid number of the merged RCNs
RCN.c.ids <- paste0("RCN", sprintf('%0.2d',1:length(unique(RCN.annotations$RCN_name))))
names(RCN.c.ids) <- unique(RCN.annotations$RCN_name)

RCN.c.ids.all <- sapply(RCN.annotations$RCN_name, function(x){
      RCN.c.ids[which(x == names(RCN.c.ids))]
})

RCN.annotations$RCN.id.collapsed <- RCN.c.ids.all
RCN.annotations$RCN_id <- as.integer(gsub("RCN","",RCN.annotations$RCN_id))


#Next lines to rename the RNC id according to the abundance. So, RCN=1 would be the most abundant 
RCN.order <- order(table(RCN.annotations[RCN.clusters$spatial_kmeans,"RCN.id.collapsed"]), decreasing = TRUE)
new.rcn.order <- paste0("RCN", sprintf('%0.2d',1:length(unique(RCN.annotations$RCN.id.collapsed))))
names(new.rcn.order) <- unique(RCN.annotations$RCN.id.collapsed)[RCN.order]

RCN.annotations$RCN.id.collapsed.ordered <- new.rcn.order[RCN.annotations$RCN.id.collapsed]
RCN.annotations
```

##Re-naming the RCN accordint to the RNC group they belong
```{r}
#New names
new.vals <- c("RCN01", "RCN02", "RCN03","RCN04", "RCN05", "RCN06","RCN07", "RCN08", "RCN09","RCN10",
          "RCN11", "RCN12", "RCN13", "RCN14")

#Original names, with the order according to the groups
order <- c("RCN01", "RCN03", "RCN11", "RCN02", "RCN05", "RCN06", "RCN09", "RCN12", "RCN04", "RCN07",
               "RCN08", "RCN10", "RCN13", "RCN14")

names(new.vals) <- order
RCN.annotations$RCN.id.collapsed.ordered <- new.vals[RCN.annotations$RCN.id.collapsed.ordered]
```


#Reading input cell dataset
```{r}
cells <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Tables/All_cells_subtype_scimap-format1_20231129.csv",
                    header = TRUE, sep=",")
```


#Adding the core.id and slide.id
```{r}
cells.aux <- cells

cells.aux$cycif.slide  <- sapply(cells$imageid, function(x){
                      splitted <- strsplit(x,"_")[[1]]
                      concat <- paste(splitted[1],splitted[2],splitted[3], sep="_")
                      return(concat)
                    })

cells.aux$cycif.core.id <- sapply(cells$imageid, function(x){
                      splitted <- strsplit(x,"_")[[1]]
                      return(splitted[4])
                    })
```

#Merging global-celltype values
```{r}
cells.aux$GlobalCellType[grep("Stromal", cells.aux$GlobalCellType)] = "Stromal"
cells.aux$GlobalCellType[grep("CancerClus1", cells.aux$GlobalCellType)] = "Can.Idle"
cells.aux$GlobalCellType[grep("CancerClus2", cells.aux$GlobalCellType)] = "Can.Proliferative"
cells.aux$GlobalCellType[grep("CancerClus3", cells.aux$GlobalCellType)] = "Can.EMT"
cells.aux$GlobalCellType[grep("CancerClus4", cells.aux$GlobalCellType)] = "Can.Immunogenic"
```

#Merging cells ids with clinical info
```{r}
#Merging cells ids with patient.id 
cells.pat <- merge(cells.aux, cycif2samples, by=c("cycif.slide","cycif.core.id"))

#Merging cells ids with clinical info
cells.clin <- merge(cells.pat, clinical.info, by.x="patient", by.y="TnumberAVL")
```

#Loop for barplots of RNC using different criteria
```{r}
method.rcn <- "radious46_k30"

cells.aux$rcn_id <- paste0("RCN",RCN.clusters[,2])
n.clust <- length(unique(cells.aux$rcn_id))
cells.aux$rcn_id <- factor(cells.aux$rcn_id,
                       levels=paste0("RCN",1:n.clust))
################################################
#Exporting data for visualization in Napari
for (slide in unique(cells.aux$cycif.slide)){
  sel.cell <- cells.aux %>%
              filter(cycif.slide == slide, cycif.core.id %in% cores.interest) %>%
              arrange(cycif.core.id, CellId)

  outout.folder.name <- paste0(project.folder, "/", slide, "/", output.napari, "_",method.rcn)
  dir.create(outout.folder.name)

  for (core in unique(sel.cell$cycif.core.id)){
    sel.cell.core <- sel.cell[which(sel.cell$cycif.core.id == core),]
    df <- data.frame(Core_Names=paste0(core,"_Probabilities_cytomask2"),
                   Cellid=sel.cell.core$CellId,
                   GlobalCellType=paste0("RCN",sel.cell.core$rcn_id))
    cell.id <- sapply(df$Cellid, function(x){
                      splitted <- strsplit(x,"_")[[1]]
                      cid <- splitted[5]
                      cid <- as.integer(gsub("c","",cid))
                      return(cid)
                    })
    df$Cellid <- cell.id
    df <- df %>% arrange(Cellid)
    write.table(df, file=paste0(outout.folder.name,"/",core,"_RCN_Napari.csv"),
                sep=",", row.names = FALSE)
  }
}
```


```{r}
cells.aux %>% group_by(imageid,rcn_id) %>%
    summarise(N=n()) %>% mutate(Proportion= N * 100 /sum(N)) %>% dplyr::filter(rcn_id=="RCN10c") %>%
    arrange(desc(Proportion))
```

#Identifying those patients with really low amount of cancer cells or super high amount of cancer cells
```{r}
cores.cancer.prop <- cells.aux %>% group_by(imageid, GlobalCellType) %>% 
  summarise(N=n()) %>% mutate(Proportion= N * 100 /sum(N)) %>% dplyr::filter(GlobalCellType=="Cancer") %>%
  arrange(desc(Proportion))

cores.stromal.prop <- cells.aux %>% group_by(imageid, GlobalCellType) %>% 
  summarise(N=n()) %>% mutate(Proportion= N * 100 /sum(N)) %>%
  dplyr::filter(GlobalCellType %in% "Stromal") %>%
  arrange(desc(Proportion))

high.cancer <- cores.cancer.prop$imageid[which(cores.cancer.prop$Proportion >= 85)]

high.stromal <- cores.stromal.prop$imageid[which(cores.stromal.prop$Proportion >= 85)]
```




#Loop for barplots of RNC after merging  using different criteria
```{r}
rcn.folder <- "Analysis_results/03_RCN_2024_metaclusters/"
input.prefix <- "Cluster_ids_metaclustersmerged_"
project.folder <- "D:/users/fperez/NKI_TMAs_AF/"
output.napari <- "RCN_Napari2024"
out.folder.name <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/03_RCN_2024_metaclusters/"
method.rcn <- "radius46_k30_merged"

# # #Order of RCN for plotting
RNC.order <- c("RCN01", "RCN02", "RCN03", "RCN04","RCN05","RCN06", "RCN07","RCN08", "RCN09", "RCN10",
                  "RCN11", "RCN12", "RCN13", "RCN14")

#Adding the merged RCNid to the cells
cells.aux$rcn_id <- RCN.annotations[RCN.clusters$spatial_kmeans,"RCN.id.collapsed.ordered"]
cells.aux$rcn_id <- factor(cells.aux$rcn_id, levels = rev(RNC.order))


# This code was used to ignore some samples for some of the first heatmaps created not used in publication
# #Ignoring cores with high amount of cancer or stromal
# cells.aux.clin2 <- cells.aux.clin[!(cells.aux.clin$imageid %in% high.cancer |
#                                       cells.aux.clin$imageid %in% high.stromal),]
# 
# #Ignoring patients with only one core
# patient.cores.nums <- cells.aux.clin2 %>% group_by(patient, imageid) %>%
#   summarise(N.cores=length(unique(imageid))) %>%
#   summarise(N.cores2=sum(N.cores))
# 
# patients.low.num.cores <- patient.cores.nums$patient[which(patient.cores.nums$N.cores2 < 2)]
# cells.aux.clin2 <- cells.aux.clin2[!cells.aux.clin2$patient %in% patients.low.num.cores,]
# 
# #Additionally ignore those patients with total amount of cancer cells very low
# tot.prop.cancer <- cells.aux.clin2 %>% group_by(patient, GlobalCellType) %>%
#                       summarise(N=n()) %>% mutate(Proportion = N * 100 /sum(N)) %>%
#                       dplyr::filter(GlobalCellType=="Cancer") %>% arrange(Proportion)
# 
# patients.low.cancer <- tot.prop.cancer$patient[which(tot.prop.cancer$Proportion < 10)]
# cells.aux.clin2 <- cells.aux.clin2[!cells.aux.clin2$patient %in% patients.low.cancer,]
# 
# #Ignoring RCN01c and RCN02c
# cells.aux.clin2 <- cells.aux.clin2[!cells.aux.clin2$rcn_id %in% c("RCN01c","RCN02c"),]
# 


# ###########################################
#Plotting barplot for abundance of clusters
barplot.rcn(cells.aux, method=method.rcn)

###########################################
#Plotting heatmap annotations with diversity and abundance of clusters
heatmap.diversity(cells.aux, method=method.rcn)

cells.aux.clin <- merge(cells.aux, M.profiles.clin, by="imageid", all.x = TRUE)

# ###########################################
# #Plotting heatmap annotations with diversity and abundance of clusters
heatmaps(cells.aux.clin, method=method.rcn)
```


```{r}
###############################################
#Exporting data for visualization in Napari
for (slide in unique(cells.aux$cycif.slide)[c(1:4,10)]){
   sel.cell <- cells.aux %>%
               dplyr::filter(cycif.slide == slide) %>% arrange(cycif.core.id, CellId)

  outout.folder.name <- paste0(project.folder, "/", slide, "/", output.napari, "_",method.rcn)
  dir.create(outout.folder.name)

  for (core in unique(sel.cell$cycif.core.id)){
    sel.cell.core <- sel.cell[which(sel.cell$cycif.core.id == core),]
    df <- data.frame(Core_Names=paste0(core,"_Probabilities_cytomask2"),
                   Cellid=sel.cell.core$CellId,
                   GlobalCellType=sel.cell.core$rcn_id)
    cell.id <- sapply(df$Cellid, function(x){
                      splitted <- strsplit(x,"_")[[1]]
                      cid <- splitted[5]
                      cid <- as.integer(gsub("c","",cid))
                      return(cid)
                    })
    df$Cellid <- cell.id
    df <- df %>% arrange(Cellid)
    write.table(df, file=paste0(outout.folder.name,"/",core,"_RCN_Napari.csv"),
                sep=",", row.names = FALSE)
  }
}
```



#Saving RCNid per cell
```{r}
write.table(cells.aux[,c(1,2,8,9,10,11,12,13)],
            file="D:/users/fperez/NKI_TMAs_AF/Tables/ALL_cells-RNCid_v20241122.csv", sep=",", row.names=FALSE)
```


#Saving proportion of RCN_id by image
```{r}
cells.by.image <- cells.aux.clin %>% group_by(imageid, rcn_id) %>% 
                    summarise(N=n()) %>%
                    mutate(Proportion = N /sum(N))
write.table(cells.by.image, file=paste0(out.folder.name,"RCN_by_image.csv"),sep=",", row.names = FALSE)
```





