---
title: "Data_preprocessing"
author: "Gayani Anandagoda"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages}
library(Seurat)
library(SeuratObject)
library(nichenetr)
library(multinichenetr)
library(SingleCellExperiment)
library(dplyr)
library(ggplot2)

```


## Do not have to run this script 
## Output of this script is saved as a Seurat object: "launonen_cancercell_allcells_allannot_Filtered.RDS" 
## Use "launonen_cancercell_allcells_allannot_Filtered.RDS" for "2_LR_Analysis.Rmd"


```{r load data}
# Load Seurat object and Ucell scores

scRNA_obj_all = readRDS("/run/user/1356082/gvfs/smb-share:server=group3.ad.helsinki.fi,share=h345/afarkkilab/Data/13_sciset/launonen_cancercell_allcells_allannot_v2.RDS")


Ucell_scores_csv = read.csv("/run/user/1356082/gvfs/smb-share:server=group3.ad.helsinki.fi,share=h345/afarkkilab/Projects/NKI/Single_Cell_LR_Annalysis/Input/correlation_15_genes_expression_vs_mhc2_scores_pseudobulk_epithelial.tsv", sep = "\t", header = TRUE)
```



```{r tertile calculation}

Ucell_scores = data.frame(unique(Ucell_scores_csv[,c(1,2)]))

tertiles = quantile(Ucell_scores$mhc2_sig_UCell, probs = c(1/3, 2/3))

mhcii_group = cut(Ucell_scores$mhc2_sig_UCell,
                  breaks = c(-Inf, tertiles[1], tertiles[2], Inf),
                  labels = c("Low","Mid","High"),
                  include.loweset = TRUE)
Ucell_scores_df = data.frame(Ucell_scores$sample,Ucell_scores$mhc2_sig_UCell,mhcii_group)

rownames(Ucell_scores_df)<-Ucell_scores_df$Ucell_scores.sample
```


### adding MHCII expression levels


```{r adding MHCII expression levels}
# MHCII expression levels : HIgh, Mid and low

scRNA_obj_all$MHCIIGroups_bulk<-"No_expression"

scRNA_obj_all$MHCIIGroups_bulk[scRNA_obj_all$sample%in%rownames(Ucell_scores_df)[Ucell_scores_df$mhcii_group=="Low"]]<-"Low"

scRNA_obj_all$MHCIIGroups_bulk[scRNA_obj_all$sample%in%rownames(Ucell_scores_df)[Ucell_scores_df$mhcii_group=="Mid"]]<-"Mid"

scRNA_obj_all$MHCIIGroups_bulk[scRNA_obj_all$sample%in%rownames(Ucell_scores_df)[Ucell_scores_df$mhcii_group=="High"]]<-"High"


```

# Meta Data  Exploration

# sample Meta Data

```{r}

combined_df_samples = data.frame(sample_ID = scRNA_obj_all$sample, MHCII_type = scRNA_obj_all$MHCIIGroups_bulk, treatment = scRNA_obj_all$treatment_stage, publication_ID = scRNA_obj_all$publication_sample_code)

df_grouped_samples = combined_df_samples %>% group_by(sample_ID,publication_ID,MHCII_type,treatment) %>% summarise(count = n())

df_grouped_samples$patient_ID = sapply(strsplit(df_grouped_samples$publication_ID, "_"),"[", 1)



#df_grouped_samples = Ucell_scores[Ucell_scores$sample%in%df_grouped_samples$sample_ID,]$mhc2_sig_UCell
#df_grouped_samples = data.frame(df_grouped_samples, Ucell_scores = Ucell_scores)
df_grouped_samples

```

```{r}
# sample counts

table(df_grouped_samples$MHCII_type)

# High           Low           Mid No_expression 
# 13            13            13            12


```

```{r}
# cell counts

combined_df_samples %>% group_by(MHCII_type) %>% summarise(count = n())

```

```{r}
#ggplot(df_grouped_samples, aes(x = reorder(patient_ID, -Ucell_scores), y = Ucell_scores, fill = MHCII_type)) #+
  #       geom_bar(stat = "identity") +
   #     theme(axis.text.x = element_text(angle = 90))
```

#######################################

# subsetting MHCII High and Low only

```{r subset seurat object2}
# subsetting MHCII High and Low only

scRNA_obj_all = subset(scRNA_obj_all, subset = MHCIIGroups_bulk %in% c("High","Low"))

table(scRNA_obj_all$treatment_stage)
# chemo-naive         IDS 
#      34657       33642 

```

```{r}
combined_df_samples = data.frame(sample_ID = scRNA_obj_all$sample, MHCII_type = scRNA_obj_all$MHCIIGroups_bulk, treatment = scRNA_obj_all$treatment_stage, publication_ID = scRNA_obj_all$publication_sample_code)

df_grouped_samples = combined_df_samples %>% group_by(sample_ID,publication_ID,MHCII_type,treatment) %>% summarise(count = n())

df_grouped_samples$patient_ID = sapply(strsplit(df_grouped_samples$publication_ID, "_"),"[", 1)



Ucell_scores= Ucell_scores[Ucell_scores$sample%in%df_grouped_samples$sample_ID,]$mhc2_sig_UCell
df_grouped_samples = data.frame(df_grouped_samples, Ucell_scores = Ucell_scores)
df_grouped_samples
```
```{r}
table(df_grouped_samples$MHCII_type)
#combined_df_samples %>% group_by(MHCII_type) %>% summarise(count = n())


# High  Low 
# 13   13
```



####################################

### editing cell types

```{r editing cell types}

combined_df_cell_types = data.frame(cell_type_detailed = scRNA_obj_all$cell_type_detailed, cell_type = scRNA_obj_all$cell_type, cell_type_cycif = scRNA_obj_all$cell_type_cycif)


combined_df_cell_types_mutated = combined_df_cell_types %>% 
  
  mutate(cell_type_cycif = case_when(cell_type %in% c("DC1","DC2","pDC","Migratory DCs") ~ "DC",
                                     cell_type %in% c("CD16+ NK cells","CD16- NK cells","NK cells") ~ "NK",
                                     TRUE ~ cell_type_cycif ))

combined_df_cell_types_mutated = combined_df_cell_types_mutated %>% 
  
  mutate(cell_type_cycif = case_when(cell_type_detailed %in% c("macrophages_CD163_high") ~ "macrophages_CD163_high",
                                     cell_type_detailed %in% c("macrophages_ITGAX_high") ~ "macrophages_ITGAX_high",
                                     TRUE ~ cell_type_cycif ))

df_grouped_cell_types = combined_df_cell_types_mutated %>% group_by(cell_type_cycif,cell_type,cell_type_detailed) %>% summarise(count = n())

df_grouped_cell_types
```


```{r}
combined_df_cell_types_mutated %>% group_by(cell_type_cycif) %>% summarise(count = n())
```


```{r adding edited cell types}
scRNA_obj_all$edited_cell_type = combined_df_cell_types_mutated$cell_type_cycif

```

# save filtered Seurat object

```{r save object}
# save filtered Seurat object

saveRDS(scRNA_obj_all, file = "/run/user/1356082/gvfs/smb-share:server=group3.ad.helsinki.fi,share=h345/afarkkilab/Projects/NKI/Single_Cell_LR_Annalysis/Input/launonen_cancercell_allcells_allannot_Filtered_MHCII_HIgh_and_low.RDS")

```


```{r}
df_MHCII_groups = readRDS("/run/user/1356082/gvfs/smb-share:server=group3.ad.helsinki.fi,share=h345/afarkkilab/Projects/NKI/Single_Cell_LR_Annalysis/input/MHCII_groups_per_sample.RDS")
```

```{r}
table(df_MHCII_groups$mhc2_group)

```

```{r}
table(Ucell_scores_df$mhcii_group)
```

