---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(survival)
library(survminer)
library(gridExtra)
library(grid)
library(forestploter)
library(reshape2)
library(ComplexHeatmap)
```

#Defining min_max function
```{r}
min.max.norm <- function(x, na.rm = TRUE) {
    return((x- min(x)) /(max(x)-min(x)))
}

median.center <- function(x, na.rm = TRUE) {
    return((x - median(x))/sd(x))
}

#This is for relative proportions
p.vals.markers2 <- function(data, Profile.test="HRD", y.val=25, alt.hypothesis="greater"){
       test.dat <- data$Proportion[data$Molecular.profile2 == Profile.test]
       p.vals <- sapply(unique(data$Molecular.profile2), function(x){
                mol.dat <- data$Proportion[data$Molecular.profile2 == x]
                wtest <- wilcox.test(test.dat, mol.dat, alternative=alt.hypothesis)
                wtest$p.value
       })
      #Organizing p-values in data.frame for plotting using geom_text
      p.vals.dataframe <- data.frame(pval=p.vals,
                           Molecular.profile2=unique(data$Molecular.profile2),
                           Y.val=y.val)
      return(p.vals.dataframe)
}
```



#Reading input
```{r}
#Reading input cell proportion table by patient
#patient.cell.prop <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Analysis_results/01_Cell-proportions_Heatmaps2/tCycif_cell-proportions-patients.csv", sep=",", header=TRUE)

core.cell.prop <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Analysis_results/01_Survival_Proportions_November_medians/Proportions_by_core.csv", sep=",", header = TRUE)

#Reading clinical info
core.patient.slide <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Clinical_intersection/All-slides_cycif2samples.txt",
                                 header = TRUE, sep="\t")

#Reading clinical info
clinical.info <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Clinical_data/T-CycIF.update.csv", sep=",", header = TRUE)

Mol.profiles.patients <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Tables/Molecular_profiles_patients_20231208.csv",
            sep=",", header=TRUE)

#out.put.folder <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/01_Survival_Proportions_November_min3cores/"
out.put.folder <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/01_Survival_Proportions_November_medians//"

dir.create(out.put.folder, showWarnings = FALSE)
```

#Selecting only patients with at leats 3 cores
```{r}
#Patients selected
# sel.patients <- core.cell.prop %>% group_by(patient) %>%
#                   summarise(n.cores=length(unique(cycif.core.id))) %>%
#                   filter (n.cores >=3 ) %>% pull(patient)
# 
# core.cell.prop <- core.cell.prop[core.cell.prop$patient %in% sel.patients,]
```


#Calculating cell proportions by patient
```{r}
patient.cell.prop <- core.cell.prop %>% group_by(patient, GlobalCellType) %>% 
                      summarise(Proportion = median(Proportion)) %>% as.data.frame() %>%
                      pivot_wider(names_from = GlobalCellType, values_from = Proportion)
patient.cell.prop[is.na(patient.cell.prop)] <- 0 
patient.cell.prop
```



#Merging clinical info and data
```{r}
#Selecting relevant columns from the clinical info
interesting.clin.columns <- c("TnumberAVL", "centre","refage","incyear","timelastfu","daystoprogression",
                              "finalstatus","progression","figo_stage")

clinical.info.sel <- clinical.info %>% select(interesting.clin.columns)

#Merging the cell proportions and selected clinical info
patient.cell.prop.clin <- merge(patient.cell.prop, clinical.info.sel,
                                by.x = "patient", by.y="TnumberAVL")

#Merging with molecular profile
patient.cell.prop.clin2 <- merge(patient.cell.prop.clin, Mol.profiles.patients, by="patient")

#Re-naming therapy sequence as IDS and PDS
patient.cell.prop.clin2$therapy_sequence <- ifelse(grepl("NACT", patient.cell.prop.clin2$therapy_sequence), "IDS","PDS")

#Factor with levels are for the survival analysis. HRP is used as reference in cox models
levels.molecular <- c("HRP", "BRCAmut/met", "HRD", "CCNE1amp", "Other")
patient.cell.prop.clin2$Molecular.profile2 <- factor(patient.cell.prop.clin2$Molecular.profile2,
                                                   levels = levels.molecular)

#Converting survival info from days to months
patient.cell.prop.clin2$timelastfu <- patient.cell.prop.clin2$timelastfu/30.4
patient.cell.prop.clin2$daystoprogression <- patient.cell.prop.clin2$daystoprogression/30.4

#Structuring data
patient.cell.prop.clin2$figo_stage2 <- NA
patient.cell.prop.clin2$figo_stage2[patient.cell.prop.clin2$figo_stage == "FIGO II"] <- 2
patient.cell.prop.clin2$figo_stage2[patient.cell.prop.clin2$figo_stage == "FIGO III"] <- 3
patient.cell.prop.clin2$figo_stage2[patient.cell.prop.clin2$figo_stage == "FIGO IV"] <- 4

#Structuring data
patient.cell.prop.clin2$refage2 <- NA
patient.cell.prop.clin2$refage2[patient.cell.prop.clin2$refage < 50] <- 0
patient.cell.prop.clin2$refage2[patient.cell.prop.clin2$refage >= 50 & patient.cell.prop.clin2$refage < 60] <- 1
patient.cell.prop.clin2$refage2[patient.cell.prop.clin2$refage >= 60 & patient.cell.prop.clin2$refage < 70] <- 2
patient.cell.prop.clin2$refage2[patient.cell.prop.clin2$refage >= 70 & patient.cell.prop.clin2$refage < 80] <- 3
patient.cell.prop.clin2$refage2[patient.cell.prop.clin2$refage >= 80] <- 4
```


#Calculating cut-off points (median values) per cell type to definy high abundance of cell.type for survival analysis
```{r}
vars.all = c("CD8.CD45RO.T.cells","CD8.PD1.T.cells","CD8.effector.T.cells",
             "B.cells","CD4.T.cells","CD4.PD1.T.cells","CD4.CD45RO.T.cells",
             "CD68.MP", "CD11c.MY","CD15.MY","CD163.MP","CD207.MY","T.regs","Other.MY")


cutpoints <- sapply(vars.all, function(x){
              var = x
              #cutval.var <- surv_cutpoint(cells_percentages.mc, time = "daystoprogression",
              #                            event = "progression", variables=var)
              #cutpoint <- cutval.var$cutpoint[1]
              #names(cutpoint) <- NULL
              median.var <- median(patient.cell.prop.clin2[,var])
              cutpoint <- median.var
              #If the cut.point is equal to 0 then take the mean
              if ((cutpoint) == 0){
                print(var)
                cutpoint <- mean(patient.cell.prop.clin2[,var])
              }
              return(cutpoint)
})
```


#Performing Kapplan-Meir plots for the abundance of each cell.type, according to therapy sequence
```{r}
output.folder = "D:/users/fperez/NKI_TMAs_AF/Analysis_results/01_Survival_Proportions_November_medians/"
#output.folder <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/01_Survival_Proportions_November_min3cores/"
dir.create(paste0(output.folder))

suffix <- c("PDS","IDS")
list.class <- list(c("PDS"),
                   c("IDS"))


for (j in 1:length(suffix)){
  data.in <- patient.cell.prop.clin2[patient.cell.prop.clin2$therapy_sequence %in% list.class[[j]],]
  lastcol <- ncol(data.in)
  colnames(data.in)[2] <- "B.cells"
  
  plots.OS <- lapply(vars.all, function(x){
  vars = x
  cut.off.value <- cutpoints[names(cutpoints) == vars]
  cat.var <- as.factor(ifelse(data.in[,vars] >  cut.off.value, "high","low"))
  data.in[,lastcol] <- cat.var
  colnames(data.in)[lastcol] <- "aux"
  fit <- survfit(Surv(timelastfu, finalstatus) ~ aux, data = data.in)
  logrank = surv_pvalue(fit, data.in)
  pval=logrank$pval.txt
  p <- ggsurvplot(fit, data = data.in, risk.table = TRUE, p.val=TRUE, conf.int = TRUE,
                  font.legend = c(18, "plain"), font.tickslab = c(16, "plain"),
                  font.x = c(18, "plain"), font.y = c(18, "plain"),
                  tables.y.text = FALSE,  break.time.by=20, xlim = c(0,80),
                  legend.labs=c(paste0(vars, "+"),paste0(vars, "-")),
                  palette = c("#ff5a36", "#4d4dff")) + 
        guides(colour = guide_legend(nrow = 2))
  p <- p + ylab("OS probability") + xlab("Time (months)")
  p$plot <- p$plot + theme(legend.key.width = unit(3, "line"), plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"), legend.key.height = unit(1.5, "line"))
  p$plot <- p$plot + ggplot2::annotate("text", x = 65, y = 0.85, label =pval, size = 5)
  p$table <- p$table + theme(plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"))
  plots <- list(plot=p$plot, table=p$table)
  return(plots)
  })
  fig.OS <- lapply(1:length(plots.OS) , function(x){plots.OS[[x]]$plot})
  tables.OS <- lapply(1:length(plots.OS) , function(x){plots.OS[[x]]$table})
  
  
  plots.PFI <- lapply(vars.all, function(x){
  vars = x
  cut.off.value <- cutpoints[names(cutpoints) == vars]
  cat.var <- as.factor(ifelse(data.in[,vars] >  cut.off.value, "high","low"))
  data.in[,lastcol] <- cat.var
  colnames(data.in)[lastcol] <- "aux"
  fit <- survfit(Surv(daystoprogression, progression) ~ aux, data = data.in)
  logrank = surv_pvalue(fit, data.in)
  pval=logrank$pval.txt
  p <- ggsurvplot(fit, data = data.in, risk.table = TRUE, p.val=TRUE, conf.int = TRUE,
                  font.legend = c(18, "plain"), font.tickslab = c(16, "plain"),
                  font.x = c(18, "plain"), font.y = c(18, "plain"),
                  tables.y.text = FALSE,  break.time.by=20, xlim = c(0,80),
                  legend.labs=c(paste0(vars, "+"),paste0(vars, "-")),
                  palette = c("#ff5a36", "#4d4dff")) + 
        guides(colour = guide_legend(nrow = 2))
  p <- p + ylab("PFI probability") + xlab("Time (months)")
  p$plot <- p$plot + theme(legend.key.width = unit(3, "line"), plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"), legend.key.height = unit(1.5, "line"))
  p$plot <- p$plot + ggplot2::annotate("text", x = 65, y = 0.85, label =pval, size = 5)
  p$table <- p$table + theme(plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"))
  plots <- list(plot=p$plot, table=p$table)
  return(plots)
  })
  fig.PFI <- lapply(1:length(plots.PFI) , function(x){plots.PFI[[x]]$plot})
  tables.PFI <- lapply(1:length(plots.PFI) , function(x){plots.PFI[[x]]$table})
  
  ggsave(file=paste0(output.folder,suffix[j], "_OS.svg"), arrangeGrob(grobs = fig.OS, ncol = 2),
         width = 23, height = 50, units = "cm")
  
  ggsave(file=paste0(output.folder,suffix[j], "_OS-table.svg"), arrangeGrob(grobs = tables.OS, ncol = 2),
         width = 23, height = 22, units = "cm")
  
  ggsave(file=paste0(output.folder,suffix[j],"_PFI.svg"), arrangeGrob(grobs = fig.PFI, ncol = 2),
         width = 23, height = 50, units = "cm")
  
  ggsave(file=paste0(output.folder,suffix[j],"_PFI-table.svg"), arrangeGrob(grobs = tables.PFI, ncol = 2),
         width = 23, height = 22, units = "cm")
}
```


#Function to calculate cox hazzard ratios
```{r}
#Indicate the variables for which is going to be calculated the cox model
Cox.regresion.variables2 <- function(data, variables, formula='Surv(timelastfu, finalstatus)~'){

   N <- sapply(variables, function(x){length(which(data[,x] == 1))})
   N <- as.data.frame(N)
   Prop <- paste0(round((N[,1] * 100 /nrow(data)),0), "%")

   univ_formulas <- sapply(variables,
                           function(x) as.formula(paste(formula, x)))

   univ_models <- lapply( univ_formulas, function(x){coxph(x, data = data, method = "exact")})

   univ_results <- lapply(univ_models,
                          function(x){
                             x <- summary(x)
                             positive <- x$nevent
                             prop.positive <- paste0(round((x$nevent) * 100 /x$n,0),"%")
                             #p.value <- signif(x$wald["pvalue"], digits=2) #p.value for the model
                             wald.test <- signif(x$wald["test"], digits=2)
                             lastrow <- nrow(x$coef) #When using several covariables in the same regression, take the last one used
                             p.value <- signif(x$coef[lastrow, "Pr(>|z|)"], digits=2) #p.value for last covariable
                             beta <- signif(x$coef[lastrow, 1], digits=2);#coeficient beta
                             HR <- signif(x$coef[lastrow, 2], digits=2);#exp(beta)
                             HR.confint.lower <- signif(x$conf.int[lastrow,"lower .95"], 2)
                             HR.confint.upper <- signif(x$conf.int[lastrow,"upper .95"],2)
                             res <- c(HR, p.value, HR.confint.lower,HR.confint.upper)
                             names(res)<-c("HR", "p.value", "HR.confint.lower", "HR.confint.upper")
                             return(res)
                          })
   res <- t(as.data.frame(univ_results, check.names = FALSE))
   result <- as.data.frame(res)
   result2 <- cbind(N, Prop, result)

   return(result2)
}
```


#Calculating individual cox hazzard ratios per cell.type and including the Molecular profile in each
```{r}
tm <- forest_theme(core=list(bg_params=list(fill = c("grey95", "#eaf8f5", "white"))))


formulas <- c('Surv(timelastfu, finalstatus)~ Molecular.profile2 + refage2  + figo_stage2 +',
              'Surv(daystoprogression, progression)~ Molecular.profile2 + refage2  + figo_stage2 +')

formula.suffix <- c('OS_mprofile_age_figo', 'PFS_mprofile_age_figo')

lymphocites <- c("CD4.T.cells","CD4.CD45RO.T.cells", "CD4.PD1.T.cells", "T.regs",
                 "CD8.effector.T.cells","CD8.CD45RO.T.cells","CD8.PD1.T.cells", "B.cells", "CD15.MY",
                 "CD11c.MY", "Other.MY")

for (index in 1:length(formulas)){
  #Calculating cox hazards in in two groups therapy sequence
  therapies <- c("IDS","PDS")
  cox.therpies <- lapply(therapies, function(x){
          dat <- patient.cell.prop.clin2[(patient.cell.prop.clin2$therapy_sequence %in% x),]
          
          #If proportion of cell.type is equal or above the median in the sampled population, then is equal to 1
          for (j in lymphocites){
           cut.off.value <- cutpoints[names(cutpoints) == j]
           dat[,j] <- ifelse(dat[,j] > cut.off.value,1,0)
          }
          #Calculating cox models
          Cox.lymphocites <- Cox.regresion.variables2(dat, lymphocites, formula=formulas[index])
          Cox.lymphocites$Treatment <- x
          Cox.lymphocites$Cell.type <- rownames(Cox.lymphocites)
          return(Cox.lymphocites)
  })
  
  
  #Reordering the cox results in an specific way necessary for the package forestplot
  spacer <- paste(rep(" ", 20), collapse = " ") #For the forest plot table
  cox.groups <- rbind(cox.therpies[[2]], cox.therpies[[1]])
  df.forest <- NULL
  for (i in unique(cox.groups$Cell.type)){
        row.class <- data.frame(Cell.type =i, N="", HR="", space=spacer, p.value=NA, HR.confint.lower=NA,
                                HR.confint.upper=NA, HR2=NA)
        set.df <- cox.groups[cox.groups$Cell.type %in% i,]
        set.df$space <- spacer
        set.df$HR2 <- set.df$HR
        set.df.reordered <- set.df %>% select(Treatment, N, HR, space, p.value, HR.confint.lower, HR.confint.upper, HR2)
        colnames(set.df.reordered) <- colnames(row.class)
        row.celltype <- rbind(row.class, set.df.reordered)
        df.forest <- rbind(df.forest, row.celltype)
  }
  colnames(df.forest)[4] <- ""
  colnames(df.forest)[3] <- "HR"
  
  #Adding * to pvalues lower than 0.05
  low.pvals <- which(df.forest$p.value < 0.05)
  df.forest$p.value[low.pvals] <- paste0(df.forest$p.value[low.pvals],"*")
  df.forest$p.value[which(is.na(df.forest$p.value))] = ""
  colnames(df.forest)[5] <- "P.val"
  
  #indent the subgroup if there is a number in the HR2 column
  df.forest$Cell.type <- ifelse(is.na(df.forest$HR2), 
                                df.forest$Cell.type,
                                #unlist(sapply(df.forest$Cell.type, function(x){cat("\t",x)})))
                                paste0("-", df.forest$Cell.type))
  
  
  p <- forest(df.forest[,c(1:5)],
              est = df.forest$HR2,
              lower = df.forest$HR.confint.lower, 
              upper = df.forest$HR.confint.upper,
              sizes = 0.5,
              ci_column = 4,
              ref_line = 1,
              xlim = c(0, 1.5),
              arrow_lab = c("Lower risk", "Higher risk"),
              ticks_at = c(0.5, 1, 1.5),
              theme = tm)
  print(p)
  
  ggplot2::ggsave(filename = paste0(output.folder, "Forest_plot_", formula.suffix[index], ".svg"), plot = p,
                   width = 15, height = 26, units = "cm", device = svglite::svglite)
  
  ggplot2::ggsave(filename = paste0(output.folder, "Forest_plot_", formula.suffix[index], ".png"), plot = p,
                   width = 15, height = 26, units = "cm")
}
```

####Forest plots with individual cutoffs for PDS and IDS samples
#Calculating cut-off points (median values) per cell type to definy high abundance of cell.type for survival analysis
```{r}
vars.all = c("CD8.CD45RO.T.cells","CD8.PD1.T.cells","CD8.effector.T.cells",
             "B.cells","CD4.T.cells","CD4.PD1.T.cells","CD4.CD45RO.T.cells",
             "CD11c.MY","CD15.MY","CD163.MP","CD207.MY","T.regs","Other.MY")

therapies <- c("IDS","PDS")
cutpoints.therapy <-  lapply(therapies, function(j){
    data.sel <- patient.cell.prop.clin2[patient.cell.prop.clin2$therapy_sequence == j,]
    sapply(vars.all, function(x){
                  var = x
                  median.var <- median(data.sel[,var])
                  cutpoint <- median.var
                  #If the cut.point is equal to 0 then take the mean
                  if ((cutpoint) == 0){
                    print(var)
                    cutpoint <- mean(patient.cell.prop.clin2[,var])
                  }
                  return(cutpoint)
    })
  })
names(cutpoints.therapy) <- therapies
cutpoints.therapy
```


###Cox models indivual cutoffs for IDS and PDS
```{r}
formulas <- c('Surv(timelastfu, finalstatus)~ Molecular.profile2 + refage2  + figo_stage2 +',
              'Surv(daystoprogression, progression)~ Molecular.profile2 + refage2  + figo_stage2 +')

formula.suffix <- c('OS_mprofile_age_figo', 'PFS_mprofile_age_figo')

formula.short.name <- c('OS', 'PFS')


for (index in 1:length(formulas)){
  for (t in therapies){
    #Calculating cox hazards in in two groups therapy sequence
    dat <- patient.cell.prop.clin2[(patient.cell.prop.clin2$therapy_sequence %in% t),]
            
    #If proportion of cell.type is equal or above the median in the sampled population, then is equal to 1
    for (j in lymphocites){
     cut.off.value <- cutpoints.therapy[[t]][names(cutpoints.therapy[[t]]) == j]
     dat[,j] <- factor(ifelse(dat[,j] > cut.off.value,1,0),levels=c(0,1))
    }
    #Calculating cox models
    Cox.lymphocites <- Cox.regresion.variables2(dat, lymphocites, formula=formulas[index])
    Cox.lymphocites$Treatment <- t
    Cox.lymphocites$Cell.type <- rownames(Cox.lymphocites)
  
    #Reordering the cox results in an specific way neccesary for the package forestplot
    spacer <- paste(rep(" ", 20), collapse = " ") #For the forest plot table
    cox.groups <- Cox.lymphocites
    df.forest <- NULL
    for (i in unique(cox.groups$Cell.type)){
          set.df <- cox.groups[cox.groups$Cell.type %in% i,]
          set.df$space <- spacer
          set.df$HR2 <- set.df$HR
          set.df.reordered <- set.df %>% select(Cell.type, N, HR, space, p.value, HR.confint.lower, HR.confint.upper, HR2)
          df.forest <- rbind(df.forest, set.df.reordered)
    }
    colnames(df.forest)[4] <- ""
    colnames(df.forest)[3] <- "HR"
    
    #Adding * to pvalues lower than 0.05
    low.pvals <- which(df.forest$p.value < 0.05)
    df.forest$p.value[low.pvals] <- paste0(df.forest$p.value[low.pvals],"*")
    df.forest$p.value[which(is.na(df.forest$p.value))] = ""
    colnames(df.forest)[5] <- "P.val"
    
    p <- forest(df.forest[,c(1:5)],
                est = df.forest$HR2,
                lower = df.forest$HR.confint.lower, 
                upper = df.forest$HR.confint.upper,
                title = paste(formula.short.name[index], t),
                sizes = 0.5,
                ci_column = 4,
                ref_line = 1,
                xlim = c(0, 1.5),
                arrow_lab = c("Lower risk", "Higher risk"),
                ticks_at = c(0.5, 1, 1.5))
    #print(p)
    
    ggplot2::ggsave(filename = paste0(output.folder, "Forest_plot_", formula.suffix[index], "_" , t, ".svg"), plot = p,
                     width = 15, height = 11, units = "cm")
    ggplot2::ggsave(filename = paste0(output.folder, "Forest_plot_", formula.suffix[index], "_" , t, ".png"), plot = p,
                     width = 15, height = 11, units = "cm")
    dev.off()
  }
}
```



###Create multivariable cox model with all the variables significan in the univariable and adding molecular profiles 
```{r}
formulas <- c('Surv(timelastfu, finalstatus) ~ .',
              'Surv(daystoprogression, progression)~ .')

formula.suffix <- c('OS_mprofile_age_figo', 'PFS_mprofile_age_figo')

vars.for.suv <- list(c("timelastfu","finalstatus"),
                     c("daystoprogression","progression"))

short.surv.name <- c("OS","PFS")

for (index in 1:length(formulas)){
  for (c in c("IntraCutoff", "GlobalCutoff")){
    for (t in therapies){
      dat <- patient.cell.prop.clin2[(patient.cell.prop.clin2$therapy_sequence %in% t),]
      for (j in lymphocites){
      if (c == "GlobalCutoff"){
       cut.off.value<- cutpoints[names(cutpoints) == j]
      }else{
       cut.off.value <- cutpoints.therapy[[t]][names(cutpoints.therapy[[t]]) == j]
      }
       dat[,j] <- factor(ifelse(dat[,j] > cut.off.value,"high","low"), levels=c("low","high"))
      }
      dat <- dat %>% select(c("Molecular.profile2", "refage2" ,  "figo_stage2", lymphocites, vars.for.suv[[index]]))
      colnames(dat)[1] <- "M.profile"
      
      model <- coxph(as.formula(formulas[index]), data=dat)
      p <- ggforest(model, data = dat, main=paste0(short.surv.name[index],"_", t,"_", c))
      ggplot2::ggsave(filename = paste0(output.folder, "Forest_plot_multivarCox_", formula.suffix[index], "_",
                                        c, "_", t, ".png"),
                      plot = p,  width = 15, height = 20, units = "cm")
    }
  }
  
  
  dat <- patient.cell.prop.clin2
  for (j in lymphocites){
    cut.off.value <- cutpoints[names(cutpoints) == j]
    #cut.off.value <- cutpoints.therapy[[t]][names(cutpoints.therapy[[t]]) == j]
    dat[,j] <- factor(ifelse(dat[,j] > cut.off.value,"high","low"), levels=c("low","high"))
  }
  
  
  dat <- dat %>% select(c("Molecular.profile2", "therapy_sequence", "figo_stage2", "refage2",
                          lymphocites, vars.for.suv[[index]]))
  colnames(dat)[1] <- "M.profile"
  model <- coxph(as.formula(formulas[index]), data=dat)
  p <- ggforest(model, data = dat, main=paste0(short.surv.name[index], " Multi-therapy Global"))
  ggplot2::ggsave(filename = paste0(output.folder, "Forest_plot_multivarCox_", formula.suffix[index], ".png"), plot = p,
                     width = 15, height = 20, units = "cm")
}
```

```{r}

# lymphocites <- c("CD4.T.cells","CD4.CD45RO.T.cells", "CD4.PD1.T.cells", "T.regs",
#                  "CD8.effector.T.cells","CD8.CD45RO.T.cells","CD8.PD1.T.cells", "B.cells", "CD15.MY",
#                  "CD11c.MY", "Other.MY")


lymphocites2 <- c( "CD11c.MY")

formulas <- c('Surv(timelastfu, finalstatus) ~ .',
              'Surv(daystoprogression, progression)~ .')

formula.suffix <- c('OS_mprofile_age_figo', 'PFS_mprofile_age_figo')

vars.for.suv <- list(c("timelastfu","finalstatus"),
                     c("daystoprogression","progression"))

short.surv.name <- c("OS","PFS")

dat <- patient.cell.prop.clin2[patient.cell.prop.clin2$therapy_sequence == "PDS",]
for (j in lymphocites){
   cut.off.value<- cutpoints[names(cutpoints) == j]
   dat[,j] <- factor(ifelse(dat[,j] > cut.off.value,"high","low"), levels=c("low","high"))
#    dat[,j] <- ntile(dat[,j], 2)
}


dat <- dat %>% select(c("Molecular.profile2", "figo_stage2", "refage2",
                        lymphocites2, "timelastfu", "finalstatus"))

colnames(dat)[1] <- "M.profile"

model <- coxph(Surv(timelastfu, finalstatus) ~ ., data=dat)

p <- ggforest(model, data = dat, main=paste0("OS Multi-therapy Global"))

print(p)
```
```{r}
dat <- patient.cell.prop.clin2 %>% filter(therapy_sequence == "PDS") %>% select(lymphocites)
Heatmap(cor(dat), name="Correlation", column_title="PDS")

dat <- patient.cell.prop.clin2 %>% filter(therapy_sequence == "IDS") %>% select(lymphocites)
Heatmap(cor(dat), name="Correlation", column_title="IDS")

dat <- patient.cell.prop.clin2 %>% select(lymphocites)
Heatmap(cor(dat), name="Correlation", column_title="All cohort")
```


```{r}
test.ph <- cox.zph(model)
test.ph
#ggcoxzph(test.ph)
```



###Next lines to calculate median the proportion of celltypes by Molecular profiles
```{r}
#Re-organizing data
melt.patient.cell.prop <- melt(patient.cell.prop)

names(melt.patient.cell.prop) <- c("patient","GlobalCelltype","Proportion")
melt.patient.cell.prop.clin  <- merge(melt.patient.cell.prop, clinical.info.sel, by.x="patient", by.y="TnumberAVL")
melt.patient.cell.prop.clin <- merge(melt.patient.cell.prop.clin, Mol.profiles.patients, by="patient")
melt.patient.cell.prop.clin$therapy_sequence <- ifelse(grepl("NACT", melt.patient.cell.prop.clin$therapy_sequence),
                                                       "IDS","PDS")


melt.patient.cell.prop.clin$GlobalCelltype <- as.vector(melt.patient.cell.prop.clin$GlobalCelltype)

#melt.patient.cell.prop.clin$GlobalCelltype[grep("CD8", melt.patient.cell.prop.clin$GlobalCelltype)] <- "CD8.T.cells"
#melt.patient.cell.prop.clin$GlobalCelltype[grep("CD4", melt.patient.cell.prop.clin$GlobalCelltype)] <- "CD4.T.cells"

#Ignoring Others from heatmap
melt.patient.cell.prop.clin <- melt.patient.cell.prop.clin[melt.patient.cell.prop.clin$GlobalCelltype != "Other",]
melt.patient.cell.prop.clin <- melt.patient.cell.prop.clin[melt.patient.cell.prop.clin$Molecular.profile2 != "Other",]


#Calculating the median proportion by molecular profile
med.prop.profile <- melt.patient.cell.prop.clin %>%  group_by(Molecular.profile2, therapy_sequence, GlobalCelltype) %>%
                    summarise(Prop=median(Proportion) * 100)

#Changing data to matrix 
mt.prop.profile <- med.prop.profile %>% pivot_wider(names_from = GlobalCelltype, values_from = Prop)
mt.prop.profile <- as.data.frame(mt.prop.profile)
mt.prop.profile[is.na(mt.prop.profile)] <- 0

#Scaling proportions using z-score
mt.prop.scaled <- as.data.frame(scale(mt.prop.profile[,c(-1,-2)]))
mt.prop.scaled$Molecular.profile2 <- mt.prop.profile$Molecular.profile2
mt.prop.scaled$therapy_sequence <- mt.prop.profile$therapy_sequence

#Preparing data for creating complex heatmap
M.profile.order <- mt.prop.scaled$Molecular.profile2 
therapy.order <- mt.prop.scaled$therapy_sequence 
mt.prop.scaled <- mt.prop.scaled %>% #select(CD4.T.cells, CD8.T.cells, T.regs, B.cells, CD11c.MY,
                                    #        CD163.MP, CD68.MP, Other.MY) %>%
                                  select(CD4.T.cells, CD4.CD45RO.T.cells, CD4.PD1.T.cells, CD8.effector.T.cells,
                                         CD8.CD45RO.T.cells, CD8.PD1.T.cells, T.regs, B.cells,
                                         CD11c.MY, CD163.MP, CD68.MP, Other.MY, Stromal, Cancer)    %>%
                  as.matrix()


colours <- list(Therapy=c("IDS"="black", "PDS"="grey90"),
                 M.profile=c("BRCAmut/met" = "#FF0000", "HRD" = "#AB1717", "CCNE1amp"="darkblue", "HRP"="#1E90FF",
                             "Other"="grey60"))

ha = rowAnnotation(Therapy=therapy.order, M.profile=M.profile.order, col=colours)

hmap <- Heatmap(mt.prop.scaled, right_annotation = ha, cluster_columns = FALSE, cluster_rows = FALSE,
        name="Median proportion \n (Z-score)")
svg(paste0(output.folder, "Heatmap_Mprofile_ImmuneStromalCancer_prop.svg"), width=8, height=5)
draw(hmap)
dev.off()



#Scaling by therapy sequence
mt.prop.scaled <- mt.prop.profile %>% group_by(therapy_sequence) %>% 
                   mutate(across(where(is.numeric), ~ as.numeric(scale(.)))) %>%
                   as.data.frame()

for (t in c("PDS","IDS")){
  #Preparing data for creating complex heatmap
  M.profile.order <- mt.prop.scaled %>% filter(therapy_sequence == {{t}}) %>%
                                    pull(Molecular.profile2)
  mt.prop.scaled.sel <- mt.prop.scaled %>% filter(therapy_sequence == {{t}}) %>%
                                    select(CD4.T.cells, CD4.CD45RO.T.cells, CD4.PD1.T.cells, CD8.effector.T.cells,
                                           CD8.CD45RO.T.cells, CD8.PD1.T.cells, T.regs, B.cells,
                                           CD11c.MY, CD163.MP, CD68.MP, Other.MY, Stromal, Cancer)    %>%
                    as.matrix()
  
  
  colours <- list(M.profile=c("BRCAmut/met" = "#FF0000", "HRD" = "#AB1717", "CCNE1amp"="darkblue", "HRP"="#1E90FF",
                               "Other"="grey60"))
  
  ha = rowAnnotation(M.profile=M.profile.order, col=colours)
  
  hmap <- Heatmap(mt.prop.scaled.sel, right_annotation = ha, cluster_columns = FALSE, cluster_rows = FALSE,
          name="Median proportion \n (Z-score)", column_title = paste0(t," samples"))
  pdf(paste0(output.folder, "Heatmap_Mprofile_ImmuneStromalCancer_prop_", t, ".pdf"), width=8, height=5)
  draw(hmap)
  dev.off()
}
```

#Making violin plots 
```{r}
patient.cell.prop.clin2$therapy_sequence[patient.cell.prop.clin2$therapy_sequence == "NACT" ] <- "IDS"

patient.cell.prop.clin2$Molecular.profile2 = as.vector(patient.cell.prop.clin2$Molecular.profile2)
cell.types.interest  <- list(c("CD163.MP"))

columns.extra <- c("patient","Molecular.profile2","therapy_sequence")

col <- rowSums(patient.cell.prop.clin2[cell.types.interest[[1]]])
dat <- patient.cell.prop.clin2[,c(columns.extra)]
dat$Proportion <- col

p.vals.markers2(subset(dat, therapy_sequence=="PDS"), Profile.test="CCNE1amp", y.val = 0.35)

p <- ggplot(dat, aes(x=Molecular.profile2,y=Proportion)) +  facet_wrap("therapy_sequence") + 
     # ylim(0,0.13) +
      geom_point(aes(col=Molecular.profile2), position= position_jitter(width= .3, height = 0), size= 1.5, alpha = 0.5) +
      geom_violin(fill=NA) + geom_boxplot(fill=NA, width=0.2, outlier.shape = NA) +
      theme(strip.placement = "outside",strip.background = element_blank(),
               axis.text.x=element_text(size=rel(1.2), colour = "black"), axis.text.y=element_text(size=rel(1.2)),
               legend.text=element_text(size=rel(1.2)),
               strip.text.x = element_text(size=rel(1.2)),
               legend.title=element_text(size=rel(1)), axis.title=element_text(size=rel(1.2)),
               panel.spacing = unit(1, "lines"),
               panel.border = element_rect(linetype = "solid", fill = NA),
               panel.background = element_rect(fill = "white"),
               panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "lightgrey"),
               panel.grid.minor = element_blank(),
               panel.grid.major.x = element_blank(),
               panel.grid.minor.x = element_blank())
print(p)
```

