---
title: "R Notebook"
author: Aleksandra
output: html_notebook
---

### Load libraries
```{r}
library(data.table)
library(dplyr)
library(RColorBrewer)
library(ggplot2)
library(ggpubr)
library(ggpmisc)
library(reshape2)
library(tibble)
library(tidyverse)
library(broom)
```

### Functions
```{r}
transform_expression_df <- function(df, metadata) {
  df %>%
    pivot_longer(
      cols = -gene_name,
      names_to = "dcc_filename",
      values_to = "expression"
    ) %>%
    left_join(metadata, by = "dcc_filename") %>%
    select(gene_name, expression, dcc_filename, Roi, Segment, Patient)
}

analyze_LR_pairs <- function(LR_df, merged_df, ligand_expr_col, sender_label, receptor_label, receptor_column) {
  
  print(LR_df)
  
  results <- list()
  
  for (i in 1:nrow(LR_df)) {
    
    print(LR_df$ligand[i])
    
    ligand   <- LR_df$ligand[i]
    
    # Define receptor in tumor as HLA gene, Ucell score or gene of matching receptor for ligand
    if (receptor_label == 'HLA-DRB1') {
      receptor <- 'HLA-DRB1'
    } else if (receptor_label == 'FER_MHC_REVISITED') {
      receptor <- 'FER_MHC_REVISITED'
    } else {
      receptor <- LR_df$receptor[i]
    }
    
    # Check if both ligand and receptor are present
    if (!(ligand %in% merged_df$gene_name) || !(receptor %in% merged_df[[receptor_column]])) {
      message(paste("Skipping", ligand, "-", receptor, ": not found in merged_df"))
      next
    }
    
    print(paste0(ligand,receptor))
    
    # Dynamically select the expression column
    ligand_df <- merged_df %>%
      filter(gene_name == ligand) %>%
      select(Roi, Patient, ligand_expr = all_of(ligand_expr_col))

    receptor_df <- merged_df %>%
      filter(.data[[receptor_column]] == receptor) %>%
      select(Roi, receptor_expr = tumor_expression)

    # Merge on Roi
    expr_df <- inner_join(ligand_df, receptor_df, by = "Roi") %>%
      na.omit()

    # Correlation
    cor_res <- cor.test(expr_df$ligand_expr, expr_df$receptor_expr, method = "pearson")
    
    cor_est <- cor_res$estimate
    cor_pval <- cor_res$p.value
    
    # Linear model
    model <- lm(receptor_expr ~ ligand_expr + Patient, data = expr_df)
    print(summary(model))
    r2 <- summary(model)$r.squared
    pval <- glance(model)$p.value
    # pval <- summary(model)$coefficients["ligand_expr", "Pr(>|t|)"]

    # Store result
    results[[length(results) + 1]] <- data.frame(
      sender = sender_label,
      ligand = ligand,
      receptor = receptor,
      correlation = cor_est,
      cor_p_value = cor_pval
    )

    # Plot
    p <- ggplot(expr_df, aes(x = ligand_expr, y = receptor_expr)) +
      geom_point(color = "steelblue", size = 3) +
      geom_smooth(method = "lm", se = TRUE, color = "red") +
      labs(
        title = paste(ligand, "(", sender_label, ") vs", receptor, "(Tumor)"),
        subtitle = paste("R² =", round(r2, 3), "| p =", signif(pval, 3)),
        x = paste("Ligand:", ligand),
        y = paste("Receptor:", receptor)
      ) +
      theme_minimal()

    # print(p)
  }
  
  print(results)
  
  # Combine results
  do.call(rbind, results)
}

# Function to check gene presence in a dataframe
check_genes <- function(genes, df, df_name) {
  if (is.null(rownames(df))) {
    stop(paste("Dataframe", df_name, "has no row names!"))
  }
  
  present_genes <- genes[genes %in% df$gene_name]
  missing_genes <- genes[!genes %in% df$gene_name]
  
  cat("====", df_name, "====\n")
  cat("Present genes:\n")
  print(present_genes)
  cat("\nMissing genes:\n")
  print(missing_genes)
  cat("\n\n")
}
```

### Read datasets
```{r}
# Read metadata and all & de-convoluted expression tables
metadata <- data.frame(fread('/../../Volumes/h345/afarkkilab/Projects/NKI/Single_Cell_LR_Analysis/Geomx_validation/Data/geomx_metadata.csv'))
gsea_all <- data.frame(fread('/../../Volumes/h345/afarkkilab/Projects/NKI/Single_Cell_LR_Analysis/Geomx_validation/Data/geomx_harmony_corr_expr.csv'))
deconv_ct_frac <- readRDS("/../../Volumes/h345/afarkkilab/Projects/NKI/Single_Cell_LR_Analysis/Geomx_validation/Data/bp_res_mid_lvl_ct_expr_mtx_cleaned_vst_harmony_batch_corr.RDS")

# Read L-R pairs 
top7LR_interactions <- data.frame(fread("/../../Volumes/h345/afarkkilab/Projects/NKI/Single_Cell_LR_Analysis/Figures_Results/top7LRinteractions.csv"))
```

### Prepare dataframes
#### Filter for selected genes from top7 L-R pairs
```{r}
# Replace '.' with '-' in ligand and receptor columns
top7LR_interactions$ligand   <- gsub("\\.", "-", top7LR_interactions$ligand)
top7LR_interactions$receptor <- gsub("\\.", "-", top7LR_interactions$receptor)

# Define subsets of L-R interactions of interest
Tcells_LR_df <- filter(top7LR_interactions, !sender %in% c('DC'))
DCs_LR_df <- filter(top7LR_interactions, sender == "DC")
# Bcells_LR_df <- filter(top7LR_interactions, sender == "Bcell")

# Select genes of interest 
tumor_receptors <- unique(c(top7LR_interactions$receptor, 'HLA-DRB1'))
stroma_ligands <- unique(top7LR_interactions$ligand)
Tcells_ligands <- unique(Tcells_LR_df$ligand)
DCs_ligands <- unique(DCs_LR_df$ligand)
# Bcells_ligands <- unique(Bcells_LR_df$ligand)
genes <- unique(c(top7LR_interactions$ligand, top7LR_interactions$receptor, 'HLA-DRB1'))

# Filter expression dataframes for selected genes only
filtered_gsea_all <- gsea_all %>% filter(gene_name %in% genes)

# Select deconvoluted matrixes for cell types of interest 
tumor_df  <- as.data.frame(deconv_ct_frac[["tumor"]])
tumor_df  <- tumor_df %>% rownames_to_column(var = "gene_name")

Tcells_df <- as.data.frame(deconv_ct_frac[["Tcells"]])
Tcells_df <- Tcells_df %>% rownames_to_column(var = "gene_name")

Bcells_df <- as.data.frame(deconv_ct_frac[["Bcells"]])
Bcells_df <- Bcells_df %>% rownames_to_column(var = "gene_name")

DCs_df    <- as.data.frame(deconv_ct_frac[["DCs"]])
DCs_df    <- DCs_df %>% rownames_to_column(var = "gene_name")

# Run the check if expression dataframes are missing any genes
check_genes(tumor_receptors, tumor_df, "tumor_df")
check_genes(Tcells_ligands, Tcells_df, "Tcells_df")
# check_genes(genes, Bcells_df, "Bcells_df")
check_genes(DCs_ligands, DCs_df, "DCs_df")

# Filter expression dataframes for selected genes only (ensure same number of genes across dfs for joining)
filtered_tumor_df  <- tumor_df %>% filter(gene_name %in% genes)
filtered_Tcells_df <- Tcells_df %>% filter(gene_name %in% genes)
# filtered_Bcells_df <- Bcells_df %>% filter(gene_name %in% genes)
filtered_DCs_df    <- DCs_df %>% filter(gene_name %in% genes)
```

```{r}
# Relabel Roi to make it unique
metadata$Roi <- paste0(metadata$Sample, '_', metadata$Roi)

# Match dcc_filanames format to metadata 
colnames(filtered_gsea_all) <- sub("\\.dcc$", "-dcc", colnames(filtered_gsea_all))           # temporarily protect '.dcc'
colnames(filtered_gsea_all) <- gsub("\\.", "-", colnames(filtered_gsea_all))                 # replace all remaining '.' with '-'
colnames(filtered_gsea_all) <- sub("-dcc$", ".dcc", colnames(filtered_gsea_all))  

all_exp_combined <- transform_expression_df(filtered_gsea_all, metadata)
tumor_combined  <- transform_expression_df(filtered_tumor_df, metadata)
Tcells_combined  <- transform_expression_df(filtered_Tcells_df, metadata)
DCs_combined     <- transform_expression_df(filtered_DCs_df, metadata)
# Bcells_combined  <- transform_expression_df(filtered_Bcells_df, metadata)

# Subset for tumor and stroma segments
stroma_combined  <- all_exp_combined[all_exp_combined$Segment == "stroma", ]
# tumor_combined   <- all_exp_combined[all_exp_combined$Segment == "tumor", ]

Tcells_combined  <- Tcells_combined[Tcells_combined$Segment == "stroma", ]
DCs_combined     <- DCs_combined[DCs_combined$Segment == "stroma", ]
```

#### Combine datasets by Roi with expression column per aoi and cell type of interest
```{r}
# Rename expression columns
tumor_combined$tumor_expression   <- tumor_combined$expression
stroma_combined$stroma_expression <- stroma_combined$expression
Tcells_combined$Tcell_expression  <- Tcells_combined$expression
DCs_combined$DCs_expression       <- DCs_combined$expression
# Bcells_combined$Bcell_expression <- Bcells_combined$expression

# Merge on gene_name and Roi to ensure same number of Rois
merged_df <- tumor_combined[, c("gene_name", "Roi", "tumor_expression", "dcc_filename", "Patient")] %>%
  left_join(stroma_combined[, c("gene_name", "Roi", "stroma_expression")], by = c("gene_name", "Roi")) %>%
  left_join(Tcells_combined[, c("gene_name", "Roi", "Tcell_expression")], by = c("gene_name", "Roi")) %>%
  left_join(DCs_combined[, c("gene_name", "Roi", "DCs_expression")], by = c("gene_name", "Roi"))
```

### Perform correlation and linear dependence analysis between L-R pairs in subsets of interest
```{r}
# For Tcells subset
Tcell_results <- analyze_LR_pairs(
  LR_df = Tcells_LR_df,
  merged_df = merged_df,
  ligand_expr_col = "Tcell_expression",
  sender_label = "Tcell",
  receptor_label = 'receptor',
  receptor_column = 'gene_name'
)

# For DCs subset
DCs_results <- analyze_LR_pairs(
  LR_df = DCs_LR_df,
  merged_df = merged_df,
  ligand_expr_col = "DCs_expression",
  sender_label = "DC",
  receptor_label = 'receptor',
  receptor_column = 'gene_name'
)

# For all (Stroma) subset
Stroma_results <- analyze_LR_pairs(
  LR_df = top7LR_interactions,
  merged_df = merged_df,
  ligand_expr_col = "stroma_expression",
  sender_label = "Stroma",
  receptor_label = 'receptor',
  receptor_column = 'gene_name'
)

all_LR_results <- bind_rows(Tcell_results, DCs_results, Stroma_results)

# Adjust p-value for multiple testing
all_LR_results$fdr <- p.adjust(all_LR_results$cor_p_value)
```
### Perform correlation and linear dependence analysis between Ligand and Tumor HLA-DRB1 pairs in subsets of interest
```{r}
# For Tcells subset
Tcell_results <- analyze_LR_pairs(
  LR_df = Tcells_LR_df,
  merged_df = merged_df,
  ligand_expr_col = "Tcell_expression",
  sender_label = "Tcell",
  receptor_label = 'HLA-DRB1',
  receptor_column = 'gene_name'
)

# For DCs subset
DCs_results <- analyze_LR_pairs(
  LR_df = DCs_LR_df,
  merged_df = merged_df,
  ligand_expr_col = "DCs_expression",
  sender_label = "DC",
  receptor_label = 'HLA-DRB1',
  receptor_column = 'gene_name'
)

# For all (Stroma) subset
Stroma_results <- analyze_LR_pairs(
  LR_df = top7LR_interactions,
  merged_df = merged_df,
  ligand_expr_col = "stroma_expression",
  sender_label = "Stroma",
  receptor_label = 'HLA-DRB1',
  receptor_column = 'gene_name'
)

all_LigHLA_results <- bind_rows(Tcell_results, DCs_results, Stroma_results)

# Adjust p-value for multiple testing
all_LigHLA_results$fdr <- p.adjust(all_LigHLA_results$cor_p_value)
```

### Read data with Ucell scores per pathways and prepare for analysis
```{r}
#GSEA calculated from deconvoluted tumor signal
gsea_deconv_tum <- data.frame(fread('/../../Volumes/h345/afarkkilab/Projects/NKI/GeoMX/Inputs/ssgsea_norm_harmony_batch_corr_deconv_tumor_custom_IFNg_pathways.csv.csv'))

# Select MHC-II Ucell score
tumor_mhcii <- gsea_deconv_tum %>% filter(pathway == "FER_MHC_REVISITED")

# Relabel Roi to make it unique
tumor_mhcii$Roi <- paste0(tumor_mhcii$Sample, '_', tumor_mhcii$Roi)

tumor_mhcii <- tumor_mhcii %>% rename_at('ssgsea_score', ~'tumor_expression')
tumor_mhcii   <- tumor_mhcii[tumor_mhcii$Segment == "tumor", ]
```

### Perform correlation and linear dependence analysis between Ligand and tumor MHC-II U cell score pairs in subsets of interest
```{r}
# Merge on gene_name and Roi to ensure same number of Rois
merged_df <- tumor_mhcii[, c("pathway", "tumor_expression", "Roi", "dcc_filename", "Patient")] %>%
  left_join(Tcells_combined[, c("gene_name", "Roi", "Tcell_expression")], by = "Roi")

# For Tcells subset
Tcell_results <- analyze_LR_pairs(
  LR_df = Tcells_LR_df,
  merged_df = merged_df,
  ligand_expr_col = "Tcell_expression",
  sender_label = "Tcell",
  receptor_label = 'FER_MHC_REVISITED',
  receptor_column = 'pathway'
)

# Merge on gene_name and Roi to ensure same number of Rois
merged_df <- tumor_mhcii[, c("pathway", "tumor_expression", "Roi", "dcc_filename", "Patient")] %>%
  left_join(DCs_combined[, c("gene_name", "Roi", "DCs_expression")], by = "Roi")

# For DCs subset
DCs_results <- analyze_LR_pairs(
  LR_df = DCs_LR_df,
  merged_df = merged_df,
  ligand_expr_col = "DCs_expression",
  sender_label = "DC",
  receptor_label = 'FER_MHC_REVISITED',
  receptor_column = 'pathway'
)

# Merge on gene_name and Roi to ensure same number of Rois
merged_df <- tumor_mhcii[, c("pathway", "tumor_expression", "Roi", "dcc_filename", "Patient")] %>%
  left_join(stroma_combined[, c("gene_name", "Roi", "stroma_expression")], by = "Roi")

# For all (Stroma) subset
Stroma_results <- analyze_LR_pairs(
  LR_df = top7LR_interactions,
  merged_df = merged_df,
  ligand_expr_col = "stroma_expression",
  sender_label = "Stroma",
  receptor_label = 'FER_MHC_REVISITED',
  receptor_column = 'pathway'
)

# all_LigMHCUcell_results <- bind_rows(Tcell_results, DCs_results, Stroma_results)

# Adjust p-value for multiple testing
Stroma_results$fdr <- p.adjust(Stroma_results$cor_p_value)
```
# Combined all results for plotting
```{r}
# Rename columns in each dataframe
all_LR_results <- all_LR_results %>%
  rename(
    r_receptor = receptor,
    correlation_receptor = correlation,
    cor_p_val_receptor = cor_p_value,
    fdr_receptor = fdr
  )

all_LigHLA_results <- all_LigHLA_results %>%
  rename(
    r_HLA_DPB1 = receptor,
    correlation_HLA_DPB1 = correlation,
    cor_p_val_HLA_DPB1 = cor_p_value,
    fdr_HLA_DPB1 = fdr
  )

all_LigMHCUcell_results <- all_LigMHCUcell_results %>%
  rename(
    r_MHCIIUcellScore = receptor,
    correlation_MHCIIUcellScore = correlation,
    cor_p_val_MHCIIUcellScore = cor_p_value,
    fdr_MHCIIUcellScore = fdr
  )

# Merge dataframes on sender and ligand using full join
# final_results_df <- all_LigHLA_results %>%
#   # full_join(all_LigHLA_results, by = c("sender", "ligand")) %>%
#   full_join(all_LigMHCUcell_results, by = c("sender", "ligand")) %>%
#   distinct()
```

# Plot correlation for MHC-II U cell Score - Fig.5 i
```{r}

# Remove duplicate rows across all columns
distinct_Stroma_results <- Stroma_results %>%
  distinct(ligand, receptor, .keep_all = TRUE) %>%
  arrange(desc(correlation))

# Convert to long format without converting to a factor
cor_est_long <- distinct_Stroma_results %>%
  pivot_longer(
    cols = starts_with("correlation"),
    names_to = "cor_type",
    values_to = "cor_value"
  ) %>%
  # Ensure cor_value is numeric and ordered based on correlation in descending order
  mutate(cor_value = as.numeric(cor_value)) %>%
  arrange(desc(cor_value))  # Sort by cor_value in descending order

# Similarly process the p-value data
pval_long <- distinct_Stroma_results %>%
  pivot_longer(
    cols = starts_with("fdr"),
    names_to = "cor_type",
    values_to = "cor_fdr"
  ) %>%
  mutate(cor_type = str_replace(cor_type, "fdr", ""))

# Merge correlation and p-value data
combined_long <- cor_est_long %>%
  mutate(cor_type = str_replace(cor_type, "correlation", "")) %>%
  left_join(pval_long, by = c("sender", "ligand", "cor_type"))

# Ensure the data is ordered by correlation values
combined_long <- combined_long %>%
  mutate(ligand = factor(ligand, levels = combined_long$ligand[order(combined_long$cor_value, decreasing = FALSE)]))

# Plot heatmap with size based on p-value
p <- ggplot(combined_long, aes(x = cor_type, y = ligand, fill = cor_value, size = -log10(cor_fdr))) +
  geom_point(shape = 21, color = "black") +
  scale_fill_gradient2(low = "#6599CD", mid = "#D9DEE4", high = "#D33F49", midpoint = 0,
                       name = "Pearson\nCorrelation") +
  scale_size_continuous(name = "-log10(FDR)") +
  theme_minimal() +
  labs(x = "MHC-II tumor ssGSEA score", y = "") +
  ggtitle("Correlation of ligands vs tumor MHC-II Ucell score")

print(p)

# Save plot
ggsave(
  filename = "corr_Ligand_vs_tumor_MHCUcellScore.svg",
  plot = p,
  width = 3, height = 3.5
)
```

