---
title: "R Notebook"
output: html_notebook
---

#Loading libraries
```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(RColorBrewer)
```


#Reading input files and metrics
```{r}
project.path = "D:/users/fperez/NKI_TMAs_AF/"
output.folder = "D:/users/fperez/NKI_TMAs_AF/Analysis_results/"
tribus.subfolder = "Tribus_Cell-segment_202311/"
celltypes.file.suffix = "_cellTypes.csv"
CellTypes2ignore = c("InsideROI", "LostCell", "LowDapi", "Out.of.core")
clinical.info <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Clinical_data/T-CycIF.csv", sep=",", header = TRUE)
cores2ignore1 <- read.table(file="D:/users/fperez/NKI_TMAs_AF/devNKI-scripts/utils/Total_cores_to_ignore.csv", sep=",")
cores2ignore2 <- read.table(file="D:/users/fperez/NKI_TMAs_AF/devNKI-scripts/utils/Total_extra-cores_to_ignore.csv",
                            sep = ",", header = TRUE)
cores2ignore3 <- read.table(file="D:/users/fperez/NKI_TMAs_AF/devNKI-scripts/utils/Total_cores_to_ignore_extra3.csv",
                            sep = ",", header = TRUE)

cores2ignore3 <- cores2ignore3[,c(1,2)]
cores2ignore2 <- cores2ignore2[,-3]


#File with patient IDs and coreIDs relationship
samples2cycif <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Clinical_intersection/All-slides_cycif2samples.txt",
                            sep = "\t", header = TRUE)

source('D:/users/fperez/NKI_TMAs_AF/devNKI-scripts/3_Cell-typeClassification/Celltype-markers_logic_NKI_2023.R')
```



```{r}
#Merging files with cores to ignore and excluding those from the samples2cycif table
colnames(cores2ignore1) <- colnames(cores2ignore2)
cores2ignore <- rbind(cores2ignore2, cores2ignore1, cores2ignore3)
cores2ignore[,2] <- paste0("core",cores2ignore[,2])

cols <- sapply(1:nrow(cores2ignore), function(x){
  col <- which(samples2cycif$cycif.slide == cores2ignore[x,1] & samples2cycif$cycif.core.id == cores2ignore[x,2])
  return(col)
})

samples2cycif.sel <- samples2cycif[-sort(unlist(cols)),] #Ignoring cores that do not passed QC
samples2cycif.sel <- samples2cycif.sel[which(samples2cycif.sel$patient != "OC liver"),] #Ignoring control samples
samples2cycif.dat <- merge(samples2cycif.sel, clinical.info, by.x="patient", by.y="TnumberAVL")
```

#Reading input tCycif data and cell.type data
```{r}
#List all the tables with celltype labels
big.data.set <- NULL
files <- list.files(path=project.path, pattern = "TMA_", full.names = T)
#Read all cell.type and signal intensity data
for (f in files){
  cycif.slide <- basename(f)
  slide.data <- read.table(file=paste0(f, "/", tribus.subfolder, "/", cycif.slide, celltypes.file.suffix), sep=",", header=TRUE)
  slide.data <- cbind(cycif.slide, slide.data)
  slide.data$Row_number <- paste0(cycif.slide,":",slide.data$Row_number)
  colnames(slide.data)[colnames(slide.data) == "CoreId"] <- "cycif.core.id"
  slide.data$cycif.core.id <- paste0("core",slide.data$cycif.core.id)
  big.data.set <- rbind(big.data.set,slide.data)
}
```


#QC and filtering of the cell.type dataset
```{r}
#Performing QC according to cell.types
#Ignoring cell with low DAPI intensity and out of the core
big.data.set <- big.data.set %>% dplyr::filter(GlobalCellType != "LowDapi")

#Ignoring non-useful columns
big.data.set <- big.data.set %>% select(-out_of_core, -lost, -Inside_ROI)
```

#Merging clinical info with tCycif and cell.type info
```{r}
interesting.clin.columns <- c("patient","cycif.slide","cycif.core.id","therapy_sequence",
                              "centre","refage","incyear")
samples2cycif.dat.sel <- samples2cycif.dat[,interesting.clin.columns]

merged.dat <- merge(big.data.set, samples2cycif.dat.sel, by=c("cycif.slide","cycif.core.id"))

merged.dat$GlobalCellType[grep("Stromal", merged.dat$GlobalCellType, ignore.case = TRUE)] <- "Stromal"
merged.dat[merged.dat$GlobalCellType == "Other", "GlobalCellType"] <- "Others"
```


##Number of cores by patient
```{r, fig.width=3, fig.height=4}
cores.by.patient <- merged.dat %>% 
                          group_by(patient) %>% 
                          summarise(n = length(unique(cycif.core.id)))

barplot(table(cores.by.patient$n), xlab="Number of cores", ylab="Number of patients")

```

#Barplot with percentage of immune, stromal and cancer cells per patient
```{r}
out.put.subfolder <- "01_Cell-proportions_Heatmaps_Nov2023/"
dir.create(paste0(output.folder, out.put.subfolder))

print(round(table(merged.dat$GlobalCellType) * 100 /nrow(merged.dat),2))

Lymphoids <- c("CD8.effector.T.cells", "CD8.CD45RO.T.cells", "CD8.PD1.T.cells",  "CD4.T.cells",
               "CD4.CD45RO.T.cells", "CD4.PD1.T.cells", "T.regs", "B.cells", "Other.immune")

Myeloids <- c("CD11c.MY","Other.MY","CD68.MP", "CD207.MY", "CD163.MP", "CD15.MY")

       
aux.dat <- merged.dat

aux.dat[aux.dat$GlobalCellType %in%  Lymphoids,"GlobalCellType"] <- "Lymphoid"
aux.dat[aux.dat$GlobalCellType %in%  Myeloids,"GlobalCellType"] <- "Myeloid"

merged.types.patients <- aux.dat %>% 
                          group_by(patient,GlobalCellType) %>% 
                          summarise(n=n()) %>%
                          mutate(Proportion=n * 100 / sum(n))


merged.types.patients$GlobalCellType <- factor(merged.types.patients$GlobalCellType,
                                               levels=c("Cancer","Lymphoid", "Myeloid", "Stromal","Others"))

df <- merged.types.patients %>% 
	      dplyr::filter(GlobalCellType=="Cancer") %>% 
	      arrange(desc(Proportion))
merged.types.patients$patient <- factor(merged.types.patients$patient, levels=df$patient)

colnames(merged.types.patients)[c(1,2)] <- c("Patient","Cell.type")

p <- ggplot(merged.types.patients, aes(x=Patient, y=Proportion, fill=Cell.type)) +
        geom_bar(stat="identity") + theme_classic() +
        scale_fill_manual(values = c("#e600e6", "#FF7F00", "#2F82B9", "#e6c200", "grey")) +
        theme(axis.text.x=element_text(size=rel(0.2), angle = 65, hjust = 1), legend.position="top",
              legend.text = element_text(size=7), legend.key.size = unit(0.4, 'cm'),  legend.title = element_text(size=5))
print(p)
ggsave(p, filename = paste0(output.folder, out.put.subfolder, "Proportions-barplot_Global.svg"),
        width = 15, height = 8.5, units = "cm")
ggsave(p, filename = paste0(output.folder, out.put.subfolder, "Proportions-barplot_Global.png"),
        width = 15, height = 8.5, units = "cm")

########################## ########################## ########################## 
########################## Now for Immune cells #################

aux.dat <- merged.dat
aux.dat <- aux.dat[aux.dat$GlobalCellType != "Cancer" & aux.dat$GlobalCellType != "Stromal" & aux.dat$GlobalCellType != "Others",]

merged.types.patients <- aux.dat %>% 
                          group_by(patient,GlobalCellType) %>% 
                          summarise(n=n()) %>%
                          mutate(Proportion=n * 100 / sum(n))

#Order patients by proportion of CD11c.MY
df <- merged.types.patients %>% 
	dplyr::filter(GlobalCellType=="CD11c.MY") %>% 
	arrange(desc(Proportion))

merged.types.patients$patient <- factor(merged.types.patients$patient, levels=df$patient)

colnames(merged.types.patients)[c(1,2)] <- c("Patient","Cell.type")

myloid.colors  <- c("#ade6e6", "#add8e6", "#0095b6", "#2F82B9", "#8DC594", "#62B74F")
lymphocite.colors <- c("#e4717a", "#E94330", "#ff0000", "#ff6e4a", "#FF7F00", "#FDB661", "yellow")
b.cells.color <-  "purple"
other.color <- "grey"

mycolors <- c(myloid.colors, lymphocite.colors, b.cells.color, other.color)

my.cell.levels <- c("CD11c.MY","CD15.MY","CD207.MY", "Other.MY", "CD163.MP", "CD68.MP",
                    "CD4.T.cells", "CD4.CD45RO.T.cells", "CD4.PD1.T.cells", "CD8.CD45RO.T.cells",
                    "CD8.effector.T.cells", "CD8.PD1.T.cells", "T.regs", "B.cells",
                    "Other.immune")

merged.types.patients$Cell.type <- factor(merged.types.patients$Cell.type,
                                    levels = my.cell.levels)

p <- ggplot(merged.types.patients, aes(x=Patient, y=Proportion, fill=Cell.type)) +
        geom_bar(stat="identity") + theme_classic() +
        scale_fill_manual(values = mycolors) +
        theme(axis.text.x=element_text(size=rel(0.2), angle = 65, hjust = 1), legend.position="top",
              legend.text = element_text(size=7), legend.key.size = unit(0.4, 'cm'), legend.title = element_text(size=5))
print(p)
ggsave(p, filename = paste0(output.folder, out.put.subfolder, "Proportions-barplot_Immune.svg"),
        width = 15, height = 9, units = "cm")
ggsave(p, filename = paste0(output.folder, out.put.subfolder, "Proportions-barplot_Immune.png"),
        width = 15, height = 9, units = "cm")

########################## ########################## ########################## 
########################## Now for Immune merged cells #################
aux.dat <- merged.dat
aux.dat <- aux.dat[aux.dat$GlobalCellType != "Cancer" & aux.dat$GlobalCellType != "Stromal" & aux.dat$GlobalCellType != "Others",]

aux.dat$GlobalCellType[grep("CD8", aux.dat$GlobalCellType)] <- "CD8.T.cells"
aux.dat$GlobalCellType[grep("CD4", aux.dat$GlobalCellType)] <- "CD4.T.cells"

merged.types.patients <- aux.dat %>% 
                          group_by(patient,GlobalCellType) %>% 
                          summarise(n=n()) %>%
                          mutate(Proportion=n * 100 / sum(n))

df <- merged.types.patients %>% 
	dplyr::filter(GlobalCellType=="CD8.T.cells") %>% 
	arrange(desc(Proportion))


merged.types.patients$patient <- factor(merged.types.patients$patient, levels=df$patient)

colnames(merged.types.patients)[c(1,2)] <- c("Patient","Cell.type")

nb.cols <- length(unique(merged.types.patients$Cell.type)) - 2
mycolors <- colorRampPalette(brewer.pal(8, "Paired"))(nb.cols)
mycolors <- c(mycolors,"purple", "grey")

my.cell.levels <- c("CD8.T.cells", "CD4.T.cells", "T.regs", "CD11c.MY",
                    "CD163.MP","CD207.MY", "CD68.MP", "CD15.MY", "Other.MY", "B.cells", "Other.immune")

merged.types.patients$Cell.type <- factor(merged.types.patients$Cell.type,
                                    levels = my.cell.levels)

p <- ggplot(merged.types.patients, aes(x=Patient, y=Proportion, fill=Cell.type)) +
        geom_bar(stat="identity") + theme_classic() +
        scale_fill_manual(values = mycolors) +
        theme(axis.text.x=element_text(size=rel(0.2), angle = 65, hjust = 1), legend.position="top",
              legend.text = element_text(size=7), legend.key.size = unit(0.4, 'cm'), legend.title = element_text(size=5))
print(p)
ggsave(p, filename = paste0(output.folder, out.put.subfolder, "Proportions-barplot_Immune_merged.svg"),
        width = 15, height = 9, units = "cm")
ggsave(p, filename = paste0(output.folder, out.put.subfolder, "Proportions-barplot_Immune_merged.png"),
        width = 15, height = 9, units = "cm")

rm(aux.dat)
```
#Function for plotting UMAPS
```{r}
#Function for scaling dataset by slide
#It will also transfor data to log2 and will truncate the outliers above the 0.99 and below the 0.01 quantiles the corresponing quantiles
scaling.sampling <- function(datain, channels, nsampling, slide.name="cycif.slide"){
                  slide.name2 <- ensym(slide.name)
                  dat.log2 <- datain[,c(channels,slide.name)]
                  dat.log2[,channels] <- log2(dat.log2[,channels])
                  #Scaling by slide
                  scaled.data <- sapply(unique(dat.log2[,slide.name]), function(x){
                    dat.slide <- dat.log2 %>% filter({{slide.name2}} == x) %>% select(-{{slide.name2}}) %>% as.matrix()
                    quantiles <- apply(dat.slide, 2, function(x){quantile(x,c(0.01, 0.99))})
                    for (x in colnames(dat.slide)){
                      dat.slide[,x][which(dat.slide[,x] <= quantiles[1,x])] = quantiles[1,x]
                      dat.slide[,x][which(dat.slide[,x] >= quantiles[2,x])] = quantiles[2,x]
                    } 
                    as.matrix(scale(dat.slide))
                  })
                  scaled.data2 <- NULL #Concatenating
                  for (i in 1:length(scaled.data)){ scaled.data2 <- rbind(scaled.data2, scaled.data[[i]])}
                  rm(scaled.data)
                  #Sampling  
                  rows.to.sample <- sample(1:nrow(scaled.data2), nsampling)
                  dat.scaled.glob <- scaled.data2[rows.to.sample,]
                  Global.types <- datain %>% slice(rows.to.sample) %>% pull(GlobalCellType)
                  return(list(dat.scaled.glob,Global.types))
              }
```


#Calculating UMAPs
```{r}
set.seed(42) #For reproducibility 
#Channels for global.cell types umaps
global.channels.UMAPs <- c("CD3d","CD8a","CD4","FOXP3","CD68", "MHCII", "aSMA", "Vimentin",
                                "CD163","CD207","CD11c", "CK7","ECadherin","PD1","CD45RO")
immune.channels.UMAPs <- c("CD3d","CD8a","CD4","FOXP3","CD68", "MHCII", "CD15",
                           "CD11b", "CD20", "CD163","CD207","CD11c", "PD1","CD45RO")
N.sampling = 50000 #Number of cells to sample for UMAP

global.colors <- c("#e600e6", "#FF7F00", "#2F82B9", "#e6c200", "grey") #Global.celltypes

myloid.colors  <- c("#ade6e6", "#add8e6", "#0095b6", "#2F82B9", "#8DC594", "#62B74F")
lymphocite.colors <- c("#e4717a", "#E94330", "#ff0000", "#ff6e4a", "#FF7F00", "#FDB661", "yellow")
b.cells.color <-  "purple"
other.color <- "grey"

immune.colors <- c(myloid.colors, lymphocite.colors, b.cells.color, other.color)

immune.levels <- c("CD11c.MY","CD15.MY","CD207.MY", "Other.MY", "CD163.MP", "CD68.MP",
                    "CD4.T.cells", "CD4.CD45RO.T.cells", "CD4.PD1.T.cells", "CD8.CD45RO.T.cells",
                    "CD8.effector.T.cells", "CD8.PD1.T.cells", "T.regs", "B.cells",
                    "Other.immune")

#Generating input data for global cell.types UMAPs
sampled.global <- scaling.sampling(merged.dat, global.channels.UMAPs, N.sampling, slide.name="cycif.slide")
dat.sampled.global <- sampled.global[[1]]
types.sampled.global <- sampled.global[[2]]
rm(sampled.global)
types.sampled.global[types.sampled.global %in% Lymphoids] <- "Lymphoid"
types.sampled.global[types.sampled.global %in% Myeloids] <- "Myeloid"
types.sampled.global <- factor(types.sampled.global, levels=c("Cancer","Lymphoid","Myeloid","Stromal","Others"))


#Generating input data immune cell.types UMAPs
merged.dat.immune <- merged.dat[merged.dat$GlobalCellType %in% c(Lymphoids, Myeloids),]
sampled.immune <- scaling.sampling(merged.dat.immune, immune.channels.UMAPs, N.sampling, slide.name="cycif.slide")
dat.sampled.immune <- sampled.immune[[1]]
types.sampled.immune <- sampled.immune[[2]]
rm(sampled.immune)
types.sampled.immune <- factor(types.sampled.immune, levels = immune.levels)


#UMAP for global celltypes
umap.global = uwot::umap(dat.sampled.global, n_neighbors = 50, scale=FALSE, spread=1.5, min_dist = 0.3, n_epochs = 60)
global.uplot = data.frame(x = umap.global[,1], y= umap.global[,2])

#UMAP for immmune cells
umap.immune = uwot::umap(dat.sampled.immune, n_neighbors = 50, scale=FALSE, spread=1.5, min_dist = 0.3, n_epochs = 60)
immune.uplot = data.frame(x = umap.immune[,1], y= umap.immune[,2])
```


#Plotting UMAPs
```{r}
#Plotting UMAP for immune cells
l = ggplot(immune.uplot,aes(x,y, color=types.sampled.immune))+ geom_point(size=0.4, stroke=0, alpha=0.7)+ 
    geom_density_2d(color="grey60", alpha=0.6, size = 0.25) +
    scale_fill_viridis(discrete = TRUE) + 
    guides(color = guide_legend(override.aes = list(size=5))) + 
    scale_color_manual(values=immune.colors) + theme_bw() + coord_fixed(ratio=1) +
    guides(colour = guide_legend(override.aes = list(size=4, shape=15, alpha=1), label.position="bottom",
                                 direction="horizontal", ncol=2,  byrow=F)) + 
    xlab("umap1") + ylab("umap2") + xlim(-9.25, 8) + ylim(-11.85,11.25) + 
    theme(legend.position="right", legend.text = element_text(size=6), legend.spacing.x = unit(0.05, 'cm'),
          legend.spacing.y = unit(0.06, 'cm'), legend.key.size = unit(0.4, 'cm'), legend.title = element_blank(),
          aspect.ratio=1)
print(l)
ggsave(l, filename = paste0(output.folder, out.put.subfolder, "UMAP_Immunes.svg"),
        width = 16, height = 8, units = "cm")
ggsave(l, filename = paste0(output.folder, out.put.subfolder, "UMAP_Immunes.png"),
        width = 16, height = 8, units = "cm")


#Plotting UMAP for global cell types
l = ggplot(global.uplot,aes(x,y, color=types.sampled.global))+ geom_point(size=0.4, stroke=0, alpha=0.7)+ 
    geom_density_2d(color="grey60", alpha=0.6, size = 0.25) +
    scale_fill_viridis(discrete = TRUE) + 
    guides(color = guide_legend(override.aes = list(size=5))) + 
    scale_color_manual(values=global.colors) + theme_bw() + coord_fixed(ratio=1) +
    guides(colour = guide_legend(override.aes = list(size=4, shape=15, alpha=1), label.position="bottom",
                                 direction="horizontal", ncol=1,  byrow=F)) + 
    xlab("umap1") + ylab("umap2") + xlim(-11, 9) + ylim(-8,11.1) + 
    theme(legend.position="right", legend.text = element_text(size=7), legend.spacing.x = unit(0.05, 'cm'),
          legend.spacing.y = unit(0.06, 'cm'), legend.key.size = unit(0.5, 'cm'), legend.title = element_blank(),
          aspect.ratio=1)
print(l)
ggsave(l, filename = paste0(output.folder, out.put.subfolder, "UMAP_Global-celltypes.svg"),
         width = 16, height = 8, units = "cm")
ggsave(l, filename = paste0(output.folder, out.put.subfolder, "UMAP_Global-celltypes.png"),
         width = 16, height = 8, units = "cm")
```



#Calculating proportion of cells by patient and plotting boxplots
```{r}
out.put.subfolder <- "01_Cell-proportions_Heatmaps_Nov2023/"

core.percentages <- merged.dat %>% group_by(patient, cycif.core.id, GlobalCellType) %>%
                    summarise(n = n()) %>%
                    mutate(Proportion = n / sum(n))

max.cell.percentages <- core.percentages  %>%
                        group_by(patient, GlobalCellType) %>%
                        summarise(Proportion=max(Proportion))

med.cell.percentages <- core.percentages  %>%
                        group_by(patient, GlobalCellType) %>%
                        summarise(Proportion=median(Proportion))


cell.patient.percentages <- merged.dat %>% group_by(patient, GlobalCellType) %>%
                    summarise(n = n()) %>%
                    mutate(Proportion = n  / sum(n))

#Maximum cell percentage by patient
p1 <- ggplot(max.cell.percentages, aes(x=GlobalCellType, y=Proportion)) +
  geom_point() + geom_boxplot(fill=NA) +
  theme_bw() + coord_flip() 

p2 <- p1 + ylim(0,35)
print(p1)
print(p2)

ggsave(p1, filename = paste0(output.folder, out.put.subfolder, "Proportions-Max-boxplot.png"),
        width = 15, height = 9, units = "cm")
ggsave(p1, filename = paste0(output.folder, out.put.subfolder, "Proportions-Max-boxplot.svg"),
        width = 15, height = 9, units = "cm")
ggsave(p2, filename = paste0(output.folder, out.put.subfolder, "Proportions-Max-boxplot_cut.png"),
        width = 15, height = 9, units = "cm")
ggsave(p2, filename = paste0(output.folder, out.put.subfolder, "Proportions-Max-boxplot_cut.svg"),
        width = 15, height = 9, units = "cm")


#Total cell percentage by patient
p1 <- ggplot(cell.patient.percentages, aes(x=GlobalCellType, y=Proportion)) +
  geom_point() + geom_boxplot(fill=NA) +
  theme_bw() + coord_flip() 

p2 <- p1 + ylim(0,0.3)
print(p1)
print(p2)

ggsave(p1, filename = paste0(output.folder, out.put.subfolder, "Proportions-total-boxplot.png"),
        width = 15, height = 9, units = "cm")
ggsave(p1, filename = paste0(output.folder, out.put.subfolder, "Proportions-total-boxplot.svg"),
        width = 15, height = 9, units = "cm")
ggsave(p2, filename = paste0(output.folder, out.put.subfolder, "Proportions-total-boxplot_cut.png"), width = 15, height = 9, units = "cm")
ggsave(p2, filename = paste0(output.folder, out.put.subfolder, "Proportions-total-boxplot_cut.svg"), width = 15, height = 9, units = "cm")
```

#Box plot for concordance between B.cells in t.Cycif and CD20 in IHC
```{r}
####################################################################
#Box plot for concordance between B.cells in t.Cycif and CD20 in IHC

out.put.subfolder <- "01_Cell-proportions_Heatmaps_Nov2023/"

#cell.percentages2 <- as.data.frame(cell.percentages)
cell.percentages2 <- as.data.frame(max.cell.percentages)
#cell.percentages2 <- cell.percentages[,-3]
cells_percentages.m <- cell.percentages2 %>% pivot_wider(names_from = GlobalCellType, values_from = Proportion)

cells_percentages.m <- as.data.frame(cells_percentages.m)
row.names(cells_percentages.m) <- cells_percentages.m[,1]
cells_percentages.m[is.na(cells_percentages.m)] <- 0

cells_percentages.mc <- merge(cells_percentages.m, clinical.info, by.x="patient", by.y="TnumberAVL")

cells_percentages.mc$CD20 <- factor(cells_percentages.mc$CD20, levels=c("<20","20-50",">=50"))

p <- ggplot(cells_percentages.mc, aes(x=CD20, y=B.cells)) + ylim(0,0.21) +
  geom_point(position= position_jitter(width= .2), size= 2, alpha = 0.5) +
  geom_boxplot(fill=NA) + theme_bw() + xlab("IHC CD20 category by pathologist") +
  ylab("t.CyCIF B.cell proportion")
print(p)
ggsave(p, filename = paste0(output.folder, out.put.subfolder, "Boxplot_CD20-B.cells.svg"),
        width = 7, height = 8, units = "cm")
ggsave(p, filename = paste0(output.folder, out.put.subfolder, "Boxplot_CD20-B.cells.png"),
        width = 7, height = 8, units = "cm")


test1 <- wilcox.test(cells_percentages.mc[cells_percentages.mc$CD20 == "<20","B.cells"],
            cells_percentages.mc[cells_percentages.mc$CD20 == "20-50","B.cells"])
test2 <- wilcox.test(cells_percentages.mc[cells_percentages.mc$CD20 == "<20","B.cells"],
            cells_percentages.mc[cells_percentages.mc$CD20 == ">=50","B.cells"])
test3 <- wilcox.test(cells_percentages.mc[cells_percentages.mc$CD20 == ">=50","B.cells"],
            cells_percentages.mc[cells_percentages.mc$CD20 == "20-50","B.cells"])
print(test1$p.value)
print(test2$p.value)
print(test3$p.value)
```


#Box plot for concordance between CD8.T cells in t.Cycif and CD8 in IHC
```{r}
####################################################################
#Box plot for concordance between CD8.T cells in t.Cycif and CD8 in IHC

out.put.subfolder <- "01_Cell-proportions_Heatmaps_Nov2023/"
aux.dat <- merged.dat
aux.dat$GlobalCellType[grep("CD8", aux.dat$GlobalCellType)] <- "CD8.T.cells"

core.percentages <- aux.dat %>% group_by(patient, cycif.core.id, GlobalCellType) %>%
                    summarise(n = n()) %>%
                    mutate(Proportion = n / sum(n))

max.cell.percentages <- core.percentages  %>%
                        group_by(patient, GlobalCellType) %>%
                        summarise(Proportion=max(Proportion))

cell.percentages2 <- as.data.frame(max.cell.percentages)
cells_percentages.m <- cell.percentages2 %>% pivot_wider(names_from = GlobalCellType, values_from = Proportion)

cells_percentages.m <- as.data.frame(cells_percentages.m)
row.names(cells_percentages.m) <- cells_percentages.m[,1]
cells_percentages.m[is.na(cells_percentages.m)] <- 0

cells_percentages.mc <- merge(cells_percentages.m, clinical.info, by.x="patient", by.y="TnumberAVL")

cells_percentages.mc$CD8 <- factor(cells_percentages.mc$CD8, levels=c("<20","20-100",">=100"))

p <- ggplot(cells_percentages.mc, aes(x=CD8, y=CD8.T.cells)) +
  geom_point(position= position_jitter(width= .2), size= 2, alpha = 0.5) +
  geom_boxplot(fill=NA) + theme_bw() + ylim(0,0.23) + xlab("IHC CD8 category by pathologist") +
  ylab("t.CyCIF CD8.T cells proportion")
print(p)
ggsave(p, filename = paste0(output.folder, out.put.subfolder, "Boxplot_CD8-CD8T.cells.svg"),
        width = 7, height = 8, units = "cm")
ggsave(p, filename = paste0(output.folder, out.put.subfolder, "Boxplot_CD8-CD8T.cells.png"),
        width = 7, height = 8, units = "cm")

test1 <- wilcox.test(cells_percentages.mc[cells_percentages.mc$CD8 == "<20","CD8.T.cells"],
            cells_percentages.mc[cells_percentages.mc$CD8 == "20-100","CD8.T.cells"])
test2 <- wilcox.test(cells_percentages.mc[cells_percentages.mc$CD8 == "<20","CD8.T.cells"],
            cells_percentages.mc[cells_percentages.mc$CD8 == ">=100","CD8.T.cells"])
test3 <- wilcox.test(cells_percentages.mc[cells_percentages.mc$CD8 == "20-100","CD8.T.cells"],
            cells_percentages.mc[cells_percentages.mc$CD8 == ">=100","CD8.T.cells"])
print(test1$p.value)
print(test2$p.value)
print(test3$p.value)

rm(aux.dat)
```


#Saving the proportion of cell by core
```{r}
out.put.subfolder="01_Survival_Proportions_November_medians"
core.percentages <- merged.dat %>% group_by(patient, cycif.core.id, GlobalCellType) %>%
                    summarise(n = n()) %>%
                    mutate(Proportion = n / sum(n))

tab <- as.data.frame(core.percentages)
write.table(tab, file=paste0(output.folder, out.put.subfolder, "Proportions_by_core.csv"), sep=",", row.names = FALSE)

```


###Performing heatmaps for the expression of data
```{r Heatmaps}
source("gates_NKI_2023.R")
source("cell_type_caller_functions.R")
source("qc_functions.R")

out.put.subfolder <- "01_Cell-proportions_Heatmaps/"

###########A function to make the heatmaps for several inputs #########################
heatmap.plots <- function(data.in, channels.of.interest, suffix="Global-celltypes",
                          Eccentricity.val=TRUE){
  
  #First normalization by slide using z-score
  aux.data.norm.slide1 <- norm.by.slide.fun(data.in, channels.interest = channels.of.interest,
                                            normalization ="z.score")
  #First normalization by slide using z-score + min.max
  aux.data.norm.slide2 <- norm.by.slide.fun(data.in, channels.interest = channels.of.interest,
                                            normalization ="min.max")
  
  print("Normalization done")
  #Scaling Eccentricity in whole dataset
  aux.data.norm.slide1$Eccentricity <- as.vector(scale(aux.data.norm.slide1$Eccentricity))
  aux.data.norm.slide2$Eccentricity <- as.vector(scale(aux.data.norm.slide2$Eccentricity))
  aux.data.norm.slide2$Eccentricity <- min_max_norm(aux.data.norm.slide2$Eccentricity)
  
  if (Eccentricity.val){
    channels.of.interest <- c(channels.of.interest, "Eccentricity")
  }
  
  ind.plots1 = index_plots(aux.data.norm.slide1, aux.data.norm.slide1,
                          gates.input=channels.of.interest,
                          scaling = FALSE)
  
  ind.plots2 = index_plots(aux.data.norm.slide2, aux.data.norm.slide2,
                          gates.input=channels.of.interest,
                          scaling = FALSE)
  print(paste0("Saving files: ", output.folder,
               out.put.subfolder, "Heatmap_",suffix,"_zscore.pdf"))
  pdf(file=paste0(output.folder, out.put.subfolder, "Heatmap_",suffix,"_zscore.pdf"),
      height = 8, width = 10)
  print(ind.plots1[[3]])
  dev.off()
  print(ind.plots1[[3]])
  pdf(file=paste0(output.folder, out.put.subfolder, "Heatmap_", suffix,"_min-max.pdf"),
      height = 8, width = 10)
  print(ind.plots2[[3]])
  dev.off()
  print(ind.plots2[[3]])
}
######################Making plots for different datasets######################
channels.for.Globalplots <- c("CK7","ECadherin", unique(c(unlist(global.gates),         
                                                          unlist(immune.gates),
                                                          unlist(Lymphoid.gate),
                                                          unlist(Myeloid.gate),
                                                          unlist(cd8.gates))))
channels.for.Globalplots <- channels.for.Globalplots[channels.for.Globalplots != "Eccentricity"]

channels.for.Globalplots <- c(channels.for.Globalplots, "CD20", "CD15", "CD68")

channels.for.Immuneplots <- channels.for.Globalplots[!channels.for.Globalplots %in%
                                          c("ECadherin", "CK7", "aSMA", "Vimentin", "CD31")]

merged.dat.immune <- merged.dat %>% dplyr::filter(GlobalCellType != "Cancer",
                                                  GlobalCellType != "Stromal",
                                                  GlobalCellType != "Others")


heatmap.plots(merged.dat, channels.for.Globalplots, suffix="Global-celltypes",
              Eccentricity.val=TRUE)

heatmap.plots(merged.dat.immune, channels.for.Immuneplots, suffix="Immune-celltypes",
              Eccentricity.val=FALSE)

```