---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(survival)
library(survminer)
library(gridExtra)
library(grid)
library(forestploter)
library(reshape2)
library(ComplexHeatmap)
library(circlize)
library(EnvStats)
```

#Defining functions
```{r}
#min_max function for each column in a matrix
min.max.norm <- function(x, na.rm = TRUE) {
    return((x- min(x)) /(max(x)-min(x)))
}

median.center <- function(x, na.rm = TRUE) {
    return((x - median(x))/sd(x))
}

#Function to calculate Wilcoxon.text for each Molecular.profile2 vs the rest of the profiles. It returns a dataframe.
#Y.val is used for plotting of the significant p.values using ggplot2
p.vals.markers2 <- function(data, Profile.test="HRD", y.val=25, alt.hypothesis="greater"){
       test.dat <- data$Proportion[data$Molecular.profile2 == Profile.test]
       p.vals <- sapply(unique(data$Molecular.profile2), function(x){
                mol.dat <- data$Proportion[data$Molecular.profile2 == x]
                wtest <- wilcox.test(test.dat, mol.dat, alternative=alt.hypothesis)
                wtest$p.value
       })
      #Organizing p-values in data.frame for plotting using geom_text
      p.vals.dataframe <- data.frame(pval=p.vals,
                           Molecular.profile2=unique(data$Molecular.profile2),
                           Y.val=y.val)
      return(p.vals.dataframe)
}
```



#Reading input
```{r}
#Reading input cell proportion table by patient
core.cell.prop <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Analysis_results/01_Cell-proportions_Heatmaps_Nov2023/Proportions_by_core.csv", sep=",", header = TRUE)

#Reading clinical info
core.patient.slide <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Clinical_intersection/All-slides_cycif2samples.txt",
                                 header = TRUE, sep="\t")

#Reading clinical info
clinical.info <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Clinical_data/T-CycIF.updated.csv",
                            sep=",", header = TRUE)

Mol.profiles.patients <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Tables/Molecular_profiles_patients_20231208.csv",
            sep=",", header=TRUE)

output.folder <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/01_Cell-proportions_Heatmaps_Nov2023/"

dir.create(output.folder, showWarnings = FALSE)
```


```{r}
Mol.profiles.patients
```



#Selecting only patient information with at leats 2 cores
```{r}
#Patients selected
print(paste("Previous number of rows", nrow(core.cell.prop)))
sel.patients <- core.cell.prop %>% group_by(patient) %>%
                  summarise(n.cores=length(unique(cycif.core.id))) %>%
                  dplyr::filter (n.cores >=2 ) %>% pull(patient)

core.cell.prop <- core.cell.prop[core.cell.prop$patient %in% sel.patients,]
print(paste("Selected number of rows", nrow(core.cell.prop)))
```
#Identifying patients with in global, a low proportion of cancer cells per all samples
```{r}
cancer.prop.patient <- core.cell.prop %>%  group_by(patient) %>% 
                                mutate(Prop.cell = n/sum(n)) %>% filter(GlobalCellType == "Cancer") %>% 
                                summarise(Cancer.prop=sum(Prop.cell)) %>% 
                                dplyr::filter(Cancer.prop >= 0.05) %>% 
                                pull(patient)


core.cell.prop %>% dplyr::filter(patient %in% cancer.prop.patient) %>%
                    mutate(patient.core=paste0(patient,"_",cycif.core.id)) %>%
                    select(patient.core) %>%
                    unique()
```



#Calculating median cell proportions by patient
```{r}
core.cell.prop$GlobalCellType[core.cell.prop$GlobalCellType == "CD8.effector.T.cells"] <- "CD8.T.cells"


#The median by patient/tumor is taken from the median proportions for each core 
patient.cell.prop <- core.cell.prop %>% group_by(patient, GlobalCellType) %>% 
                      summarise(Proportion = median(Proportion)) %>% as.data.frame() %>%
                      pivot_wider(names_from = GlobalCellType, values_from = Proportion)
patient.cell.prop[is.na(patient.cell.prop)] <- 0 
patient.cell.prop
```





#Calculating correlation between cell type proportions
```{r}
library(corrplot)
library(Hmisc)

#Analysis for Supplementary Figure 2a
vars.all = c("CD8+T.memory","CD8+T.exhausted","CD8+T.effector",
             "B.cells","CD4+T.effector","CD4+T.exhausted","CD4+T.memory",
             "Macrophages", "CD11c+MY","Neutrophils","CD163+MP","Dendritic","T.regs","Other.MY")

cols.fun= circlize::colorRamp2(c(0, 0.001, 1), c("white", "yellow","#b30000"))
dat <- patient.cell.prop %>% select(any_of(vars.all)) %>% as.matrix()

#Correlation and p value correlations
corr.dat <- cor(dat)
p.mat <- ggcorrplot::cor_pmat(dat)

#Witting Supplementary Table 2
write.table(corr.dat, file = paste0(output.folder, "Correlation_matrix_cells.csv"), sep=",")

#Lower and upper matrix p.values and p.value adjustment
#Done this independently because is a symmetrical matrix
p.vals.lower <- p.mat[lower.tri(p.mat)]
p.vals.upper <- p.mat[upper.tri(p.mat)]
p.mat[lower.tri(p.mat)] = p.adjust(p.vals.lower)
p.mat[upper.tri(p.mat)] = p.adjust(p.vals.upper)

#Witting Supplementary Table 3
write.table(p.mat, file = paste0(output.folder, "Correlation_matrix_cells_FDR.csv"), sep=",")


#Ingoring values for FDR values lower than 0.05
corr.dat[p.mat >= 0.05] <- 0

#For Supplementary Figure2a
hmap <- ComplexHeatmap::Heatmap(corr.dat, name="Pearson\ncorrelation", col = cols.fun, na_col="white",
                        column_names_rot = 60, rect_gp = gpar(col = "white", lwd = 1))

pdf(paste0(output.folder, "Heatmap_cellprop_correlations_all.pdf"), width=6, height=5)
draw(hmap)
dev.off()

```



#Merging clinical info and data
```{r}
#Selecting relevant columns from the clinical info
interesting.clin.columns <- c("TnumberAVL", "centre","refage","incyear","timelastfu","daystoprogression",
                              "finalstatus","progression","figo_stage")

clinical.info.sel <- clinical.info %>% select(interesting.clin.columns)

#Merging the cell proportions and selected clinical info
patient.cell.prop.clin <- merge(patient.cell.prop, clinical.info.sel,
                                by.x = "patient", by.y="TnumberAVL")

#Merging with molecular profile
patient.cell.prop.clin2 <- merge(patient.cell.prop.clin, Mol.profiles.patients, by="patient")

#Re-naming therapy sequence as IDS and PDS
navals <- which(is.na(patient.cell.prop.clin2$therapy_sequence))
patient.cell.prop.clin2$therapy_sequence <- ifelse(grepl("NACT", patient.cell.prop.clin2$therapy_sequence),
                                                   "IDS","PDS")
patient.cell.prop.clin2$therapy_sequence[navals] <- NA

#Factor with levels are for the survival analysis. HRP is used as reference in cox models
levels.molecular <- c("HRP", "BRCAmut/met", "HRD", "CCNE1amp", "Other")
patient.cell.prop.clin2$Molecular.profile2 <- factor(patient.cell.prop.clin2$Molecular.profile2,
                                                   levels = levels.molecular)

#Converting survival info from days to months
patient.cell.prop.clin2$timelastfu <- patient.cell.prop.clin2$timelastfu/30.4
patient.cell.prop.clin2$daystoprogression <- patient.cell.prop.clin2$daystoprogression/30.4

#Structuring data
patient.cell.prop.clin2$figo_stage2 <- NA
patient.cell.prop.clin2$figo_stage2[patient.cell.prop.clin2$figo_stage == "FIGO II"] <- 2
patient.cell.prop.clin2$figo_stage2[patient.cell.prop.clin2$figo_stage == "FIGO III"] <- 3
patient.cell.prop.clin2$figo_stage2[patient.cell.prop.clin2$figo_stage == "FIGO IV"] <- 4

#Structuring Age
patient.cell.prop.clin2$refage2 <- NA
patient.cell.prop.clin2$refage2[patient.cell.prop.clin2$refage < 50] <- 0
patient.cell.prop.clin2$refage2[patient.cell.prop.clin2$refage >= 50 & patient.cell.prop.clin2$refage < 60] <- 1
patient.cell.prop.clin2$refage2[patient.cell.prop.clin2$refage >= 60 & patient.cell.prop.clin2$refage < 70] <- 2
patient.cell.prop.clin2$refage2[patient.cell.prop.clin2$refage >= 70 & patient.cell.prop.clin2$refage < 80] <- 3
patient.cell.prop.clin2$refage2[patient.cell.prop.clin2$refage >= 80] <- 4
```



###Next lines to calculate median the proportion of celltypes by Molecular profiles
```{r}
#Output folder
output.folder = "D:/users/fperez/NKI_TMAs_AF/Analysis_results/01_Cell-proportions_Heatmaps_Nov2023/"


order.celltypes = c("CD8+T.effector", "CD8+T.memory","CD8+T.exhausted",
             "CD4+T.effector","CD4+T.memory","CD4+T.exhausted", "T.regs", "B.cells",
              "Dendritic", "CD11c+MY","Neutrophils","Macrophages","CD163+MP","Other.MY")



#Re-organizing data
melt.patient.cell.prop <- melt(patient.cell.prop)

names(melt.patient.cell.prop) <- c("patient","GlobalCelltype","Proportion")
melt.patient.cell.prop.clin  <- merge(melt.patient.cell.prop, clinical.info.sel, by.x="patient", by.y="TnumberAVL")
melt.patient.cell.prop.clin <- merge(melt.patient.cell.prop.clin, Mol.profiles.patients, by="patient")
melt.patient.cell.prop.clin$therapy_sequence <- ifelse(grepl("NACT", melt.patient.cell.prop.clin$therapy_sequence),
                                                       "IDS","PDS")
melt.patient.cell.prop.clin$GlobalCelltype <- as.vector(melt.patient.cell.prop.clin$GlobalCelltype)
melt.patient.cell.prop.clin$Molecular.profile2[melt.patient.cell.prop.clin$Molecular.profile2 == "BRCAmut/met"] <- "BRCAloss"

#Ignoring cell.types and Molecular.profile "Others" from heatmaps
melt.patient.cell.prop.clin <- melt.patient.cell.prop.clin[melt.patient.cell.prop.clin$GlobalCelltype != "Other",]
melt.patient.cell.prop.clin <- melt.patient.cell.prop.clin[melt.patient.cell.prop.clin$Molecular.profile2 != "Other",]
melt.patient.cell.prop.clin$Molecular.profile2 <- factor(melt.patient.cell.prop.clin$Molecular.profile2,
                                                            levels=c("BRCAloss", "HRD", "HRP","CCNE1amp"))

#Performing statistics independently for IDS and PDS
krustal.df <- NULL
for (t in c("IDS","PDS")){
  dat.sel <- melt.patient.cell.prop.clin[melt.patient.cell.prop.clin$therapy_sequence == t,]
  #Calculating kruskal wallis test, getting p.values
  kruskal.pvals <- lapply(order.celltypes, function(x){
                df <- dat.sel[which(dat.sel$GlobalCelltype == x),]
                k.res <- kruskal.test(Proportion ~ Molecular.profile2, data=df)
                res <- data.frame(Celltype=x, pval=k.res$p.value, y.val = 0.8,
                                  therapy_sequence=t)
                return(res)
  })
  for (i in 1:length(kruskal.pvals)){krustal.df <- rbind(krustal.df, kruskal.pvals[[i]])}
}

#Adjusting p.values
krustal.df$FDR <- p.adjust(krustal.df$pval, method = "BH")

#Calculating the median proportion by molecular profiles for each therapy sequence for plotting
med.prop.profile <- melt.patient.cell.prop.clin %>%
                                group_by(Molecular.profile2, therapy_sequence, GlobalCelltype) %>%
                                summarise(Prop=median(Proportion))


#Changing data to matrix 
mt.prop.profile <- med.prop.profile %>% pivot_wider(names_from = GlobalCelltype, values_from = Prop)
mt.prop.profile <- as.data.frame(mt.prop.profile)
mt.prop.profile[is.na(mt.prop.profile)] <- 0

mt.prop.scaled <- mt.prop.profile #Not really scaled


#Merging the median proportion matrix with the FDR for supplementary table
table <- NULL
for (t in c("IDS","PDS")){
    mt.props.sel <- mt.prop.scaled %>% filter(therapy_sequence == {t}) %>% select(-therapy_sequence)
    row.names(mt.props.sel) <- mt.props.sel[,1]
    mt.props.sel <- mt.props.sel[,-1]
    t.propps <- as.data.frame(t(mt.props.sel))
    t.propps <- rownames_to_column(t.propps, "Celltype")
    t.propps$therapy_sequence = t
    table <- rbind(table, t.propps)
}
table <- table %>% filter(Celltype %in% order.celltypes)

table.fdrs <- merge(table, krustal.df, by=c("Celltype","therapy_sequence")) %>% 
                                    arrange(therapy_sequence) %>% select(-y.val)
write.table(table.fdrs, file=paste0(output.folder,"/Median_cell_proportions_Mprofile.csv"), sep=",",
            row.names = FALSE)

#Generating plots for Supplementary Figure b
for (t in c("PDS","IDS")){
  #Preparing data for creating complex heatmap
  M.profile.order <- mt.prop.scaled %>% filter(therapy_sequence == {{t}}) %>%
                                    pull(Molecular.profile2)
  mt.prop.scaled.sel <- mt.prop.scaled %>% filter(therapy_sequence == {{t}}) %>%
                                    select(all_of(order.celltypes)) %>%
                                    as.matrix()
  mt.prop.scaled.sel <- t(mt.prop.scaled.sel)
  colnames(mt.prop.scaled.sel) <- as.character(M.profile.order)
  
  pvalue <- krustal.df %>% filter(therapy_sequence == t) %>% pull(FDR)
  is_sig = pvalue < 0.05
  pch = rep("P<0.05", 10)
  pch[!is_sig] = NA
  pvalue_col_fun = colorRamp2(c(1, 0.05, 0.001), c("blue", "white", "red"))
  
  ha = rowAnnotation(Significance=pch, col=list(Significance=c("P<0.05"="cyan4")), na_col="grey95")

  col_fun = colorRamp2(c(0, 0.03), c("white", "#9a0000"))

  
  hmap <- Heatmap(mt.prop.scaled.sel, cluster_columns = FALSE, cluster_rows = FALSE, row_names_side = "left",
          name="Median proportion", column_names_rot = 45, rect_gp = gpar(col = "white", lwd = 0.5),
          column_title =t, col = col_fun, right_annotation = ha, na_col="white")
  
  #see how we define the legend for pvalue
  lgd_pvalue = Legend(title = "p-value", col_fun = pvalue_col_fun, at = c(1, 0.05, 0.01, 0.001), 
  labels = c("1", "0.05", "0.01", "0.001"))
  # and one for the significant p-values
  lgd_sig = Legend(pch = "*", type = "points", labels = "< 0.05")
  
  pdf(paste0(output.folder, "Heatmap_Mprofile_ImmuneStromalCancer_prop_", t, ".pdf"), width=4.5, height=5)
  draw(hmap)
  dev.off()
}
```

#Making violin plots with p.values compared to BRCAloss
```{r}

#Selecting data for plotting
dat <- patient.cell.prop.clin2
dat$therapy_sequence[dat$therapy_sequence == "NACT" ] <- "IDS"
dat$Molecular.profile2 <- as.vector(dat$Molecular.profile2)
dat$Molecular.profile2[dat$Molecular.profile2 == "BRCAmut/met"] <- "BRCAloss"
dat <- dat %>% filter(Molecular.profile2 != "Other")
dat$Molecular.profile2 <- factor(dat$Molecular.profile2, levels=c("BRCAloss", "HRD", "HRP","CCNE1amp"))
dat <- dat %>% select(all_of(c("patient", order.celltypes, "Molecular.profile2", "therapy_sequence")))

# #Making shorter some column names
# colnames(dat)[colnames(dat) == "CD8.CD45RO.T.cells"] <- "CD8.CD45RO.T"
# colnames(dat)[colnames(dat) == "CD4.CD45RO.T.cells"] <- "CD4.CD45RO.T"
# colnames(dat)[colnames(dat) == "CD4.PD1.T.cells"] <- "CD4.PD1.T"
# colnames(dat)[colnames(dat) == "CD8.PD1.T.cells"] <- "CD8.PD1.T"

order.celltypes2 <- c("T.cells","CD4+T.effector", "CD4+T.memory", "CD4+T.exhausted",
                      "CD8+T.effector", "CD8+T.memory", "CD8+T.exhausted", "T.regs", "B.cells",
                                 "Neutrophils", "Dendritic", "CD11c+MY",  "CD163+MP", "Macrophages", "Other.MY",
                                  "Other.immune")

for (t in c("PDS","IDS")){
  dat.t <- dat[dat$therapy_sequence == t,]
  cell.types.interest  <- colnames(dat.t)[2:13]
  columns.extra <- c("patient","Molecular.profile2","therapy_sequence")
  
  #Calculating Wilcoxon test, for proportions of cells per molecular profile compared to BRCAloss
  wilxcox.df <- NULL
  wilcox.pvals <-   lapply(cell.types.interest, function(x){
              dat.sel <- dat.t[,c(columns.extra)]
              dat.sel$Proportion <- dat.t[,x]
              res <- p.vals.markers2(dat.sel, Profile.test="BRCAloss", y.val = 0.11)
              res$GlobalCelltype <- x
              return(res)
  })
  for (i in 1:length(wilcox.pvals)){wilxcox.df <- rbind(wilxcox.df, wilcox.pvals[[i]])}
  wilxcox.df$FDR <- p.adjust(wilxcox.df$pval, method ="BH") 
  wilxcox.df$Molecular.profile2 <- factor(wilxcox.df$Molecular.profile2,
                                          levels=c("BRCAloss", "HRD", "HRP","CCNE1amp"))
  
  #Selecting data for plotting
  df.selected <- melt.patient.cell.prop.clin %>% filter(GlobalCelltype %in% order.celltypes2,
                                                        therapy_sequence == t)
  
  df.selected$GlobalCelltype <- factor(df.selected$GlobalCelltype, levels=order.celltypes2)
  

  
  p <- ggplot(df.selected, aes(x=Molecular.profile2,y=Proportion)) + 
        geom_point(aes(col=Molecular.profile2), position= position_jitter(width= .3, height = 0),
                   size= 1.5, alpha = 0.5) +
        geom_violin(fill=NA) + geom_boxplot(fill=NA, width=0.2, outlier.shape = NA) +
        theme(strip.placement = "outside",strip.background = element_blank(),
                 axis.title.y =element_text(size=rel(1.2)),
                 axis.text.x=element_blank(),
                 legend.text=element_text(size=rel(1.1)),
                 strip.text.x = element_text(size=rel(1.2)),
                 legend.title=element_blank(),
                 panel.spacing = unit(1, "lines"),
                 plot.title = element_text(size=rel(1.6)),
                 strip.text = element_text(size=rel(0.6)),
                 panel.border = element_rect(linetype = "solid", fill = NA),
                 panel.background = element_rect(fill = "white"),
                 panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "lightgrey"),
                 panel.grid.minor = element_blank(),
                 panel.grid.major.x = element_blank(),
                 panel.grid.minor.x = element_blank()) + 
         geom_text(data = wilxcox.df[wilxcox.df$pval < 0.05,],
                   aes(x=Molecular.profile2, y=Y.val), label="*", size=8) +
         facet_wrap(~forcats::fct_relevel(GlobalCelltype, order.celltypes2)) + ylim(0,0.13) +
                ggtitle(t)
                
  print(p)
  ggsave(p, file=paste0(output.folder, "Boxplots_cell_proportions_comparisons_", t, ".png"),
         height = 12, width = 15, units = "cm")
  
  
  ##Selecting cell types of interest that have show high abundance in heat map
  dat.t.cells <- dat.t %>% mutate(T.cells =  `CD4+T.effector`, `CD4+T.memory`, `CD4+T.exhausted`,
                                              `CD8+T.effector`, `CD8+T.memory`, `CD8+T.exhausted`, `T.regs`)
  
  melted.dat <- melt(dat.t.cells)
  colnames(melted.dat)[4] <- "GlobalCelltype"
  colnames(melted.dat)[5] <- "Proportion"
  
  
  #For Supplementary Figure 2c
  ##For BRCAloss comparison vs other molecular profiles
  celltypes.of.interest <- c("T.cells", "CD11c+MY", "CD163+MP", "Other.MY")
  df <- melted.dat %>% filter(GlobalCelltype %in% c(celltypes.of.interest))
  
  my_comparisons <- list( c("BRCAloss", "HRD"), c("BRCAloss", "HRP"), c("BRCAloss", "CCNE1amp") )
  p <- ggboxplot(df, "Molecular.profile2", "Proportion") + 
            stat_compare_means(comparisons = my_comparisons, label = "p.signif") +
            geom_violin(fill=NA) +
            xlab("") + ylab("Median proportion in tumor") +
            stat_n_text(size=3) +
            theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size=rel(0.8))) +
            facet_wrap(~forcats::fct_relevel(GlobalCelltype, order.celltypes2))
  ggsave(p, file=paste0(output.folder, "Boxplots_cell_proportions_comparisons_BRCAloss_", t, ".png"),
       height = 19, width = 12, units = "cm")
    ggsave(p, file=paste0(output.folder, "Boxplots_cell_proportions_comparisons_BRCAloss_", t, ".svg"),
       height = 19, width = 12, units = "cm")
}
```










#Calculating cut-off points (median values) per cell type to definy high abundance of cell.type for survival analysis
```{r}
vars.all = c("CD8+T.memory","CD8+T.exhausted","CD8+T.effector",
             "B.cells","CD4+T.effector","CD4+T.exhausted","CD4+T.memory",
             "Macrophages", "CD11c+MY","Neutrophils","CD163+MP","Dendritic","T.regs","Other.MY")


cutpoints <- sapply(vars.all, function(x){
              var = x
              median.var <- median(patient.cell.prop.clin2[,var])
              cutpoint <- median.var
              #If the cut.point is equal to 0 then take the mean
              if ((cutpoint) == 0){
                print(var)
                cutpoint <- mean(patient.cell.prop.clin2[,var])
              }
              return(cutpoint)
})
cutpoints
```


#Functions related with plotting survival plots
```{r}
#Function to arrange a list of survival plots and risk tables in a list
personal.arrange.ggsurvplots <- function(list.surv.plots, list.risk.tables){
                      out.put.figures <- list()
                      count = 0
                      m = 0
                      while (m < length(list.risk.tables)/2){
                        if ((count %% 2 == 0)){
                         for (h in 1:2){
                             index.1 <- 2 * count + h
                             index.2 <- m * 2 + h
                             out.put.figures[[index.1]] <- list.surv.plots[[index.2]]
                         }
                         count = count + 1
                        }else{
                          for (h in 1:2){
                             index.1 <- 2 * count +  h 
                             index.2 <- m * 2 + h
                             out.put.figures[[index.1]] <- list.risk.tables[[index.2]]
                          }
                          m = m + 1
                          count = count + 1
                        }
                      }
                      return(out.put.figures)
}
```


#Performing Kapplan-Meir plots for the abundance of each cell.type, according to therapy sequence
```{r}
output.folder = "D:/users/fperez/NKI_TMAs_AF/Analysis_results/01_Survival_Proportions_November_medians/"

dir.create(paste0(output.folder))

suffix <- c("PDS","IDS") #For naming the output files
list.class <- list(c("PDS"),
                   c("IDS"))


for (j in 1:length(suffix)){
  data.in <- patient.cell.prop.clin2[patient.cell.prop.clin2$therapy_sequence %in% list.class[[j]],]
  lastcol <- ncol(data.in)

  #Generating plots for OS
  plots.OS <- lapply(vars.all, function(x){
    vars = x
    cut.off.value <- cutpoints[names(cutpoints) == vars]
    cat.var <- as.factor(ifelse(data.in[,vars] >  cut.off.value, "high","low"))
    data.in[,lastcol] <- cat.var
    colnames(data.in)[lastcol] <- "aux"
    fit <- survfit(Surv(timelastfu, finalstatus) ~ aux, data = data.in)
    logrank = surv_pvalue(fit, data.in)
    pval=logrank$pval.txt
    p <- ggsurvplot(fit, data = data.in, risk.table = TRUE, p.val=TRUE, conf.int = TRUE,
                    font.legend = c(18, "plain"), font.tickslab = c(16, "plain"),
                    font.x = c(18, "plain"), font.y = c(18, "plain"),
                    tables.y.text = FALSE,  break.time.by=20, xlim = c(0,80),
                    legend.labs=c(paste0(vars, "+"),paste0(vars, "-")),
                    palette = c("#ff5a36", "#4d4dff"))
    p$plot <- p$plot +   guides(colour = guide_legend(nrow = 2))   
    p <- p + ylab("OS probability") + xlab("Time (months)")
    p$plot <- p$plot + theme(legend.key.width = unit(3, "line"),
                             plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"),
                             legend.key.height = unit(1.5, "line"))
    p$plot <- p$plot + ggplot2::annotate("text", x = 65, y = 0.85, label =pval, size = 5)
    p$table <- p$table + theme(plot.margin = margin(t = 0.5,  r = 2, b = 4, l = 3, unit = "cm"),
                               axis.title.y = element_blank(), axis.title.x = element_blank())
    plots <- list(plot=p$plot, table=p$table)
    return(plots)
  })
  fig.OS <- lapply(1:length(plots.OS) , function(x){plots.OS[[x]]$plot})
  tables.OS <- lapply(1:length(plots.OS) , function(x){plots.OS[[x]]$table})
  
  
  #Generating plots for PFI
  plots.PFI <- lapply(vars.all, function(x){
    vars = x
    cut.off.value <- cutpoints[names(cutpoints) == vars]
    cat.var <- as.factor(ifelse(data.in[,vars] >  cut.off.value, "high","low"))
    data.in[,lastcol] <- cat.var
    colnames(data.in)[lastcol] <- "aux"
    fit <- survfit(Surv(daystoprogression, progression) ~ aux, data = data.in)
    logrank = surv_pvalue(fit, data.in)
    pval=logrank$pval.txt
    p <- ggsurvplot(fit, data = data.in, risk.table = TRUE, p.val=TRUE, conf.int = TRUE,
                    font.legend = c(18, "plain"), font.tickslab = c(16, "plain"),
                    font.x = c(18, "plain"), font.y = c(18, "plain"),
                    tables.y.text = FALSE,  break.time.by=20, xlim = c(0,80),
                    legend.labs=c(paste0(vars, "+"),paste0(vars, "-")),
                    palette = c("#ff5a36", "#4d4dff"))
    p$plot <- p$plot + guides(colour = guide_legend(nrow = 2))
    p <- p + ylab("PFI probability") + xlab("Time (months)")
    p$plot <- p$plot + theme(legend.key.width = unit(3, "line"),
                      plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"),
                      legend.key.height = unit(1.5, "line"))
    p$plot <- p$plot + ggplot2::annotate("text", x = 65, y = 0.85, label =pval, size = 5)
    p$table <- p$table + theme(plot.margin = margin(t = 0.5,  r = 2, b = 4, l = 3, unit = "cm"),
                               axis.title.y = element_blank(), axis.title.x = element_blank())
    plots <- list(plot=p$plot, table=p$table)
    return(plots)
  })
  fig.PFI <- lapply(1:length(plots.PFI) , function(x){plots.PFI[[x]]$plot})
  tables.PFI <- lapply(1:length(plots.PFI) , function(x){plots.PFI[[x]]$table})
  
  #Arranging the plots in a list
  out.put.PFI.figures <- personal.arrange.ggsurvplots(fig.PFI, tables.PFI)
  out.put.OS.figures <- personal.arrange.ggsurvplots(fig.OS, tables.OS)

  #Saving plots
  ggsave(file=paste0(output.folder,suffix[j], "_table-fig_PFI.svg"),
         arrangeGrob(grobs = out.put.PFI.figures, ncol = 2),
         width = 23, height = 105, units = "cm")
  
    ggsave(file=paste0(output.folder,suffix[j], "_table-fig_OS.svg"),
         arrangeGrob(grobs = out.put.OS.figures, ncol = 2),
         width = 23, height = 105, units = "cm")
}
```


#Function to calculate cox hazzard ratios
```{r}
#Indicate the variables for which is going to be calculated the cox model
Cox.regresion.variables2 <- function(data, variables, formula='Surv(timelastfu, finalstatus)~'){

   univ_formulas <- sapply(variables,
                           function(x) as.formula(paste(formula, x)))

   univ_models <- lapply( univ_formulas, function(x){coxph(x, data = data, method = "exact")})

   univ_results <- lapply(univ_models,
                          function(x){
                             x <- summary(x)
                             positive <- x$nevent
                             prop.positive <- paste0(round((x$nevent) * 100 /x$n,0),"%")
                             #p.value <- signif(x$wald["pvalue"], digits=2) #p.value for the model
                             wald.test <- signif(x$wald["test"], digits=2)
                             lastrow <- nrow(x$coef) #When using several covariables in the same regression, take the last one used
                             p.value <- signif(x$coef[lastrow, "Pr(>|z|)"], digits=2) #p.value for last covariable
                             beta <- signif(x$coef[lastrow, 1], digits=2);#coeficient beta
                             HR <- signif(x$coef[lastrow, 2], digits=2);#exp(beta)
                             HR.confint.lower <- signif(x$conf.int[lastrow,"lower .95"], 2)
                             HR.confint.upper <- signif(x$conf.int[lastrow,"upper .95"],2)
                             res <- c(HR, p.value, HR.confint.lower,HR.confint.upper)
                             names(res)<-c("HR", "p.value", "HR.confint.lower", "HR.confint.upper")
                             return(res)
                          })
   res <- t(as.data.frame(univ_results, check.names = FALSE))
   result <- as.data.frame(res)
   return(result)
}
```


#Calculating individual cox hazzard ratios per cell.type and including the Molecular profile in each
```{r}
#For Supplementary Figure2d and Supplementary Figure2e
tm <- forest_theme(core=list(bg_params=list(fill = c("grey90", "#eaf8f5", "white"))))


formulas <- c('Surv(timelastfu, finalstatus)~ Molecular.profile2 + refage2  + figo_stage2 +',
              'Surv(daystoprogression, progression)~ Molecular.profile2 + refage2  + figo_stage2 +')

formula.suffix <- c('OS_mprofile_age_figo', 'PFS_mprofile_age_figo')

therapies <- c("IDS","PDS")

df <- patient.cell.prop.clin2
colnames(df) <- gsub("[+]",".",colnames(df)) 

vars.all = c("CD8.T.effector","CD8.T.memory","CD8.T.exhausted",
             "CD4.T.effector","CD4.T.memory", "CD4.T.exhausted","T.regs","B.cells",
             "Dendritic","CD11c.MY","Neutrophils","Macrophages","CD163.MP","Other.MY")


for (index in 1:length(formulas)){
  #Calculating cox hazards in in two groups therapy sequence
  cox.therpies <- lapply(therapies, function(x){
          dat <- df[(df$therapy_sequence %in% x),]
          t.rows <- which(df$therapy_sequence %in% x)

          
          #If proportion of cell.type is equal or above the median in the sampled population, then is equal to 1
          for (j in vars.all){
           #cut.off.value <- cutpoints[names(cutpoints) == j]
           #dat[,j] <- ifelse(dat[,j] > cut.off.value,1,0)
            dat[,j] <- ntile(df[,j], 4)[t.rows]
          }
          #Calculating cox models
          Cox.lymphocites <- Cox.regresion.variables2(dat, vars.all, formula=formulas[index])
          Cox.lymphocites$Treatment <- x
          Cox.lymphocites$Cell.type <- rownames(Cox.lymphocites)
          Cox.lymphocites$p.value <- signif(p.adjust(Cox.lymphocites$p.value, method ="BH"),2)
          return(Cox.lymphocites)
  })
  
  
  #Reordering the cox results in an specific way necessary for the package forestplot
  spacer <- paste(rep(" ", 18), collapse = " ") #For the forest plot table
  cox.groups <- rbind(cox.therpies[[2]], cox.therpies[[1]])
  df.forest <- NULL
  for (i in unique(cox.groups$Cell.type)){
        row.class <- data.frame(Cell.type =i, HR="", space=spacer, p.value=NA, HR.confint.lower=NA,
                                HR.confint.upper=NA, HR2=NA)
        set.df <- cox.groups[cox.groups$Cell.type %in% i,]
        set.df$space <- spacer
        set.df$HR2 <- set.df$HR
        set.df.reordered <- set.df %>%
                                select(Treatment, HR, space, p.value,
                                       HR.confint.lower, HR.confint.upper, HR2)
        colnames(set.df.reordered) <- colnames(row.class)
        row.celltype <- rbind(row.class, set.df.reordered)
        df.forest <- rbind(df.forest, row.celltype)
  }
  colnames(df.forest)[3] <- ""
  colnames(df.forest)[2] <- "HR"
  
  #Adding * to pvalues lower than 0.05
  low.pvals <- which(df.forest$p.value < 0.05)
  tiny.pvals <- which(df.forest$p.value < 0.001)
  df.forest$p.value[low.pvals] <- paste0(df.forest$p.value[low.pvals],"*")
  df.forest$p.value[tiny.pvals] <- "<0.001**"
  df.forest$p.value[which(is.na(df.forest$p.value))] = ""
  colnames(df.forest)[4] <- "FDR"
  
  #indent the subgroup if there is a number in the HR2 column
  df.forest$Cell.type <- ifelse(is.na(df.forest$HR2), 
                                df.forest$Cell.type,
                                #unlist(sapply(df.forest$Cell.type, function(x){cat("\t",x)})))
                                paste0("", df.forest$Cell.type))
  
  
  p <- forest(df.forest[,c(1:4)],
              est = df.forest$HR2,
              lower = df.forest$HR.confint.lower, 
              upper = df.forest$HR.confint.upper,
              sizes = 0.5,
              ci_column = 3,
              ref_line = 1,
              xlim = c(0, 1.5),
              arrow_lab = c("Lower risk", "Higher risk"),
              ticks_at = c(0.5, 1, 1.5),
              theme = tm)
  print(p)
  
  ggplot2::ggsave(filename = paste0(output.folder, "Forest_plot_", formula.suffix[index], ".svg"), plot = p,
                   width = 15, height = 28, units = "cm", device = svglite::svglite)
  
  ggplot2::ggsave(filename = paste0(output.folder, "Forest_plot_", formula.suffix[index], ".png"), plot = p,
                   width = 15, height = 28, units = "cm")
}
```






###Create multivariable cox model with all the variables
```{r}
#For Supplementary Figure2f
#Formulas for OS and PFI
formulas <- c('Surv(timelastfu, finalstatus) ~ .',
              'Surv(daystoprogression, progression)~ .')

#Names for output files
formula.suffix <- c('OS_mprofile_age_figo', 'PFS_mprofile_age_figo')

#Variables for OS and PFI analysis
vars.for.suv <- list(c("timelastfu","finalstatus"),
                     c("daystoprogression","progression"))

#Variables to name files
short.surv.name <- c("OS","PFS")
therapies <- c("IDS","PDS")

#For a second analysis the CD4+ and Macrophages were excluded
df <- patient.cell.prop.clin2 %>% select(-any_of(c("CD4+T.exhausted","CD4+T.memory","Macrophages")))

vars.all = c("CD8+T.memory","CD8+T.exhausted","CD8+T.effector",
             "B.cells","CD4+T.effector",
             "CD11c+MY","Neutrophils","CD163+MP","Dendritic","T.regs","Other.MY")

#Renaming columns for plots
colnames(df)[which(colnames(df) == "refage2")] <- "Age"
colnames(df)[which(colnames(df) == "figo_stage2")] <- "Stage"
colnames(df)[which(colnames(df) == "Molecular.profile2")] <- "Molecular.profile"
colnames(df)[which(colnames(df) == "therapy_sequence")] <- "Therapy.sequence"
df[df$Molecular.profile == "BRCAmut/met","Molecular.profile"] <- "BRCAloss"
#df$Molecular.profile <- as.character(df$Molecular.profile2)



#First iteration for PFS or OS cox regression
for (index in 1:length(formulas)){
  for (c in c("IntraCutoff", "GlobalCutoff")){
    #Generating multivariable cox regression for each therapy sequence
    for (t in therapies){
      dat <- df[(df$Therapy.sequence %in% t),]
      t.rows <- which(df$Therapy.sequence %in% t)
      #Stratifiying patients in four groups according to each cell proportion
      for (j in vars.all){
        if (c == "GlobalCutoff"){
          dat[,j] <- ntile(df[,j],4)[t.rows]
        }else{
          dat[,j] <- ntile(dat[,j], 4) #Intra therapy ntails
        }
      }
      #Selecting columns of intereset for the forest plots
      dat <- dat %>% select(all_of(c("Molecular.profile", "Age" ,
                                     "Stage", vars.all, vars.for.suv[[index]])))
      #Cox model
      model <- coxph(as.formula(formulas[index]), data=dat) 
      #Forest plot
      p <- ggforest(model, data = dat, main=paste0(short.surv.name[index],"_", t,"_", c))
      ggsave(filename = paste0(output.folder, "Forest_plot_multivarCox_", formula.suffix[index], "_",
                                        c, "_", t, ".png"),
                      plot = p,  width = 15, height = 20, units = "cm")
    }
  }
  
  #Cox regression using the entire dataset
  dat <- df
  #Diving all cohort in four groups for each variable
  for (j in vars.all){
    dat[,j] <- ntile(dat[,j], 4)
  }
  
  dat <- dat %>% select(all_of(c("Molecular.profile", "Therapy.sequence", "Stage", "Age",
                          vars.all, vars.for.suv[[index]])))
  #Cox model
  model <- coxph(as.formula(formulas[index]), data=dat)
  #Forest plot
  p <- ggforest(model, data = dat, main=paste0(short.surv.name[index], " Multi-therapy Global"))
  ggsave(filename = paste0(output.folder, "Forest_plot_multivarCox_",
                                    formula.suffix[index], "_NOcd4.png"), plot = p,
                     width = 15, height = 20, units = "cm")
  ggsave(filename = paste0(output.folder, "Forest_plot_multivarCox_",
                                    formula.suffix[index], "_NOcd4.svg"), plot = p,
                     width = 15, height = 20, units = "cm")
}
```








