---
title: "R Notebook"
output: html_notebook
---

# Loading libraries 
```{r}
library(ggplot2)
library(tidyverse)
library(reshape2)
library(survminer)
library(survival)
library(gridExtra)
library(ComplexHeatmap)
library(forestploter)
library(EnvStats)
library(circlize)
```

# Loading relevant files
```{r}
project.folder <- "D:/users/fperez/NKI_TMAs_AF/"
out.put.folder <- "/Analysis_results/02_MetaClusters_November/"

# project.folder <- "/home/fernpere/NKI/"
# out.put.folder <- "/02_MetaClusters_April/"


cells.counts.by.image <- read.table(file=paste0(project.folder,out.put.folder, 
                                               "Metaclusters_abundance_per_core_20231129.csv"), header = TRUE, sep=",")

molecular.profiles <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Tables/Molecular_profiles_patients_20231208.csv",
                                 header=TRUE, sep=",")

cycif2patientIds <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Clinical_intersection/All-slides_cycif2samples.txt",
                               header=TRUE, sep="\t")

cycif2patientIds$imageid <- paste0(cycif2patientIds$cycif.slide, "_"
                                   ,cycif2patientIds$cycif.core.id)

clinical.data <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Clinical_data/T-CycIF.updated.csv", header=TRUE,
                            sep=",")
interesting.columns <- c("TnumberAVL", "figo_stage", "refage", "timelastfu", "finalstatus", "daystoprogression",
                         "progression","therapy_sequence")
clinical.data <- clinical.data[,colnames(clinical.data) %in%  interesting.columns]
clinical.data$therapy_sequence <- ifelse(grepl("NACT", clinical.data$therapy_sequence),"IDS","PDS")
clinical.data$daystoprogression <- clinical.data$daystoprogression / 30.4
clinical.data$timelastfu <- clinical.data$timelastfu / 30.4


# #Structuring data
clinical.data$figo_stage2 <- NA
clinical.data$figo_stage2[clinical.data$figo_stage == "FIGO II"] <- 2
clinical.data$figo_stage2[clinical.data$figo_stage == "FIGO III"] <- 3
clinical.data$figo_stage2[clinical.data$figo_stage == "FIGO IV"] <- 4

# #Structuring Age
clinical.data$refage2 <- NA
clinical.data$refage2[clinical.data$refage < 50] <- 0
clinical.data$refage2[clinical.data$refage >= 50 & clinical.data$refage < 60] <- 1
clinical.data$refage2[clinical.data$refage >= 60 & clinical.data$refage < 70] <- 2
clinical.data$refage2[clinical.data$refage >= 70 & clinical.data$refage < 80] <- 3
clinical.data$refage2[clinical.data$refage >= 80] <- 4

clinical.data
```


```{r}
length(unique(cells.counts.by.image$imageid))
```

#Selecting only patients with at least two cores
```{r}
counts.by.image.p.id <- merge(cells.counts.by.image, cycif2patientIds, by="imageid")
print(paste0("Cell types before selection: ", nrow(counts.by.image.p.id)))

patients.sel <- counts.by.image.p.id %>% group_by(patient) %>%
                        summarise(n.cores=length(unique(imageid))) %>%
                        dplyr::filter(n.cores >= 2) %>% pull(patient)

counts.by.image.p.id <- counts.by.image.p.id[counts.by.image.p.id$patient %in% patients.sel,]
print(paste0("Cell types after selection: ", nrow(counts.by.image.p.id)))
```


#Also for subsequent analysis are ignored patients with less that 5% of cancer cells
```{r}
cutoff.cancer.prop <- 5
# #Selecting only cancer metaclusters
# counts.by.image.p.id <- counts.by.image.p.id[grep("Cancer", counts.by.image.p.id$SubCelltype),]

print(paste0("Input number of patients: ", length(unique(counts.by.image.p.id$patient))))

Cancer.prop.by.patient <- counts.by.image.p.id %>% group_by(patient) %>%
                          mutate(Total.cells=sum(N)) %>%
                          dplyr::filter(grepl("Cancer", SubCelltype)) %>%
                          reframe(Cancer.prop=sum(N) * 100 /Total.cells) %>% unique() %>% arrange(Cancer.prop)

selected.patients <- Cancer.prop.by.patient$patient[Cancer.prop.by.patient$Cancer.prop >= cutoff.cancer.prop]

counts.by.image.p.id <- counts.by.image.p.id %>% dplyr::filter(patient %in% selected.patients,
                                       grepl("Cancer", SubCelltype)) 

print(paste0("Selected number of patients: ", length(unique(counts.by.image.p.id$patient))))
```


# Calculating total proportion by patient
```{r}
#Calculating the relative proportion of cancer cluster by patient
proportions.by.patient  <- counts.by.image.p.id %>% group_by(patient, SubCelltype) %>%
                                        summarise(N.sum=sum(N)) %>% 
                                        mutate(Rel.Prop = N.sum /sum(N.sum))


proportions.by.patient <- proportions.by.patient[,-3]

proportions.by.patient.m <- as.data.frame(proportions.by.patient)  %>% 
                        pivot_wider(names_from = SubCelltype, values_from = Rel.Prop)

proportions.by.patient.m <- as.data.frame(proportions.by.patient.m)
proportions.by.patient.m[is.na(proportions.by.patient.m)] <- 0


proportions.by.patient <- melt(proportions.by.patient.m)
colnames(proportions.by.patient) <- c("patient","SubCelltype","Rel.Prop")

#Adding molecular profile info and therapy sequence
proportions.by.patient.clin <- merge(proportions.by.patient, molecular.profiles, by="patient")

#Renaming IDS instead of NACT
proportions.by.patient.clin$therapy_sequence <- ifelse(grepl("NACT",
                                                             proportions.by.patient.clin$therapy_sequence),"IDS","PDS")

proportions.by.patient.clin$therapy_sequence  <- factor(proportions.by.patient.clin$therapy_sequence,
                                                           levels=c("PDS","IDS"))
#Some patients had NA information for therapy sequence
na.therapies.patients <- molecular.profiles[is.na(molecular.profiles$therapy_sequence),"patient"] 
proportions.by.patient.clin[proportions.by.patient.clin$patient %in% na.therapies.patients,"therapy_sequence"] <- NA


proportions.by.patient.clin$Molecular.profile2[proportions.by.patient.clin$Molecular.profile2 == "BRCAmut/met"] <- "BRCAloss"

proportions.by.patient.clin$Molecular.profile2 <- factor(proportions.by.patient.clin$Molecular.profile2,
                                                         levels=c("BRCAloss","HRD","HRP","CCNE1amp","Other"))

#Renaming clusters for their meaninful names
proportions.by.patient.clin <- proportions.by.patient.clin %>% mutate(SubCelltype = case_when(
                                                            SubCelltype == "CancerClus1" ~ "Idle",
                                                            SubCelltype == "CancerClus2" ~ "Proliferative",
                                                            SubCelltype == "CancerClus3" ~ "EMT",
                                                            SubCelltype == "CancerClus4" ~ "Immunogenic"))
proportions.by.patient.clin
```




# Making plots of general proportion
```{r}
output.folder <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/02_MetaClusters_Feb24/"

therapy.sequence.name <- c("IDS","PDS","ALL")
therapy.sequence.id <- c("IDS","PDS",".")

#Comparisons for Wilcoxon test
my_comparisons <- list( c("BRCAloss", "HRD"), c("BRCAloss", "HRP"), c("BRCAloss", "CCNE1amp") )
order.celltypes2 <- c("Idle", "Immunogenic", "EMT","Proliferative")

#Generating plot for each therapy sequence
for (i in 1:length(therapy.sequence.name)){
  t <- therapy.sequence.name[i]
  j <- therapy.sequence.id[i]
  proportions.by.patient.sel <- proportions.by.patient.clin[grep(j, proportions.by.patient.clin$therapy_sequence),]

  proportions.by.patient.sel <- proportions.by.patient.sel %>% dplyr::filter(Molecular.profile2 != "Other")
  
  p <- ggboxplot(proportions.by.patient.sel, x="Molecular.profile2", y="Rel.Prop") +
          stat_compare_means(comparisons = my_comparisons, label = "p.signif") +
          geom_violin(fill=NA) + ylim(-0.1,1.25) +
                  xlab("") + ylab("Relative proportion in tumor") +
                  stat_n_text(size=3) +
                  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size=rel(0.8))) +
                  facet_wrap(~forcats::fct_relevel(SubCelltype, order.celltypes2))
  print(p)
  ggsave(p, filename=paste0(output.folder, "Boxplot_prop-relative_Allclusters_", t,".png"),
        height = 12, width = 9, units = "cm")
  ggsave(p, filename=paste0(output.folder, "Boxplot_prop-relative_Allclusters_", t,".svg"),
      height = 12, width = 9, units = "cm")
}
```

#Functions to calculate wilcox.test for each molecular profile compared to the Profile.test
```{r}
#This is for proportion
p.vals.markers1 <- function(data, Profile.test="HRD", y.val=25, alt.hypothesis="greater"){
       test.dat <- data$Proportion[data$Molecular.profile2 == Profile.test]
       p.vals <- sapply(unique(data$Molecular.profile2), function(x){
                mol.dat <- data$Proportion[data$Molecular.profile2 == x]
                wtest <- wilcox.test(test.dat, mol.dat, alternative=alt.hypothesis)
                wtest$p.value
       })
      #Organizing p-values in data.frame for plotting using geom_text
      p.vals.dataframe <- data.frame(pval=p.vals,
                           Molecular.profile2=unique(data$Molecular.profile2),
                           Y.val=y.val)
      return(p.vals.dataframe)
}

#This is for relative proportions
p.vals.markers2 <- function(data, Profile.test="HRD", y.val=25){
       test.dat <- data$Rel.Prop[data$Molecular.profile2 == Profile.test]
       p.vals <- sapply(unique(data$Molecular.profile2), function(x){
                mol.dat <- data$Rel.Prop[data$Molecular.profile2 == x]
                wtest <- wilcox.test(test.dat, mol.dat)
                wtest$p.value
       })
      #Organizing p-values in data.frame for plotting using geom_text
      p.vals.dataframe <- data.frame(pval=p.vals,
                           Molecular.profile2=unique(data$Molecular.profile2),
                           Y.val=y.val)
      return(p.vals.dataframe)
}
```


#Heatmaps with median proportion per cancer metacluster
```{r}
#For complex heatmaps
col_fun = colorRamp2(c(0, 0.35), c("white", "#9a0000"))

#Factor with clusters order
proportions.c <- proportions.by.patient.clin
clusters.order <- rev(c("Idle","Immunogenic", "EMT", "Proliferative"))

proportions.c$SubCelltype <- as.character(proportions.c$SubCelltype)
proportions.c$SubCelltype <- factor(proportions.c$SubCelltype, levels=rev(clusters.order))


#Calculating the median value per molecular profile
median.by.profiles <- proportions.c %>% group_by(Molecular.profile2, therapy_sequence, SubCelltype) %>%
                                   summarise(Prop=median(Rel.Prop))

#Changing data to matrix 
mt.prop.profile <- median.by.profiles %>% pivot_wider(names_from = SubCelltype, values_from = Prop)
mt.prop.profile <- as.data.frame(mt.prop.profile)
mt.prop.profile[is.na(mt.prop.profile)] <- 0

#Scaling by therapy sequence
# mt.prop.scaled <- mt.prop.profile %>% group_by(therapy_sequence) %>% 
#                    mutate(across(where(is.numeric), ~ as.numeric(scale(.)))) %>%
#                    as.data.frame()
mt.prop.scaled <- mt.prop.profile

mt.prop.scaled <- mt.prop.scaled[mt.prop.scaled$Molecular.profile2 != "Other",]


for (t in c("PDS","IDS")){
  #Preparing data for creating complex heatmap
  M.profile.order <- mt.prop.scaled %>% dplyr::filter(therapy_sequence == {{t}}) %>%
                                    pull(Molecular.profile2)
  mt.prop.scaled.sel <- mt.prop.scaled %>% dplyr::filter(therapy_sequence == {{t}}) %>%
                                    select(-Molecular.profile2, -therapy_sequence) %>% as.matrix()
  
  mt.prop.scaled.sel <- t(mt.prop.scaled.sel)
  colnames(mt.prop.scaled.sel) <- as.character(M.profile.order)
  
  #Calculating kruskal.test for each Metacluster_proportion by patient
  dat.sel <- proportions.c %>% filter(Molecular.profile2 != "Other",
                                      therapy_sequence=={{t}})
  #Calculating p.values
  krustal.df <- NULL
  kruskal.pvals <- lapply(clusters.order, function(x){
                df <- dat.sel[which(dat.sel$SubCelltype == x),]
                k.res <- kruskal.test(Rel.Prop ~ Molecular.profile2, data=df)
                res <- data.frame(SubCelltype=x, pval=k.res$p.value, y.val = 0.8,
                                  therapy_sequence=t)
                return(res)
  })
  for (i in 1:length(kruskal.pvals)){krustal.df <- rbind(krustal.df, kruskal.pvals[[i]])}
  print(krustal.df)
  

  hmap <- Heatmap(mt.prop.scaled.sel, cluster_columns = FALSE, cluster_rows = FALSE, row_names_side = "left",
          name="Median\nproportion", rect_gp = gpar(col = "white", lwd = 0.5), col = col_fun,
          column_title = t, column_title_gp = gpar(fontsize = 16), row_names_gp = gpar(fontsize = 12),
          column_names_gp = gpar(fontsize = 14))
  
  pdf(paste0(output.folder, "Heatmap_CancerMetaclusters_relative_prop_", t, ".pdf"), width=3, height=3)
  draw(hmap)
  dev.off()
}

```


#Function to create survival plots for each type of metacluster
```{r}
#The function takes as input the variables (vars) that will be used for Kaplan-Meir. The vars should be columns of data.in
#Second input is data.in, a matrix with each row by patients and columns are numerical values.
#A data.frame with cutoff points to stratify each var. Values above the cutoff point are defined as high, otherwise as low.

#For OS
plot.OS <- function(vars, data.in, cutpoints){
    cut.off.value <- cutpoints[names(cutpoints) == vars]
    cat.var <- as.factor(ifelse(data.in[,vars] >  cut.off.value, "high","low"))
    data.in <- cbind(data.in, cat.var)
    fit <- survfit(Surv(timelastfu, finalstatus) ~ cat.var, data = data.in)
    logrank = surv_pvalue(fit, data.in)
    pval=logrank$pval.txt
    p <- ggsurvplot(fit, data = data.in, risk.table = TRUE, pval=TRUE, conf.int = TRUE,
                    font.legend = c(18, "plain"), font.tickslab = c(16, "plain"),
                    font.x = c(18, "plain"), font.y = c(18, "plain"),
                    tables.y.text = FALSE,  break.time.by=30, xlim = c(0,120),
                    legend.labs=c(paste0(vars, "_high"), paste0(vars, "_low")),
                    palette = c("#ff5a36", "#4d4dff"))
    p$plot <- p$plot +   guides(colour = guide_legend(nrow = 2))
    p <- p + ylab("OS probability") + xlab("Time (days)")
    p$plot <- p$plot + theme(legend.key.width = unit(3, "line"),
                             plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"),
                             legend.key.height = unit(1.5, "line"),
                             legend.title = element_blank())
    p$plot <- p$plot + ggplot2::annotate("text", x = 90, y = 0.85, label =pval, size = 5)
    p$table <- p$table + theme(plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"),
                               axis.title.y = element_blank())
    plots <- list(plot=p$plot, table=p$table)
    return(plots)
}

#For PFS
plot.PFS <- function(vars, data.in, cutpoints){
    cut.off.value <- cutpoints[names(cutpoints) == vars]
    cat.var <- as.factor(ifelse(data.in[,vars] >  cut.off.value, "high","low"))
    data.in <- cbind(data.in, cat.var)
    fit <- survfit(Surv(daystoprogression, progression) ~ cat.var, data = data.in)
    logrank = surv_pvalue(fit, data.in)
    pval=logrank$pval.txt
    p <- ggsurvplot(fit, data = data.in, risk.table = TRUE, p.val=TRUE, conf.int = TRUE,
                    font.legend = c(18, "plain"), font.tickslab = c(16, "plain"),
                    font.x = c(18, "plain"), font.y = c(18, "plain"),
                    tables.y.text = FALSE,  break.time.by=15, xlim = c(0,60),
                    legend.labs=c(paste0(vars, "+"),paste0(vars, "-")),
                    palette = c("#ff5a36", "#4d4dff"))  
    p$plot <- p$plot +  guides(colour = guide_legend(nrow = 2))
    p <- p + ylab("PFS probability") + xlab("Time (days)")
    p$plot <- p$plot + theme(legend.key.width = unit(3, "line"),
                             plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"),
                             legend.key.height = unit(1.5, "line"),
                             legend.title = element_blank())
    p$plot <- p$plot + ggplot2::annotate("text", x = 40, y = 0.85, label =pval, size = 5)
    p$table <- p$table + theme(plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"),
                               axis.title.y = element_blank())
    plots <- list(plot=p$plot, table=p$table)
    return(plots)
}


#Indicate the variables for which is going to be calculated the cox model
Cox.regresion.variables2 <- function(data, variables, formula='Surv(timelastfu, finalstatus)~'){

   univ_formulas <- sapply(variables,
                           function(x) as.formula(paste(formula, x)))

   univ_models <- lapply( univ_formulas, function(x){coxph(x, data = data, method = "exact")})

   univ_results <- lapply(univ_models,
                          function(x){
                             x <- summary(x)
                             positive <- x$nevent
                             prop.positive <- paste0(round((x$nevent) * 100 /x$n,0),"%")
                             wald.test <- signif(x$wald["test"], digits=2)
                             lastrow <- nrow(x$coef) #When using several covariables in the same regression, take the last one used
                             p.value <- signif(x$coef[lastrow, "Pr(>|z|)"], digits=2) #p.value for last covariable
                             beta <- signif(x$coef[lastrow, 1], digits=2);#coeficient beta
                             HR <- signif(x$coef[lastrow, 2], digits=2);#exp(beta)
                             HR.confint.lower <- signif(x$conf.int[lastrow,"lower .95"], 2)
                             HR.confint.upper <- signif(x$conf.int[lastrow,"upper .95"],2)
                             res <- c(HR, p.value, HR.confint.lower,HR.confint.upper)
                             names(res)<-c("HR", "p.value", "HR.confint.lower", "HR.confint.upper")
                             return(res)
                          })
   res <- t(as.data.frame(univ_results, check.names = FALSE))
   result <- as.data.frame(res)
   return(result)
}
```




#Generating survival analysis using the plots
```{r}
out.put.folder <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/02_MetaClusters_Feb24/Survival/"


proportions.by.patient.m <- proportions.c %>% select(-Molecular.profile2, -therapy_sequence) %>%
                                          pivot_wider(names_from = SubCelltype, values_from = Rel.Prop)

clusters.order <- c("Idle","Immunogenic", "EMT","Proliferative")


proportions.rel.surv <- merge(proportions.by.patient.m, clinical.data, by.x="patient",
                                     by.y="TnumberAVL", all.x=TRUE)

#Calculating median value in  the cohort for each cluster
cut.points.rel <- sapply(clusters.order, function(x){
        cutpoint <- median(proportions.rel.surv[,x])
        return(cutpoint)
})


#Generating survival plots independently for PDS and IDS
suffix <- c("PDS","IDS")
for (i in suffix){
  data.surv <- proportions.rel.surv[proportions.rel.surv$therapy_sequence == i,]
  
  surv.tables.os <- lapply(clusters.order, plot.OS, data.surv, cut.points.rel)
  
  #Overall survival analysis
  fig.OS <- lapply(1:length(surv.tables.os) , function(x){surv.tables.os[[x]]$plot})
  tables.OS <- lapply(1:length(surv.tables.os) , function(x){surv.tables.os[[x]]$table})
  
  ggsave(file=paste0(out.put.folder, "KM_OS_Cancer_metacluster", "_", i, ".svg"),
           arrangeGrob(grobs = fig.OS, ncol = 2), width = 21, height = 18, units = "cm")
  ggsave(file=paste0(out.put.folder, "KM_OS_metacluster", "_", i, "-table.svg"),
         arrangeGrob(grobs = tables.OS, ncol = 2), width = 21, height = 7, units = "cm")
  
  ##PFS survival analysis
  surv.tables.pfs <- lapply(clusters.order, plot.PFS, data.surv, cut.points.rel)
  fig.pfs <- lapply(1:length(surv.tables.pfs) , function(x){surv.tables.pfs[[x]]$plot})
  tables.pfs <- lapply(1:length(surv.tables.pfs) , function(x){surv.tables.pfs[[x]]$table})

  ggsave(file=paste0(out.put.folder, "KM_PFS_Cancer_metacluster", "_", i, ".svg"),
           arrangeGrob(grobs = fig.pfs, ncol = 2), width = 21, height = 18, units = "cm")
  ggsave(file=paste0(out.put.folder, "KM_PFS_Cancer_metacluster", "_", i, "-table.svg"),
         arrangeGrob(grobs = tables.pfs, ncol = 2), width = 21, height = 7, units = "cm")
}
```




#Calculating individual cox hazzard ratios per cell.type and including the Molecular profile in each
```{r}
tm <- forest_theme(core=list(bg_params=list(fill = c("grey90", "#eaf8f5", "white"))))


molecular.profile.set <- molecular.profiles[,-3]
proportions.rel.surv2 <- merge(proportions.rel.surv, molecular.profile.set, by="patient")

formulas <- c('Surv(timelastfu, finalstatus)~ Molecular.profile2 + refage2  + figo_stage2 +',
              'Surv(daystoprogression, progression)~ Molecular.profile2 + refage2  + figo_stage2 +')

formula.suffix <- c('OS_mprofile_age_figo', 'PFS_mprofile_age_figo')

clusters.order <- c("Idle","Immunogenic","EMT","Proliferative")


therapies <- c("IDS","PDS")

#Plotting forest plots idependently for PDS and IDS
for (index in 1:length(formulas)){
  #Calculating cox hazards in in two groups therapy sequence
  cox.therpies <- lapply(therapies, function(x){
          dat <- proportions.rel.surv2[(proportions.rel.surv2$therapy_sequence %in% x),]
          t.rows <- which(proportions.rel.surv2$therapy_sequence %in% x)

          #Stratifying the whole dataset by quartiles
          for (j in clusters.order){
            dat[,j] <- ntile(proportions.rel.surv2[,j], 4)[t.rows]
          }
          #Calculating cox models
          Cox.metacluster <- Cox.regresion.variables2(dat, clusters.order, formula=formulas[index])
          Cox.metacluster$Treatment <- x
          Cox.metacluster$Cell.type <- rownames(Cox.metacluster)
          #Cox.metacluster$p.value <- signif(p.adjust(Cox.metacluster$p.value, method ="BH"),2)
          return(Cox.metacluster)
  })
  
  
  #Reordering the cox results in an specific way necessary for the package forestplot
  spacer <- paste(rep(" ", 15), collapse = " ") #For the forest plot table
  cox.groups <- rbind(cox.therpies[[2]], cox.therpies[[1]])
  
  df.forest <- NULL
  for (i in unique(cox.groups$Cell.type)){
        row.class <- data.frame(Cell.type =i, HR="", space=spacer, p.value=NA, HR.confint.lower=NA,
                                HR.confint.upper=NA, HR2=NA)
        set.df <- cox.groups[cox.groups$Cell.type %in% i,]
        set.df$space <- spacer
        set.df$HR2 <- set.df$HR
        set.df.reordered <- set.df %>% select(Treatment, HR, space, p.value, HR.confint.lower, HR.confint.upper, HR2)
        colnames(set.df.reordered) <- colnames(row.class)
        row.celltype <- rbind(row.class, set.df.reordered)
        df.forest <- rbind(df.forest, row.celltype)
  }
  colnames(df.forest)[3] <- ""
  colnames(df.forest)[2] <- "HR"
  
  #Adding * to pvalues lower than 0.05
  low.pvals <- which(df.forest$p.value < 0.05)
  tiny.pvals <- which(df.forest$p.value < 0.001)
  df.forest$p.value[low.pvals] <- paste0(df.forest$p.value[low.pvals],"*")
  df.forest$p.value[tiny.pvals] <- "<0.001**"
  df.forest$p.value[which(is.na(df.forest$p.value))] = ""
  colnames(df.forest)[4] <- "P.val"
  
  #indent the subgroup if there is a number in the HR2 column
  df.forest$Cell.type <- ifelse(is.na(df.forest$HR2), 
                                df.forest$Cell.type,
                                #unlist(sapply(df.forest$Cell.type, function(x){cat("\t",x)})))
                                paste0("", df.forest$Cell.type))
  
  
  p <- forest(df.forest[,c(1:4)],
              est = df.forest$HR2,
              lower = df.forest$HR.confint.lower, 
              upper = df.forest$HR.confint.upper,
              sizes = 0.5,
              ci_column = 3,
              ref_line = 1,
              xlim = c(0, 2),
              arrow_lab = c("Lower risk", "Higher risk"),
              ticks_at = c(0.5, 1, 1.5, 2.0),
              theme = tm)
  print(p)
  
  ggplot2::ggsave(filename = paste0(out.put.folder, "Forest_plot_", formula.suffix[index], ".svg"), plot = p,
                   width = 11, height = 11, units = "cm", device = svglite::svglite)

  ggplot2::ggsave(filename = paste0(out.put.folder, "Forest_plot_", formula.suffix[index], ".png"), plot = p,
                   width = 11, height = 11, units = "cm")
}
```






#Calculating individual cox hazzard ratios per cell.type and including the Molecular profile in each
```{r}
out.put.folder <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/02_MetaClusters_Feb24/Survival/"


proportions.by.patient.m <- proportions.by.patient.clin %>% select(-therapy_sequence) %>%
                                                      pivot_wider(names_from = SubCelltype, values_from = Rel.Prop)

clusters.order <- c("Idle","Immunogenic", "EMT","Proliferative")


proportions.rel.surv <- merge(proportions.by.patient.m, clinical.data, by.x="patient",
                                     by.y="TnumberAVL", all.x=TRUE)


colnames(proportions.rel.surv)[colnames(proportions.rel.surv) == "Molecular.profile2"] <- "Molecular.profile"
colnames(proportions.rel.surv)[colnames(proportions.rel.surv) == "therapy_sequence"] <- "Therapy.sequence"
colnames(proportions.rel.surv)[colnames(proportions.rel.surv) == "figo_stage2"] <- "Stage"
colnames(proportions.rel.surv)[colnames(proportions.rel.surv) == "refage2"] <- "Age"

proportions.rel.surv$Molecular.profile <- factor(proportions.rel.surv$Molecular.profile,
                                                 levels=c("HRP","BRCAloss","HRD", "CCNE1amp"))


#For Supplementary Figure2f
#Formulas for OS and PFS
formulas <- c('Surv(timelastfu, finalstatus) ~ .',
              'Surv(daystoprogression, progression)~ .')

#Names for output files
formula.suffix <- c('OS_mprofile_age_figo', 'PFS_mprofile_age_figo')

#Variables for OS and PFI analysis
vars.for.suv <- list(c("timelastfu","finalstatus"),
                     c("daystoprogression","progression"))

#Variables to name files
short.surv.name <- c("OS","PFS")
therapies <- c("IDS","PDS")

#One plot for PFS and another for OS
for (index in 1:length(formulas)){
  #Cox regression using the entire dataset
  dat <- proportions.rel.surv
  #Diving all cohort in four groups for each variable
  for (j in clusters.order){
    dat[,j] <- ntile(dat[,j], 2)
  }
  
  dat <- dat %>% select(all_of(c("Molecular.profile", "Therapy.sequence", "Stage", "Age",
                          clusters.order, vars.for.suv[[index]])))
  #Cox model
  model <- coxph(as.formula(formulas[index]), data=dat)
  #Forest plot
  p <- ggforest(model, data = dat, main=paste0(short.surv.name[index], " Multi-therapy Global"))
  ggsave(filename = paste0(out.put.folder, "Forest_plot_MetaClusters_multivarCox_",
                                    formula.suffix[index], ".png"), plot = p,
                     width = 15, height = 14, units = "cm")
  ggsave(filename = paste0(out.put.folder, "Forest_plot_MetaClusters_multivarCox_",
                                    formula.suffix[index], ".svg"), plot = p,
                     width = 15, height = 14, units = "cm")
}
```
