---
title: "R Notebook"
output: html_notebook
---

# Loading libraries 
```{r}
library(ggplot2)
library(tidyverse)
library(reshape2)
library(survminer)
library(survival)
library(gridExtra)
library(ComplexHeatmap)
library(forestploter)
```

# Loading relevant files
```{r}
project.folder <- "D:/users/fperez/NKI_TMAs_AF/"
out.put.folder <- "/Analysis_results/02_MetaClusters_November/"

# project.folder <- "/home/fernpere/NKI/"
# out.put.folder <- "/02_MetaClusters_April/"

proportions.by.image <- read.table(file=paste0(project.folder,out.put.folder, 
                                               "Metaclusters_abundance_per_core_20231129.csv"), header = TRUE, sep=",")

molecular.profiles <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Tables/Molecular_profiles_patients_20231208.csv",
                                 header=TRUE, sep=",")

cycif2patientIds <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Clinical_intersection/All-slides_cycif2samples.txt",
                               header=TRUE, sep="\t")

cycif2patientIds$imageid <- paste0(cycif2patientIds$cycif.slide, "_"
                                   ,cycif2patientIds$cycif.core.id)

clinical.data <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Clinical_data/T-CycIF.updated.csv", header=TRUE,
                            sep=",")
interesting.columns <- c("TnumberAVL", "figo_stage", "refage", "timelastfu", "finalstatus", "daystoprogression",
                         "progression","therapy_sequence")
clinical.data <- clinical.data[,colnames(clinical.data) %in%  interesting.columns]
clinical.data$therapy_sequence <- ifelse(grepl("NACT", clinical.data$therapy_sequence),"IDS","PDS")
clinical.data$daystoprogression <- clinical.data$daystoprogression / 30.4
clinical.data$timelastfu <- clinical.data$timelastfu / 30.4


# #Structuring data
clinical.data$figo_stage2 <- NA
clinical.data$figo_stage2[clinical.data$figo_stage == "FIGO II"] <- 2
clinical.data$figo_stage2[clinical.data$figo_stage == "FIGO III"] <- 3
clinical.data$figo_stage2[clinical.data$figo_stage == "FIGO IV"] <- 4

# #Structuring Age
clinical.data$refage2 <- NA
clinical.data$refage2[clinical.data$refage < 50] <- 0
clinical.data$refage2[clinical.data$refage >= 50 & clinical.data$refage < 60] <- 1
clinical.data$refage2[clinical.data$refage >= 60 & clinical.data$refage < 70] <- 2
clinical.data$refage2[clinical.data$refage >= 70 & clinical.data$refage < 80] <- 3
clinical.data$refage2[clinical.data$refage >= 80] <- 4

clinical.data
```


```{r}
length(unique(proportions.by.image$imageid))
```

#Selecting only patients with at least two cores
```{r}
proportions.by.image.p.id <- merge(proportions.by.image, cycif2patientIds, by="imageid")
print(dim(proportions.by.image.p.id))

patients.sel <- proportions.by.image.p.id %>% group_by(patient) %>%
                        summarise(n.cores=length(unique(imageid))) %>%
                        dplyr::filter(n.cores >= 2) %>% pull(patient)

proportions.by.image.p.id <- proportions.by.image.p.id[proportions.by.image.p.id$patient %in% patients.sel,]
print(dim(proportions.by.image.p.id))
```


# Calculating total proportion by patient
```{r}
#Proportions.by.image
proportions.by.patient  <- proportions.by.image.p.id %>% group_by(patient, SubCelltype) %>%
                                        summarise(N.sum=sum(N)) %>% 
                                        mutate(Proportion = N.sum * 100 /sum(N.sum))

proportions.by.patient <- proportions.by.patient[,-3]

proportions.by.patient.m <- as.data.frame(proportions.by.patient)  %>% 
                        pivot_wider(names_from = SubCelltype, values_from = Proportion)

proportions.by.patient.m <- as.data.frame(proportions.by.patient.m)
proportions.by.patient.m[is.na(proportions.by.patient.m)] <- 0


proportions.by.patient <- melt(proportions.by.patient.m)
colnames(proportions.by.patient) <- c("patient","SubCelltype","Proportion")

na.therapies.patients <- molecular.profiles[is.na(molecular.profiles$therapy_sequence),"patient"] 
proportions.by.patient.clin <- merge(proportions.by.patient, molecular.profiles, by="patient")

proportions.by.patient.clin$therapy_sequence <- ifelse(grepl("NACT",
                                                             proportions.by.patient.clin$therapy_sequence),"IDS","PDS")

proportions.by.patient.clin$therapy_sequence  <- factor(proportions.by.patient.clin$therapy_sequence,
                                                           levels=c("PDS","IDS"))

proportions.by.patient.clin[proportions.by.patient.clin$patient %in% na.therapies.patients,"therapy_sequence"] <- NA


proportions.by.patient.clin$Molecular.profile2[proportions.by.patient.clin$Molecular.profile2 == "BRCAmut/met"] <- "BRCAloss"

proportions.by.patient.clin$Molecular.profile2 <- factor(proportions.by.patient.clin$Molecular.profile2,
                                                         levels=c("BRCAloss","HRD","HRP","CCNE1amp","Other"))

```




# Making plots of general proportion
```{r}
output.folder <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/02_MetaClusters_November/"

therapy.sequence.name <- c("IDS","PDS","ALL")
therapy.sequence.id <- c("IDS","PDS",".")

for (i in 1:length(therapy.sequence.name)){
  t <- therapy.sequence.name[i]
  j <- therapy.sequence.id[i]
  proportions.by.patient.sel <- proportions.by.patient.clin[grep(j, proportions.by.patient.clin$therapy_sequence),]

  c <- c("Cancer")
  proportions.c <- proportions.by.patient.sel[grep(c, proportions.by.patient.sel$SubCelltype),]
  
  proportions.c <- proportions.c %>% group_by(patient) %>% mutate(Rel.Prop = Proportion/sum(Proportion)) %>%
                   dplyr::filter(Molecular.profile2 != "Other")
  
  p1 <- ggplot(proportions.c, aes(x=Molecular.profile2,y=Rel.Prop, fill=Molecular.profile2)) +
    geom_violin() + geom_boxplot(outlier.shape = NA, width=0.1, fill=NA)  +
    scale_fill_manual(values=c("red3","#ffa500","green3", "dodgerblue3","#9932CC")) + facet_wrap(~SubCelltype) +
    theme_bw() + xlab("Molecular profile") + theme(axis.text.x = element_blank())
  print(p1)
  ggsave(p1, filename=paste0(output.folder, "Boxplot_prop-relative_Allclusters_", c,"_",t,".svg"),
          height = 10, width = 11, units = "cm")
}
```


```{r}
#This is for proportion
p.vals.markers1 <- function(data, Profile.test="HRD", y.val=25, alt.hypothesis="greater"){
       test.dat <- data$Proportion[data$Molecular.profile2 == Profile.test]
       p.vals <- sapply(unique(data$Molecular.profile2), function(x){
                mol.dat <- data$Proportion[data$Molecular.profile2 == x]
                wtest <- wilcox.test(test.dat, mol.dat, alternative=alt.hypothesis)
                wtest$p.value
       })
      #Organizing p-values in data.frame for plotting using geom_text
      p.vals.dataframe <- data.frame(pval=p.vals,
                           Molecular.profile2=unique(data$Molecular.profile2),
                           Y.val=y.val)
      return(p.vals.dataframe)
}

#This is for relative proportions
p.vals.markers2 <- function(data, Profile.test="HRD", y.val=25){
       test.dat <- data$Rel.Prop[data$Molecular.profile2 == Profile.test]
       p.vals <- sapply(unique(data$Molecular.profile2), function(x){
                mol.dat <- data$Rel.Prop[data$Molecular.profile2 == x]
                wtest <- wilcox.test(test.dat, mol.dat)
                wtest$p.value
       })
      #Organizing p-values in data.frame for plotting using geom_text
      p.vals.dataframe <- data.frame(pval=p.vals,
                           Molecular.profile2=unique(data$Molecular.profile2),
                           Y.val=y.val)
      return(p.vals.dataframe)
}
```


# Boxplots for cancer cells
```{r}
# Boxplots for cancer cells metaclusters by mollecular profiles
output.folder <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/02_MetaClusters_November/"
colours <- c("red3","#ffa500","green3", "dodgerblue3","#9932CC")

#Relative proportion by patient
c <- "Cancer"
proportions.c <- proportions.by.patient.clin[grep(c, proportions.by.patient.clin$SubCelltype),]
proportions.c <- proportions.c %>% group_by(patient) %>% mutate(Rel.Prop = Proportion/sum(Proportion)) %>%
                 dplyr::filter(Molecular.profile2 != "Other") #Ignoring "Others also ingores NA for therapy sequence

therapies <- c("IDS","PDS")

for (clus in c("CancerClus3","CancerClus4")){
  dat <- proportions.c[proportions.c$SubCelltype == clus,]
  #Calculating p.values
  p.vals.dataframe <- NULL
  p.vals <- lapply(therapies, function(x){
        dat.sel <- dat[dat$therapy_sequence == x,]
        p.vals.dataframe <- p.vals.markers2(dat.sel, Profile.test="BRCAloss",
                                          y.val=0.65)
        p.vals.dataframe$therapy_sequence  <- x
        return(p.vals.dataframe)
        })
  for (i in 1:length(p.vals)){p.vals.dataframe <- rbind(p.vals.dataframe, p.vals[[i]])}
  print(p.vals.dataframe)
  #Plotting boxplots parired by therapy
  p <- ggplot(proportions.c[proportions.c$SubCelltype == clus,],
              aes(x=Molecular.profile2,y=Rel.Prop, fill=Molecular.profile2)) +
     geom_violin() + geom_boxplot(outlier.shape = NA, width=0.1, fill=NA)  + theme_bw() + xlab("Molecular profile") +
    theme(axis.text.x = element_blank()) + scale_fill_manual(values=colours) +
      geom_text(data = p.vals.dataframe[p.vals.dataframe$pval < 0.05,], aes(x=Molecular.profile2, y=Y.val),
              label="*", size=8) + ylim(0,0.85) + ggtitle(clus) + facet_wrap(~therapy_sequence)
  print(p)
  ggsave(p, filename=paste0(output.folder, "Boxplot_", clus,".svg"), height = 8, width = 14, units = "cm")
}

```


#Heatmaps with median proportion per cancer metacluster
```{r}
clusters.order <- c("Basal","EMT","Proliferative","Immunogenic")

map <- setNames(clusters.order, c("CancerClus1","CancerClus3","CancerClus2","CancerClus4"))

c <- "Cancer"
proportions.c <- proportions.by.patient.clin[grep(c, proportions.by.patient.clin$SubCelltype),]
proportions.c$SubCelltype <- as.character(proportions.c$SubCelltype)
proportions.c$SubCelltype <- map[proportions.c$SubCelltype]
proportions.c$SubCelltype <- factor(proportions.c$SubCelltype, levels=rev(clusters.order))
proportions.c <- proportions.c %>% group_by(patient) %>% mutate(Rel.Prop = Proportion/sum(Proportion))

median.by.profiles <- proportions.c %>% group_by(Molecular.profile2, therapy_sequence, SubCelltype) %>%
                                   summarise(Prop=median(Rel.Prop))

#Changing data to matrix 
mt.prop.profile <- median.by.profiles %>% pivot_wider(names_from = SubCelltype, values_from = Prop)
mt.prop.profile <- as.data.frame(mt.prop.profile)
mt.prop.profile[is.na(mt.prop.profile)] <- 0


#Scaling by therapy sequence
mt.prop.scaled <- mt.prop.profile %>% group_by(therapy_sequence) %>% 
                   mutate(across(where(is.numeric), ~ as.numeric(scale(.)))) %>%
                   as.data.frame()

mt.prop.scaled <- mt.prop.scaled[mt.prop.scaled$Molecular.profile2 != "Other",]

for (t in c("PDS","IDS")){
  #Preparing data for creating complex heatmap
  M.profile.order <- mt.prop.scaled %>% dplyr::filter(therapy_sequence == {{t}}) %>%
                                    pull(Molecular.profile2)
  mt.prop.scaled.sel <- mt.prop.scaled %>% dplyr::filter(therapy_sequence == {{t}}) %>%
                                    select(-Molecular.profile2, -therapy_sequence) %>% as.matrix()
  
  mt.prop.scaled.sel <- t(mt.prop.scaled.sel)
  colnames(mt.prop.scaled.sel) <- as.character(M.profile.order)

  hmap <- Heatmap(mt.prop.scaled.sel, cluster_columns = FALSE, cluster_rows = FALSE, row_names_side = "left",
          name="Median proportion \n (Z-score)", rect_gp = gpar(col = "white", lwd = 0.5),
          column_title = paste0(t," samples"), row_names_gp = gpar(fontsize = 12), column_names_gp = gpar(fontsize = 14))
  pdf(paste0(output.folder, "Heatmap_CancerMetaclusters_relative_prop_", t, ".pdf"), width=3.5, height=3)
  draw(hmap)
  dev.off()
}

```


#Function to create survival plots for each type of metacluster
```{r}
plot.OS <- function(vars, data.in, cutpoints){
    cut.off.value <- cutpoints[names(cutpoints) == vars]
    cat.var <- as.factor(ifelse(data.in[,vars] >  cut.off.value, "high","low"))
    data.in <- cbind(data.in, cat.var)
    fit <- survfit(Surv(timelastfu, finalstatus) ~ cat.var, data = data.in)
    logrank = surv_pvalue(fit, data.in)
    pval=logrank$pval.txt
    p <- ggsurvplot(fit, data = data.in, risk.table = TRUE, p.val=TRUE, conf.int = TRUE,
                    font.legend = c(18, "plain"), font.tickslab = c(16, "plain"),
                    font.x = c(18, "plain"), font.y = c(18, "plain"),
                    tables.y.text = FALSE,  break.time.by=30, xlim = c(0,120),
                    legend.labs=c(paste0(vars, "+"),paste0(vars, "-")),
                    palette = c("#ff5a36", "#4d4dff")) + 
          guides(colour = guide_legend(nrow = 2))
    p <- p + ylab("OS probability") + xlab("Time (days)")
    p$plot <- p$plot + theme(legend.key.width = unit(3, "line"), plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"), legend.key.height = unit(1.5, "line"))
    p$plot <- p$plot + ggplot2::annotate("text", x = 90, y = 0.85, label =pval, size = 5)
    p$table <- p$table + theme(plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"))
    plots <- list(plot=p$plot, table=p$table)
    return(plots)
}


plot.PFS <- function(vars, data.in, cutpoints){
    cut.off.value <- cutpoints[names(cutpoints) == vars]
    cat.var <- as.factor(ifelse(data.in[,vars] >  cut.off.value, "high","low"))
    data.in <- cbind(data.in, cat.var)
    fit <- survfit(Surv(daystoprogression, progression) ~ cat.var, data = data.in)
    logrank = surv_pvalue(fit, data.in)
    pval=logrank$pval.txt
    p <- ggsurvplot(fit, data = data.in, risk.table = TRUE, p.val=TRUE, conf.int = TRUE,
                    font.legend = c(18, "plain"), font.tickslab = c(16, "plain"),
                    font.x = c(18, "plain"), font.y = c(18, "plain"),
                    tables.y.text = FALSE,  break.time.by=15, xlim = c(0,60),
                    legend.labs=c(paste0(vars, "+"),paste0(vars, "-")),
                    palette = c("#ff5a36", "#4d4dff")) + 
          guides(colour = guide_legend(nrow = 2))
    p <- p + ylab("PFS probability") + xlab("Time (days)")
    p$plot <- p$plot + theme(legend.key.width = unit(3, "line"), plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"), legend.key.height = unit(1.5, "line"))
    p$plot <- p$plot + ggplot2::annotate("text", x = 40, y = 0.85, label =pval, size = 5)
    p$table <- p$table + theme(plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"))
    plots <- list(plot=p$plot, table=p$table)
    return(plots)
}


#Indicate the variables for which is going to be calculated the cox model
Cox.regresion.variables2 <- function(data, variables, formula='Surv(timelastfu, finalstatus)~'){

   univ_formulas <- sapply(variables,
                           function(x) as.formula(paste(formula, x)))

   univ_models <- lapply( univ_formulas, function(x){coxph(x, data = data, method = "exact")})

   univ_results <- lapply(univ_models,
                          function(x){
                             x <- summary(x)
                             positive <- x$nevent
                             prop.positive <- paste0(round((x$nevent) * 100 /x$n,0),"%")
                             #p.value <- signif(x$wald["pvalue"], digits=2) #p.value for the model
                             wald.test <- signif(x$wald["test"], digits=2)
                             lastrow <- nrow(x$coef) #When using several covariables in the same regression, take the last one used
                             p.value <- signif(x$coef[lastrow, "Pr(>|z|)"], digits=2) #p.value for last covariable
                             beta <- signif(x$coef[lastrow, 1], digits=2);#coeficient beta
                             HR <- signif(x$coef[lastrow, 2], digits=2);#exp(beta)
                             HR.confint.lower <- signif(x$conf.int[lastrow,"lower .95"], 2)
                             HR.confint.upper <- signif(x$conf.int[lastrow,"upper .95"],2)
                             res <- c(HR, p.value, HR.confint.lower,HR.confint.upper)
                             names(res)<-c("HR", "p.value", "HR.confint.lower", "HR.confint.upper")
                             return(res)
                          })
   res <- t(as.data.frame(univ_results, check.names = FALSE))
   result <- as.data.frame(res)
   return(result)
}
```






```{r}
out.put.folder <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/02_MetaClusters_November/Survival/"


proportions.by.patient.m <- proportions.c %>% select(-Molecular.profile2, -Proportion, -therapy_sequence) %>%
                              pivot_wider(names_from = SubCelltype, values_from = Rel.Prop)

clusters.order <- c("Basal","EMT","Proliferative","Immunogenic")


proportions.rel.surv <- merge(proportions.by.patient.m, clinical.data, by.x="patient",
                                     by.y="TnumberAVL")

cut.points.rel <- sapply(clusters.order, function(x){
        cutpoint <- median(proportions.rel.surv[,x])
        return(cutpoint)
})


suffix <- c("PDS","IDS")
for (i in suffix){
  data.surv <- proportions.rel.surv[proportions.rel.surv$therapy_sequence == i,]
  
  surv.tables.os <- lapply(sel.cols, plot.OS, data.surv, cut.points.rel)
  
  #Overall survival analysis
  fig.OS <- lapply(1:length(surv.tables.os) , function(x){surv.tables.os[[x]]$plot})
  tables.OS <- lapply(1:length(surv.tables.os) , function(x){surv.tables.os[[x]]$table})
  
  ggsave(file=paste0(out.put.folder, "KM_OS_metacluster", s, "_", i, ".svg"),
           arrangeGrob(grobs = fig.OS, ncol = 2), width = 21, height = 18, units = "cm")
  ggsave(file=paste0(out.put.folder, "KM_OS_metacluster", s, "_", i, "-table.svg"),
         arrangeGrob(grobs = tables.OS, ncol = 2), width = 21, height = 7, units = "cm")
  
  ##PFS survival analysis
  surv.tables.pfs <- lapply(sel.cols, plot.PFS, data.surv, cut.points.rel)
  fig.pfs <- lapply(1:length(surv.tables.pfs) , function(x){surv.tables.pfs[[x]]$plot})
  tables.pfs <- lapply(1:length(surv.tables.pfs) , function(x){surv.tables.pfs[[x]]$table})

  ggsave(file=paste0(out.put.folder, "KM_PFS_metacluster", s, "_", i, ".svg"),
           arrangeGrob(grobs = fig.pfs, ncol = 2), width = 21, height = 18, units = "cm")
  ggsave(file=paste0(out.put.folder, "KM_PFS_metacluster", s, "_", i, "-table.svg"),
         arrangeGrob(grobs = tables.pfs, ncol = 2), width = 21, height = 7, units = "cm")
}
```




#Calculating individual cox hazzard ratios per cell.type and including the Molecular profile in each
```{r}
tm <- forest_theme(core=list(bg_params=list(fill = c("grey90", "#eaf8f5", "white"))))


molecular.profile.set <- molecular.profiles[,-3]
proportions.rel.surv2 <- merge(proportions.rel.surv, molecular.profile.set, by="patient")

formulas <- c('Surv(timelastfu, finalstatus)~ Molecular.profile2 + refage2  + figo_stage2 +',
              'Surv(daystoprogression, progression)~ Molecular.profile2 + refage2  + figo_stage2 +')

formula.suffix <- c('OS_mprofile_age_figo', 'PFS_mprofile_age_figo')

clusters.order <- c("Basal","EMT","Proliferative","Immunogenic")


therapies <- c("IDS","PDS")


for (index in 1:length(formulas)){
  #Calculating cox hazards in in two groups therapy sequence
  cox.therpies <- lapply(therapies, function(x){
          dat <- proportions.rel.surv2[(proportions.rel.surv2$therapy_sequence %in% x),]
          t.rows <- which(proportions.rel.surv2$therapy_sequence %in% x)

          #If proportion of cell.type is equal or above the median in the sampled population, then is equal to 1
          for (j in clusters.order){
            dat[,j] <- ntile(proportions.rel.surv2[,j], 4)[t.rows]
          }
          #Calculating cox models
          Cox.metacluster <- Cox.regresion.variables2(dat, clusters.order, formula=formulas[index])
          Cox.metacluster$Treatment <- x
          Cox.metacluster$Cell.type <- rownames(Cox.metacluster)
          Cox.metacluster$p.value <- signif(p.adjust(Cox.metacluster$p.value, method ="BH"),2)
          return(Cox.metacluster)
  })
  
  
  #Reordering the cox results in an specific way necessary for the package forestplot
  spacer <- paste(rep(" ", 15), collapse = " ") #For the forest plot table
  cox.groups <- rbind(cox.therpies[[2]], cox.therpies[[1]])
  df.forest <- NULL
  for (i in unique(cox.groups$Cell.type)){
        row.class <- data.frame(Cell.type =i, HR="", space=spacer, p.value=NA, HR.confint.lower=NA,
                                HR.confint.upper=NA, HR2=NA)
        set.df <- cox.groups[cox.groups$Cell.type %in% i,]
        set.df$space <- spacer
        set.df$HR2 <- set.df$HR
        set.df.reordered <- set.df %>% select(Treatment, HR, space, p.value, HR.confint.lower, HR.confint.upper, HR2)
        colnames(set.df.reordered) <- colnames(row.class)
        row.celltype <- rbind(row.class, set.df.reordered)
        df.forest <- rbind(df.forest, row.celltype)
  }
  colnames(df.forest)[3] <- ""
  colnames(df.forest)[2] <- "HR"
  
  #Adding * to pvalues lower than 0.05
  low.pvals <- which(df.forest$p.value < 0.05)
  tiny.pvals <- which(df.forest$p.value < 0.001)
  df.forest$p.value[low.pvals] <- paste0(df.forest$p.value[low.pvals],"*")
  df.forest$p.value[tiny.pvals] <- "<0.001**"
  df.forest$p.value[which(is.na(df.forest$p.value))] = ""
  colnames(df.forest)[4] <- "FDR"
  
  #indent the subgroup if there is a number in the HR2 column
  df.forest$Cell.type <- ifelse(is.na(df.forest$HR2), 
                                df.forest$Cell.type,
                                #unlist(sapply(df.forest$Cell.type, function(x){cat("\t",x)})))
                                paste0("", df.forest$Cell.type))
  
  
  p <- forest(df.forest[,c(1:4)],
              est = df.forest$HR2,
              lower = df.forest$HR.confint.lower, 
              upper = df.forest$HR.confint.upper,
              sizes = 0.5,
              ci_column = 3,
              ref_line = 1,
              xlim = c(0, 1.5),
              arrow_lab = c("Lower risk", "Higher risk"),
              ticks_at = c(0.5, 1, 1.5),
              theme = tm)
  print(p)
  
  ggplot2::ggsave(filename = paste0(out.put.folder, "Forest_plot_", formula.suffix[index], ".svg"), plot = p,
                   width = 11, height = 11, units = "cm", device = svglite::svglite)

  ggplot2::ggsave(filename = paste0(out.put.folder, "Forest_plot_", formula.suffix[index], ".png"), plot = p,
                   width = 11, height = 11, units = "cm")
}
```
