---
title: "MHCII expression in RCNs"
output: html_notebook
---


```{r}
library(ggplot2)
library(gridExtra)
library("dplyr")
library(ComplexHeatmap)
library(tidyr)
library(paletteer)
library(vegan)
library(circlize)

source("D:/users/fperez/NKI_TMAs_AF/devNKI-scripts/5_Spatial-interactions/RCN_inspection_functions.R")
```




#Reading clinical info
```{r}
rcn.folder <- "Analysis_results/03_RCN/"
input.prefix <- "Cluster_ids_metaclustersmerged_"
project.folder <- "D:/users/fperez/NKI_TMAs_AF/"
output.napari <- "RCN_Napary2023_2"
out.folder.name <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/03_RCN_2024/"
method.rcn <- "radious46_k30"


cycif2samples <- read.table(file="D:/users/fperez/NKI_TMAs_AF/devNKI-scripts/3_Cell-typeClassification/tCicyf_patients-cores_pass.csv", sep=",", header = TRUE)
clinical.info <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Clinical_data/T-CycIF.updated.csv", sep=",",
                            header = TRUE)

clin.tcycif <- merge(cycif2samples, clinical.info, by.x="patient", by.y="TnumberAVL")

#Loading Molecular profiles
M.profiles <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Tables/Molecular_profiles_patients_20231208.csv",
             sep=",", header = TRUE)

out.folder.name <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/05_ML_validations/"

```


#Merging M.profile info with tCycif core ID
```{r}
M.profiles.clin  <- merge(M.profiles, cycif2samples, by="patient")
M.profiles.clin$imageid <- paste(M.profiles.clin$cycif.slide, M.profiles.clin$cycif.core.id, sep="_")
M.profiles.clin <- M.profiles.clin %>% select(patient,Molecular.profile2,therapy_sequence,imageid)
M.profiles.clin
```


#Reading input RCN table file
```{r}
#Reading RCN id per cell
RCN.clusters <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Analysis_results/Spatial_analysis/Cluster_ids_radius46_k40_metaclusters_20240113.csv", header = TRUE, sep=",")

#Next lines to rename the RNC id according to the abundance. So, RCN=1 would be the most abundant 
RCN.order <- order(table(RCN.clusters[,2]), decreasing = TRUE)
RCN.order.values <- sapply(1:length(RCN.order), function(x){which(RCN.order == x)})
RCN.clusters[,2] <- RCN.order.values[c(RCN.clusters[,2] + 1)]

#Reading the merged RCNs ids
RCN.annotations <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Analysis_results/03_RCN_2024_metaclusters/Merged_RCNs.csv",
                              sep=",",header=TRUE)

#Adding an RCNid number of the merged RCNs
RCN.c.ids <- paste0("RCN", sprintf('%0.2d',1:length(unique(RCN.annotations$RCN_name))))
names(RCN.c.ids) <- unique(RCN.annotations$RCN_name)

RCN.c.ids.all <- sapply(RCN.annotations$RCN_name, function(x){
      RCN.c.ids[which(x == names(RCN.c.ids))]
})

RCN.annotations$RCN.id.collapsed <- RCN.c.ids.all
RCN.annotations$RCN_id <- as.integer(gsub("RCN","",RCN.annotations$RCN_id))


#Next lines to rename the RNC id according to the abundance. So, RCN=1 would be the most abundant 
RCN.order <- order(table(RCN.annotations[RCN.clusters$spatial_kmeans,"RCN.id.collapsed"]), decreasing = TRUE)
new.rcn.order <- paste0("RCN", sprintf('%0.2d',1:length(unique(RCN.annotations$RCN.id.collapsed))))
names(new.rcn.order) <- unique(RCN.annotations$RCN.id.collapsed)[RCN.order]

RCN.annotations$RCN.id.collapsed.ordered <- new.rcn.order[RCN.annotations$RCN.id.collapsed]
RCN.annotations
```



##Re-naming the RCN accordint to the RNC group they belong
```{r}
#New names
new.vals <- c("RCN01", "RCN02", "RCN03","RCN04", "RCN05", "RCN06","RCN07", "RCN08", "RCN09","RCN10",
          "RCN11", "RCN12", "RCN13", "RCN14")

#Original names, with the order according to the groups
order <- c("RCN01", "RCN03", "RCN11", "RCN02", "RCN05", "RCN06", "RCN09", "RCN12", "RCN04", "RCN07",
               "RCN08", "RCN10", "RCN13", "RCN14")

names(new.vals) <- order
RCN.annotations$RCN.id.collapsed.ordered <- new.vals[RCN.annotations$RCN.id.collapsed.ordered]
RCN.annotations
```


#Reading big dataset with signal intensity per cell.type
```{r}
cells <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Tables/All_cells_subtype-info_20231129.csv",
                    header = TRUE, sep=",")
```

#Adding to cell table the RCNid
```{r}
cells.aux <- cells

#Order of RCN for plotting
RNC.order <- c("RCN01", "RCN02", "RCN03", "RCN04","RCN05","RCN06", "RCN07","RCN08", "RCN09", "RCN10",
                "RCN11", "RCN12", "RCN13", "RCN14")


#Adding the merged RCNid to the cells
cells.aux$rcn_id <- RCN.annotations[RCN.clusters$spatial_kmeans,"RCN.id.collapsed.ordered"]
cells.aux$rcn_id <- factor(cells.aux$rcn_id, levels = RNC.order)
```

#Selecting only cancer cells for analysis and from cancer rich RCNs
```{r}
#Selecting only cancer cells
cancers <- cells.aux %>% filter(GlobalCellType == "Cancer",
                                rcn_id %in% c("RCN04","RCN05","RCN06","RCN07","RCN08","RCN09","RCN10"))

#Merging cells ids with patient.id 
cancers.pat <- merge(cancers, cycif2samples, by=c("cycif.slide","cycif.core.id"))

#Merging cells ids with clinical info
cancers.clin <- merge(cancers.pat, M.profiles.clin, by="patient")

cancers.clin$Molecular.profile2 <- factor(cancers.clin$Molecular.profile2,
                                          levels=c("BRCAmut/met","HRD","HRP","CCNE1amp","Other"))

cancers.clin$therapy_sequence2 <- ifelse(cancers.clin$therapy_sequence == "NACT", "IDS", "PDS")
```



#Violing plots for MHCII expression per RCN in PDS, BRCAloss, cancer cells
```{r}
rnc_mycolors <- c("#FFE4C4", "tan", "#FFE4C4", "#ffc1ff", "#b400b4", "#f7f0f7","#ffc1ff","#ff4eff",
              "#FFE4C4","#ff3333","#ff6e4a", "#00ffff","#8a2be2","#00d800")

mycolors <- c("#ffc1ff", "#b400b4", "#f7f0f7","#ffc1ff","#ff4eff",
              "#FFE4C4","#ff3333","#ff6e4a", "#00ffff","#8a2be2","#00d800")


for (t in c("IDS", "PDS")){
cancers.pds <- cancers.clin %>% filter(therapy_sequence2 == {t},
                                       !is.na(therapy_sequence2))

df <- cancers.pds

#P.value that RNC05 expression is different from RCN04
pval.1 <- wilcox.test(df[df$rcn_id == "RCN05","MHCII"],
         df[df$rcn_id == "RCN08","MHCII"])
print(pval.1$p.value)

#P.value that RNC05 expression is different from RCN04
pval.2 <- wilcox.test(df[df$rcn_id == "RCN10","MHCII"],
         cancers[df$rcn_id == "RCN09","MHCII"])


p <- ggplot(df, aes(x=rcn_id, y=log2(MHCII))) +
        geom_violin(aes(fill=rcn_id)) + geom_boxplot(fill=NA, width=0.1, outlier.shape = NA) +
        stat_summary(geom = "errorbar", fun.y = "median", aes(ymax = ..y.., ymin = ..y..), width=0.6, size = 0.8) + 
        theme_bw() + xlab("RCNid") + ylab("log2(MHCII) in cancer cells") +
        theme(axis.text.x = element_text(angle = 45, vjust = 1.1, hjust=1, size=rel(1.2)),
              axis.text.y = element_text(size=rel(1.2)),
              axis.title.y = element_text(size=rel(1.2)),
              axis.title.x = element_blank(),
              legend.text = element_text(size=rel(1.2))) +
        scale_fill_manual(values = mycolors) + ylim(9.5,17)
print(p)
 ggsave(p, filename = paste0(out.folder.name,"/MHCII_expression_RNCs_",t,".svg"),
        height = 10, width = 16, units = "cm")
}
```


###Calculating median expression of MHCII per relevant RCN per molecular profiles
```{r}
#For complex heatmaps
col_fun = colorRamp2(c(0, 0.35), c("white", "#9a0000"))

signal.dat <- cancers.clin %>% filter(!grepl("NACT",therapy_sequence),
                                       !is.na(therapy_sequence),
                                       Molecular.profile2 != "Other") %>%
                                       mutate(MHCII.log2 = log2(MHCII)) %>% 
                                       select(Molecular.profile2, rcn_id, MHCII.log2)


#Calculating median value of MHCII by molecular profiles in PDS
cancers.mhcii.med <- cancers.clin %>% filter(!grepl("NACT",therapy_sequence),
                                       !is.na(therapy_sequence),
                                       Molecular.profile2 != "Other") %>%
                                       mutate(MHCII.log2 = log2(MHCII)) %>% 
                                       mutate(MHCII.scaled = 
                                              (MHCII.log2 - min(MHCII.log2))/(max(MHCII.log2) - min(MHCII.log2))) %>% 
                                       group_by(rcn_id, Molecular.profile2) %>% 
                                       summarise(med.MHCII=median(MHCII.scaled))

#Changing data to matrix 
mt.med.exp <- cancers.mhcii.med %>% pivot_wider(names_from = rcn_id , values_from = med.MHCII)
mt.med.exp <- as.data.frame(mt.med.exp)
mt.med.exp[is.na(mt.med.exp)] <- 0
mt.med.exp <- t(as.matrix(mt.med.exp[,-1]))


#Preparing data for creating complex heatmap
M.profile.order <- c("BRCAloss","HRD","HRP","CCNE1amp")
colnames(mt.med.exp) <- M.profile.order

#Calculating p.values
krustal.df <- NULL
kruskal.pvals <- lapply(unique(cancers.mhcii.med$rcn_id), function(x){
              df <- signal.dat[which(signal.dat$rcn_id == x),]
              k.res <- kruskal.test(MHCII.log2 ~ Molecular.profile2, data=df)
              res <- data.frame(RNCid=x, pval=k.res$p.value, y.val = 0.8)
              return(res)
})
for (i in 1:length(kruskal.pvals)){krustal.df <- rbind(krustal.df, kruskal.pvals[[i]])}
print(krustal.df)

is_sig = krustal.df$pval < 0.05
pch = rep("P<0.05", length(is_sig))
pch[!is_sig] = NA
  
ha = rowAnnotation(Significance=pch, col=list(Significance=c("P<0.05"="cyan4")), na_col="grey95",
                   gp = gpar(col = "white", lwd = 0.5))


hmap <- Heatmap(mt.med.exp, cluster_columns = FALSE, cluster_rows = FALSE, row_names_side = "left",
        name="Median\nexpression\n(Min.max)", rect_gp = gpar(col = "white", lwd = 0.5), col = col_fun,
        column_title_gp = gpar(fontsize = 16), row_names_gp = gpar(fontsize = 12),
        column_names_gp = gpar(fontsize = 14),  column_names_rot = 45, right_annotation = ha)

pdf(paste0(out.folder.name, "Heatmap_MHCII_cancer_Mprofiles.pdf"), width=3, height=3)
draw(hmap)
dev.off()
draw(hmap)
```