---
title: "RNA deconvolution validation"
output: html_notebook
---

```{r}
library(ggplot2)
library(tidyverse)
library(survival)
library(survminer)
library(reshape2)
library(gridExtra)
library(grid)
library(ComplexHeatmap)
```


#Reading input RNA deconvolution files
```{r}
input.folder <- "D:/users/fperez/NKI_TMAs_AF/TCGA_PRISM_RNAdeconvolution/"

g <- read.table(file=paste0(input.folder,"G.tsv"))
z <- read.table(file=paste0(input.folder,"Z.tsv"), header = TRUE)
w <- read.table(file=paste0(input.folder,"W.tsv"), header = TRUE)
```

#Reading HRD status shared in Perez-Villatoro et al 2022 and clinical data from TCGA
```{r}
clin.dat <- read.csv(file = paste0(input.folder,"TCGA-CDR-SupplementalTableS1_stage-survival.csv"),
                      header = TRUE, row.names = 1)

out.put.folder <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/05_ML_validations/TCGA_RNA/"
dir.create(out.put.folder)

HRD.tab <- read.table(file=paste0(input.folder,"Categories_Konstantinopoulos_HR_HGSC_germline.csv"),
                      sep=",", header=TRUE)
```
#Stratifiyin samples acording to tumoral molecular profiles
```{r}
HRD.tab$Molecular.profile <- "HRP"
HRD.tab$Molecular.profile[HRD.tab$categories == "CCNE1 amplification"] <- "CCNE1amp"
HRD.tab$Molecular.profile[(HRD.tab$HRDsum >= 54 & !grepl("BRCA", HRD.tab$categories))] <- "HRD"
HRD.tab$Molecular.profile[grep("BRCA", HRD.tab$categories)] <- "BRCAloss"
HRD.tab.cat <- HRD.tab[,c("Sample","Molecular.profile")]
```




#Structuring clinical data
```{r}
clin.dat <- clin.dat[clin.dat$type == "OV",]
clin.dat$PFI.time <- as.numeric(clin.dat$PFI.time)/30.4
clin.dat$OS.time <- as.numeric(clin.dat$OS.time)/30.4
clin.dat$OS <- as.numeric(clin.dat$OS)
clin.dat$PFI <- as.numeric(clin.dat$PFI)
clin.dat$age <- as.numeric(clin.dat$age_at_initial_pathologic_diagnosis)

clin.dat$age.bin <- NA
clin.dat$age.bin[clin.dat$age < 40] = 0
clin.dat$age.bin[clin.dat$age >= 40 & clin.dat$age < 50] = 1
clin.dat$age.bin[clin.dat$age >= 50 & clin.dat$age < 60] = 2
clin.dat$age.bin[clin.dat$age >= 60 & clin.dat$age < 70] = 3
clin.dat$age.bin[clin.dat$age >= 70 & clin.dat$age < 80] = 4
clin.dat$age.bin[clin.dat$age >= 80] = 5

#Selecting only high patients with high grade tumors 
clin.dat <- clin.dat[clin.dat$histological_grade %in% c("G3","G4"),]

#Ignoring low clinical stages
clin.dat <- clin.dat[!clin.dat$histological_grade %in% c("Stage IA","Stage IB", "Stage IC",
                                                         "Stage IIA","Stage IIB","Stage IIC"),]
```

#Visualizing RNA deconvoluted expression
```{r}
head(z)
```


#Normalizing data according to PRISM developers
```{r}
w.dat <- as.matrix(w[,-1])
z.dat <- as.matrix(z[,-c(1,2)])

deconv <- t( t(z.dat) / c(t(g * w.dat )))
deconv <- as.data.frame(deconv)

deconv$symbol <- z$symbol

deconv <- deconv %>% select(symbol, everything())
```


#Selecting HLA genes and organizing data table
```{r}
hla.deconv <- deconv
#hla.deconv <- deconv %>% filter(grepl("HLA",symbol ))

#Selecting the expression columns that comes from Epithelial Ovarian Cancer
hla.eoc <- hla.deconv[,c(1,grep("EOC", colnames(hla.deconv)))]
hla.genes <- hla.eoc$symbol
hla.eoc <- as.data.frame(t(hla.eoc[,-1]))
colnames(hla.eoc) <- gsub("-","_", hla.genes)

#Selecting those samples taken in primary setting
setting <-sapply(row.names(hla.eoc), function(x){strsplit(x, "[.]")[[1]][4]})
hla.eoc.p  <- hla.eoc[grepl("01", setting),]

#Selecting those samples taken in recurrence setting
hla.eoc.r  <- hla.eoc[grep("02", setting),]


#Making shorter names from TCGA.04.1348.01A.01R.1565.13..EOC to TCGA.04.1348
names.short.p <- sapply(row.names(hla.eoc.p), function(x){
                  paste(strsplit(x, "[.]")[[1]][1:3], collapse = "-")})

row.names(hla.eoc.p) <- names.short.p
hla.eoc.p
```


#Merging RNA expression with clinical data and HRD status
```{r}
hla.eoc.clin <- merge(hla.eoc.p, clin.dat, by.x="row.names", by.y="bcr_patient_barcode")
colnames(hla.eoc.clin)[1] <- "Sample"
hla.eoc.hr.clin <- merge(hla.eoc.clin, HRD.tab.cat, by="Sample")
dim(hla.eoc.hr.clin)
```

#Stratiifying samples according to MHCII expression and performing survival analysis
```{r}
formulas.names <- c("OS","PFI")
formulas.labs <- c("OS probability","PFI probability")
xlim.max.val <- c(150,120)
dat <- hla.eoc.hr.clin

formulas.cox <- c('Surv(OS.time, OS) ~ age.bin + Molecular.profile',
              'Surv(PFI.time, PFI)~ age.bin + Molecular.profile')

formulas.kapplan <- c('Surv(OS.time, OS) ~  Expression',
              'Surv(PFI.time, PFI)~ Expression')

genes.to.test <- c("MKI67","HLA_DPB1","HLA_A","KRT7")
for (gene in genes.to.test){
  for (i in 1:length(formulas.kapplan)){
    #For cox model data is stratified by tiles 
    dat$Expression <- ntile(dat[,gene],3)
    colnames(dat)[colnames(dat) == "Expression"] = paste0(gene,"exp")
  
    model <- coxph(as.formula(paste0(formulas.cox[i]," + ", gene,"exp")), data=dat)
    p1 <- ggforest(model, data = dat,  main=formulas.labs[i])
    print(p1)
    ggsave(p1, file=paste0(out.put.folder, "Cox_",gene,"_age_",formulas.names[i],".png"),
           width = 15, height = 10, units = "cm")
    
    #For kapplan data is stratified by median expression 
    dat$Expression <- ifelse(dat[,gene] > median(dat[,gene]), "High","Low")
    
    fit <- survfit(as.formula(formulas.kapplan[i]), data = dat)
    logrank = surv_pvalue(fit, dat)
    pval=logrank$pval.txt
    p <- ggsurvplot(fit, data = dat, risk.table = TRUE, p.val=TRUE, conf.int = TRUE,
                    palette = c("#ff5a36","#4d4dff"),
                   legend.labs=c(paste0(gene,".High"), paste0(gene,".Low")),
                   tables.y.text = FALSE, legend.title="",
                   xlim=c(0,xlim.max.val[i]), break.time.by=30, ylab=formulas.labs[i])
    p$plot <- p$plot + theme(legend.key.width = unit(3, "line"),
                             plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"),
                             legend.key.height = unit(1.5, "line"))
    p$plot <- p$plot + ggplot2::annotate("text", x = 90, y = 0.85, label =pval, size = 5)
    p$table <- p$table + theme(plot.margin = margin(t = 0,  r = 2, b = 5, l = 2, unit = "cm"))
    plot.to.save <- list(p$plot, p$table)
    ggsave(file=paste0(out.put.folder,"Kapplan_",gene,"_",formulas.names[i],".png"),
            arrangeGrob(grobs = plot.to.save, ncol = 1),
          width = 13, height = 16, units = "cm")
  }
}
```
#Survival only with HRDs or HRPs
```{r}
classes <- list(c("BRCAloss","HRD"), 
                c("CCNE1amp","HRP"))

profile.names <- c("HRDs","HRPs")

formulas.names <- c("OS","PFI")
formulas.labs <- c("OS probability","PFI probability")
xlim.max.val <- c(150,120)

formulas.cox <- c('Surv(OS.time, OS) ~  HLA_DPB1.exp + age.bin',
              'Surv(PFI.time, PFI)~ HLA_DPB1.exp + age.bin')

formulas.kapplan <- c('Surv(OS.time, OS) ~  HLA_DPB1.exp',
              'Surv(PFI.time, PFI)~ HLA_DPB1.exp')

for (index in 1:length(classes)){
  dat <- hla.eoc.hr.clin[hla.eoc.hr.clin$Molecular.profile %in% classes[[index]],]
  
  for (i in 1:length(formulas.kapplan)){
    #For cox model data is stratified by tiles 
    dat$HLA_DPB1.exp <- ntile(dat$HLA_DPB1,3)
  
    model <- coxph(as.formula(formulas.cox[i]), data=dat)
    p1 <- ggforest(model, data = dat,  main=paste0(formulas.labs[i]," ", profile.names[index]))
    print(p1)
    ggsave(p1, file=paste0(out.put.folder, "Cox_HLA-DPB1_age_",profile.names[index], "_", formulas.names[i],".png"),
           width = 15, height = 8, units = "cm")
    
    #For kapplan data is stratified by median expression 
    dat$HLA_DPB1.exp <- ifelse(dat$HLA_DPB1 > median(dat$HLA_DPB1), "HLA-DPB1.High","HLA-DPB1.Low")
    
    fit <- survfit(as.formula(formulas.kapplan[i]), data = dat)
    logrank = surv_pvalue(fit, dat)
    pval=logrank$pval.txt
    p <- ggsurvplot(fit, data = dat, risk.table = TRUE, p.val=TRUE,
                    conf.int = TRUE, palette = c("#ff5a36","#4d4dff"),
                   legend.labs=c("HLA-DPB1.High", "HLA-DPB1.Low"), tables.y.text = FALSE, legend.title="",
                   xlim=c(0,xlim.max.val[i]), break.time.by=30, ylab=formulas.labs[i])
    p$plot <- p$plot + theme(legend.key.width = unit(3, "line"),
                             plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"),
                             legend.key.height = unit(1.5, "line"))
    p$plot <- p$plot + ggplot2::annotate("text", x = 90, y = 0.85, label =pval, size = 5)
    p$table <- p$table + theme(plot.margin = margin(t = 0,  r = 2, b = 5, l = 2, unit = "cm"))
    plot.to.save <- list(p$plot, p$table)
      ggsave(file=paste0(out.put.folder,"Kapplan_HLA-DPB1_",profile.names[index], "_", formulas.names[i],".png"),
             arrangeGrob(grobs = plot.to.save, ncol = 1),
           width = 13, height = 16, units = "cm")
  }
}
```


#Calculating mean expression of MHCII and MHCI, and survival analysis of it
```{r}
MHCII.genes <- c("HLA-DPA1","HLA-DPA2","HLA-DPA3","HLA-DPB1","HLA-DPB2","HLA-DQA1","HLA-DQA2","HLA-DQB1",
                 "HLA-DQB2","HLA-DQB3","HLA-DRA","HLA-DRB1","HLA-DRB2","HLA-DRB3","HLA-DRB4","HLA-DRB5",
                 "HLA-DRB6","HLA-DRB7","HLA-DRB8","HLA-DRB9")
MHCII.genes <- gsub("-","_", MHCII.genes)

#MHCII.genes <- c("HLA_DPB1")

MHCI.genes <- c("HLA-A","HLA-B","HLA-C","HLA-E","HLA-F","HLA-G","HLA-H","HLA-J","HLA-K","HLA-L","HLA-N",
                "HLA-P","HLA-S","HLA-T","HLA-U","HLA-V","HLA-W","HLA-X","HLA-Y","HLA-Z")
MHCI.genes <- gsub("-","_", MHCI.genes)

MHCII.mean.expression <- hla.eoc.p %>% mutate(Sample=row.names(.)) %>%
                            select(Sample, any_of(MHCII.genes)) %>% melt() %>% 
                            group_by(Sample) %>% summarise(MHCII.mean.exp=median(value))

MHCI.mean.expression <- hla.eoc.p %>% mutate(Sample=row.names(.)) %>%
                            select(Sample, any_of(MHCI.genes)) %>% melt() %>% 
                            group_by(Sample) %>% summarise(MHCI.mean.exp=median(value))

mhc.mean.exp <- cbind(MHCII.mean.expression, MHCI.mean.expression[,-1])


#Merging with clinical data
mhc.mean.exp.clin <- merge(mhc.mean.exp, clin.dat, by.x="Sample", by.y="bcr_patient_barcode")
mhc.mean.exp.clin <- merge(mhc.mean.exp.clin, HRD.tab.cat, by="Sample")

#Stratifiying the mean expression in three groups
dat <- mhc.mean.exp.clin
dat$MHCII.exp <- ntile(dat$MHCII.mean.exp,3)

#Doing Cox models
formulas <- c('Surv(OS.time, OS) ~  MHCII.exp + age.bin + Molecular.profile',
              'Surv(PFI.time, PFI)~ MHCII.exp + age.bin + Molecular.profile')

model <- coxph(as.formula(formulas[1]), data=dat)
p1 <- ggforest(model, data = dat)
print(p1)
```


#Plotting expression of MHCII accoding to Molecular profile
```{r}
df <- hla.eoc.hr.clin
df$Molecular.profile <- factor(df$Molecular.profile, levels=c("BRCAloss","HRD","HRP","CCNE1amp"))

df$Molecular.profile2 <- ifelse(df$Molecular.profile == "HRP" | df$Molecular.profile == "CCNE1amp", "HRPs","HRDs")

p <- ggplot(df, aes(x=Molecular.profile, y=HLA_DPB1)) +
      geom_point(aes(col=Molecular.profile), position= position_jitter(width= .1), alpha = 0.5) +
      geom_boxplot(outlier.shape=NA, fill=NA, width=0.3) +
      geom_violin(fill=NA) +
      labs(title="TCGA cancer-cells deconvoluted expression", subtitle="N=252") +
      ylab("HLA_DPB1 deconv RNA expression") +
      theme_bw() + ylim(0,0.00065)
print(p)
ggsave(p, filename=paste0(out.put.folder,"Boxplot_HLA-DPB1_molecular_profiles.png"), width = 13,
                      height = 11, units = "cm")
```


###Selecting deconvoluted RNA expression of immune cells
```{r}

#Selecting the expression columns that comes from Epithelial Ovarian Cancer
immune.exp <- deconv[,c(1,grep("Immune", colnames(deconv)))]
immune.genes <- immune.exp$symbol
immune.exp <- as.data.frame(t(immune.exp[,-1]))
colnames(immune.exp) <- gsub("-","_", immune.genes)

#Selecting those samples taken in primary setting
setting <-sapply(row.names(immune.exp), function(x){strsplit(x, "[.]")[[1]][4]})
immune.exp.p  <- immune.exp[grepl("01", setting),]

#Selecting those samples taken in recurrence setting
immune.exp.r  <- immune.exp[grep("02", setting),]


#Making shorter names from TCGA.04.1348.01A.01R.1565.13..immune to TCGA.04.1348
names.short.p <- sapply(row.names(immune.exp.p), function(x){
                  paste(strsplit(x, "[.]")[[1]][1:3], collapse = "-")})

row.names(immune.exp.p) <- names.short.p
immune.exp.p
```


#Merging RNA expression with clinical data and HRD status
```{r}
immune.exp.clin <- merge(immune.exp.p, clin.dat, by.x="row.names", by.y="bcr_patient_barcode")
colnames(immune.exp.clin)[1] <- "Sample"
immune.exp.hr.clin <- merge(immune.exp.clin, HRD.tab.cat, by="Sample")
dim(immune.exp.hr.clin)
```


#Stratiifying samples according to MHCII expression and performing survival analysis
```{r}
total.genes.to.test <- c("HLA_DPB1","HLA_A", "CD74", "CXCL9", "CXCL10", "CXCL11",
                   "IDO1", "ADAMDEC1", "SLAMF7", "CCL5", "CD3D")

hrp.genes.to.test <- c("CRABP1", "CD2", "SLAMF7", "ADAMDEC1", "CD3E", "CD27", "CTLA4", 
                      "CD3D", "CXCL9", "CXCL10", "CXCL11", "IDO1")

genes.to.test <- unique(hrp.genes.to.test, total.genes.to.test)

out.put.folder2 <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/05_ML_validations/TCGA_RNA/Immune_deconv/"


formulas.names <- c("OS","PFI")
formulas.labs <- c("OS probability","PFI probability")
xlim.max.val <- c(150,120)
dat <- immune.exp.hr.clin

formulas.cox <- c('Surv(OS.time, OS) ~ age.bin + Molecular.profile',
              'Surv(PFI.time, PFI)~ age.bin + Molecular.profile')

formulas.kapplan <- c('Surv(OS.time, OS) ~  Expression',
              'Surv(PFI.time, PFI)~ Expression')

for (gene in genes.to.test){
  for (i in 1:length(formulas.kapplan)){
    #For cox model data is stratified by tiles 
    dat$Expression <- ntile(dat[,gene],3)
    colnames(dat)[colnames(dat) == "Expression"] = paste0(gene,"exp")
  
    model <- coxph(as.formula(paste0(formulas.cox[i]," + ", gene,"exp")), data=dat)
    p1 <- ggforest(model, data = dat,  main=formulas.labs[i])
    print(p1)
    ggsave(p1, file=paste0(out.put.folder2, "IMMUNE_Cox_",gene,"_age_",formulas.names[i],".png"),
           width = 15, height = 10, units = "cm")
    
    #For kapplan data is stratified by median expression 
    dat$Expression <- ifelse(dat[,gene] > median(dat[,gene]), "High","Low")
    
    fit <- survfit(as.formula(formulas.kapplan[i]), data = dat)
    logrank = surv_pvalue(fit, dat)
    pval=logrank$pval.txt
    p <- ggsurvplot(fit, data = dat, risk.table = TRUE, p.val=TRUE, conf.int = TRUE,
                    palette = c("#ff5a36","#4d4dff"),
                   legend.labs=c(paste0(gene,".High"), paste0(gene,".Low")),
                   tables.y.text = FALSE, legend.title="",
                   xlim=c(0,xlim.max.val[i]), break.time.by=30, ylab=formulas.labs[i])
    p$plot <- p$plot + theme(legend.key.width = unit(3, "line"),
                             plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"),
                             legend.key.height = unit(1.5, "line"))
    p$plot <- p$plot + ggplot2::annotate("text", x = 90, y = 0.85, label =pval, size = 5)
    p$table <- p$table + theme(plot.margin = margin(t = 0,  r = 2, b = 5, l = 2, unit = "cm"))
    plot.to.save <- list(p$plot, p$table)
    ggsave(file=paste0(out.put.folder2,"IMMUNE_Kapplan_",gene,"_",formulas.names[i],".png"),
            arrangeGrob(grobs = plot.to.save, ncol = 1),
          width = 13, height = 16, units = "cm")
  }
}
```




#Survival using immune expression for HRDs or HRPs
```{r}
classes <- list(c("BRCAloss","HRD"), 
                c("CCNE1amp","HRP"))

profile.names <- c("HRDs","HRPs")

formulas.names <- c("OS","PFI")
formulas.labs <- c("OS probability","PFI probability")
xlim.max.val <- c(150,120)

formulas.cox <- c('Surv(OS.time, OS) ~ age.bin + Molecular.profile',
              'Surv(PFI.time, PFI)~ Molecular.profile')

formulas.kapplan <- c('Surv(OS.time, OS) ~  Expression',
              'Surv(PFI.time, PFI)~ Expression')

dat2 <- immune.exp.hr.clin
dat2$HLA_DPB1 <- ntile(hla.eoc.hr.clin$HLA_DPB1,3)
dat2$MHCII.bin <- ntile(hla.eoc.hr.clin$HLA_DPB1,2)

dat2 <- dat2 %>% filter(HLA_DPB1 > 1)


for (index in 1:length(classes)){
  dat <- dat2[dat2$Molecular.profile %in% classes[[index]],]
  
  for (gene in genes.to.test){
  for (i in 1:length(formulas.kapplan)){
    #For cox model data is stratified by tiles 
    dat$Expression <- ntile(dat[,gene],3)
    colnames(dat)[colnames(dat) == "Expression"] = paste0(gene,"exp")
  
    model <- coxph(as.formula(paste0(formulas.cox[i]," + ", gene,"exp")), data=dat)
    p1 <- ggforest(model, data = dat,  main=formulas.labs[i])
    ggsave(p1, file=paste0(out.put.folder2, "IMMUNE_Cox_",gene,"_age_",profile.names[index],
                           "_", formulas.names[i],".png"),
           width = 15, height = 10, units = "cm")
    
    #For kapplan data is stratified by median expression 
    dat$Expression <- ifelse(dat[,gene] > median(dat[,gene]), "High","Low")
    
    fit <- survfit(as.formula(formulas.kapplan[i]), data = dat)
    logrank = surv_pvalue(fit, dat)
    pval=logrank$pval.txt
    p <- ggsurvplot(fit, data = dat, risk.table = TRUE, p.val=TRUE, conf.int = TRUE,
                    #palette = c("#ff5a36","#4d4dff"),
                   #legend.labs=c(paste0(gene,".High"), paste0(gene,".Low")),
                   tables.y.text = FALSE, legend.title="",
                   xlim=c(0,xlim.max.val[i]), break.time.by=30, ylab=formulas.labs[i])
    p$plot <- p$plot + theme(legend.key.width = unit(3, "line"),
                             plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"),
                             legend.key.height = unit(1.5, "line"))
    p$plot <- p$plot + ggplot2::annotate("text", x = 90, y = 0.85, label =pval, size = 5)
    p$table <- p$table + theme(plot.margin = margin(t = 0,  r = 2, b = 5, l = 2, unit = "cm"))
    p$plot <-  p$plot + guides(colour = guide_legend(nrow = 5))
    plot.to.save <- list(p$plot, p$table)
    ggsave(file=paste0(out.put.folder2,"IMMUNE_Kapplan_",gene,"_", profile.names[index], "_", formulas.names[i],".png"),
            arrangeGrob(grobs = plot.to.save, ncol = 1),
          width = 13, height = 21, units = "cm")
  }
 }
}
```




#####Exploring the relationship of MHCII expression in cancer cells and immune gen expression
```{r}
#Remove outlier function
out.rem <- function(x){
    iqr.x <- IQR(x)
    quantile3 <- quantile(x, 0.75)
    cut.val <- quantile3 + 1.5 * iqr.x
    x[which(x > cut.val)] <- cut.val
    return(x)
}

dat2 <- immune.exp.hr.clin
dat2$HLA.eoc <- ntile(hla.eoc.hr.clin$HLA_DPB1,3)

total.genes.to.test <- c("CD74", "CXCL9", "CXCL10", "CXCL11",
                   "IDO1", "ADAMDEC1", "SLAMF7", "CCL5", "CD3D")

hrp.genes.to.test <- c("CRABP1", "CD2", "SLAMF7", "ADAMDEC1", "CD3E", "CD27", "CTLA4", 
                      "CD3D", "CXCL9", "CXCL10", "CXCL11")

genes.to.test <- unique(hrp.genes.to.test, total.genes.to.test)

mat <- dat2 %>% select(any_of(c(genes.to.test))) %>% 
                mutate_all(., out.rem) %>% scale() %>% t()

ha = HeatmapAnnotation(
    MHCII.eoc = dat2$HLA.eoc,
    Profile = factor(immune.exp.hr.clin$Molecular.profile,levels=c("BRCAloss","HRD","HRP","CCNE1amp")),
    col = list(Profile = c("CCNE1amp" = "blue3", "HRP" = "cyan", "HRD" = "red", BRCAloss="gold")
    ),
    gp = gpar(col = "grey")
)


png(paste0(out.put.folder, "Heatmap_HLA-DPB1_Immune_genes_expresssion.png"), res=300, width=8, height=7, units = "in")
Heatmap(mat, top_annotation = ha, name="Z-score\nexpression")
dev.off()
```
#Boxplots for expression accross molecular profiles
```{r}
genes.exp.immune <- as.data.frame(t(mat))
df <- data.frame(Sample=dat2$Sample,  MHCII.eoc.strat=dat2$HLA.eoc,
                 Molecular.profile=factor(dat2$Molecular.profile, levels=c("BRCAloss","HRD","HRP","CCNE1amp")))
df <- cbind(df, genes.exp.immune)

#genes.exp.immune
colnames(df)[1] <- "Sample"
df <- df %>% mutate(Class=paste0(Molecular.profile,MHCII.eoc.strat)) %>% 
             select(-MHCII.eoc.strat) %>% 
             melt()

colnames(df)[4]<- "Gene"
colnames(df)[5]<- "Expression"

df$Molecular.profile <- factor(df$Molecular.profile, levels=c("BRCAloss","HRD","HRP","CCNE1amp"))

my_levels <- c("BRCAloss1","BRCAloss2","BRCAloss3","HRD1","HRD2","HRD3","HRP1","HRP2","HRP3",
              "CCNE1amp1","CCNE1amp2","CCNE1amp3")

df$Class <- factor(df$Class, levels=my_levels)

p <- ggplot(df, aes(x=Class, y=Expression, col=Molecular.profile)) +
        geom_boxplot() + xlab("") + theme_bw() +
        theme(axis.text.x = element_text(angle = 75, hjust = 1)) +
        facet_wrap(~Gene)
ggsave(p, filename=paste0(out.put.folder,"Boxplot_HLA-DPB1_Immune_genes_expresssion_Mprofiles.png"), width = 20,
                      height = 16, units = "cm")
```

#Cox model and Kapplan-Mair using sum of all immune gene expressions
```{r}
total.genes.to.test <- c("CD74", "CXCL9", "CXCL10", "CXCL11",
                   "IDO1", "ADAMDEC1", "SLAMF7", "CCL5", "CD3D")

hrp.genes.to.test <- c("CRABP1", "CD2", "SLAMF7", "ADAMDEC1", "CD3E", "CD27", "CTLA4", 
                      "CD3D", "CXCL9", "CXCL10", "CXCL11")

genes.to.test <- unique(hrp.genes.to.test, total.genes.to.test)

out.put.folder2 <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/05_ML_validations/TCGA_RNA/Immune_deconv/"


formulas.names <- c("OS","PFI")
formulas.labs <- c("OS probability","PFI probability")
xlim.max.val <- c(150,120)

dat <- immune.exp.hr.clin
dat <- dat %>% select(any_of(c("Sample", genes.to.test))) %>%
                melt() %>%
                group_by(Sample) %>% 
                summarise(Total.exp = median(value)) %>% 
                as.data.frame()

dat.clin <- merge(dat, clin.dat, by.x="Sample", by.y="bcr_patient_barcode")
colnames(immune.exp.clin)[1] <- "Sample"
dat.clin.hr.clin <- merge(dat.clin, HRD.tab.cat, by="Sample")

dat <- dat.clin.hr.clin
dat$HLA.eoc <- ntile(hla.eoc.hr.clin$HLA_DPB1,3)

dat <- dat %>% filter(HLA.eoc > 1)


formulas.cox <- c('Surv(OS.time, OS) ~ age.bin + Molecular.profile + Expression',
              'Surv(PFI.time, PFI)~ age.bin + Molecular.profile + Expression')

formulas.kapplan <- c('Surv(OS.time, OS) ~  Expression',
              'Surv(PFI.time, PFI)~ Expression')

for (i in 1:length(formulas.kapplan)){
    #For cox model data is stratified by tiles 
    dat$Expression <- ntile(dat$Total.exp,3)
    model <- coxph(as.formula(formulas.cox[i]), data=dat)
    p1 <- ggforest(model, data = dat,  main=formulas.labs[i])
    print(p1)
    ggsave(p1, file=paste0(out.put.folder2, "TOTAL_Cox_age_",formulas.names[i],".png"),
           width = 15, height = 10, units = "cm")
    
    #For kapplan data is stratified by median expression 
    dat$Expression <- ifelse(dat$Total.exp > median(dat$Total.exp), "High","Low")
    
    fit <- survfit(as.formula(formulas.kapplan[i]), data = dat)
    logrank = surv_pvalue(fit, dat)
    pval=logrank$pval.txt
    p <- ggsurvplot(fit, data = dat, risk.table = TRUE, p.val=TRUE, conf.int = TRUE,
                   palette = c("#ff5a36","#4d4dff"),
                   legend.labs=c("High", "Low"),
                   tables.y.text = FALSE, legend.title="",
                   xlim=c(0,xlim.max.val[i]), break.time.by=30, ylab=formulas.labs[i])
    p$plot <- p$plot + theme(legend.key.width = unit(3, "line"),
                             plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"),
                             legend.key.height = unit(1.5, "line"))
    p$plot <- p$plot + ggplot2::annotate("text", x = 90, y = 0.85, label =pval, size = 5)
    p$table <- p$table + theme(plot.margin = margin(t = 0,  r = 2, b = 5, l = 2, unit = "cm"))
    plot.to.save <- list(p$plot, p$table)
    ggsave(file=paste0(out.put.folder2,"TOTAL_Kapplan_",formulas.names[i],".png"),
            arrangeGrob(grobs = plot.to.save, ncol = 1),
          width = 18, height = 16, units = "cm")
}
```

