---
title: "R Notebook"
output: html_notebook
---


```{r}
library(ggplot2)
library(tidyverse)
library(survival)
library(survminer)
library(gridExtra)
```
#Defining input variables
```{r}
clinical.data <- read.table(file="L:/ltdk_farkkila/Projects/4_CellCycle/Samples_clinical_data.csv",
                            sep=",", header = TRUE)

input.annotation.folder <- "P:/h345/afarkkilab/Projects/NKI/Whole_slide_validation/tribus_analysis/output_data/"
annotation.suffix <- "_annotated.csv"

#Listing all input files
input.annotation.files <- list.files(input.annotation.folder, pattern = annotation.suffix, full.names =TRUE)


#MHCII gating values
gates.mhcii <- data.frame(sample=c("S057_iOme","S081_iOme","S073_iOme1","S072_iOme","S118_iOme", "S083_iOme1",
                                   "S098_iOme","S112_iOme","S091_iOme1","S100_iOme","S107_iOme",
                                   "S110_iTubR","S113_iOme","S121_iOme","S123_iOme","S130_iOme","S131_iOme",
                                   "S188_iOme", "S197_iOme","S268_iOme","S050_iAdn","S065_iOme"),
                         gate.val=c(3.0,3.2,3.4,3.2,2.8,3.2,2.6,2.7,
                                    2.8,3.1,2.9,2.9,3.0,3.0,3.3,2.9,
                                    3.1,3.15,3.1,3.1,2.9,3.0))
input.annotation.files
gates.mhcii2 <- gates.mhcii
```

##Reading spatial data
```{r}
scimap.output.folder <- "P:/h345/afarkkilab/Projects/NKI/Whole_slide_validation/scimap_analysis/output_data/"
scimap.file <- "Spatial-cell-proportions_radious30px_new_fuction.csv"

neighboors <- read.table(file=paste0(scimap.output.folder, scimap.file), sep="," , header = TRUE)

```


```{r}
unique(strsplit(neighboors$X,"_")[[1]][1])
```



```{r}
dim(neighboors)
```


#Reading input annotation files
```{r}
cells <- NULL
for (f in input.annotation.files){
  file.name <- basename(f)
  sampleid <- paste(strsplit(file.name, "_")[[1]][1:2], collapse = "_")
  aux <- read.table(file=f, header = TRUE, sep=",")
  aux <- aux[,-1]
  aux[,1] <- sampleid
  colnames(aux)[1] <- "sampleID"
  if(any(colnames(aux) == "Cancer")){
    aux <- aux %>% select(-Cancer)
  }
  cells <- rbind(cells, aux)
}
rm(aux)
head(cells)
```

```{r}
cells %>% group_by(sampleID) %>% 
          summarise(max(HLA.DPB1))
```


#Cancer cell proportion by patient
```{r}
cancer.prop <- cells %>% group_by(Sample, Global) %>% 
            summarise(n = n()) %>%
            mutate(Cancer.prop = n * 100 / sum(n)) %>% 
            filter(Global == "Cancer") %>% 
            separate(Sample, c("Patient", "Tissue"), sep = "_") %>% 
            select(Patient, Tissue, Cancer.prop)

cancer.prop
```


#Merging cell info with neighboor information, selecting cancer cells
```{r}
cells.n <- cbind(cells, neighboors)


cells.sel <- cells.n %>%
                  filter(Sample %in% gates.mhcii$sample) %>%
                  filter(Global=="Cancer") %>% 
                  select(-Sample)

cells.sel$Patient <- sapply(cells.sel$sampleID, function(x){strsplit(x,"_")[[1]][1]})
```



```{r}
reduced.gate.vals <- c(0, 0.01,0.02,0.03,0.04,0.05)
cut.pfs.vals <- c(8:16)
#reduced.gate.vals <- 0.04
#cut.pfs.vals <- 11
gates.mhcii <- gates.mhcii2
stop_all <- FALSE

formulas.kapplan <- c('Surv(PFS, PFS.bin) ~ MHCII.bin2')


suppressWarnings(
  for (r in reduced.gate.vals){
    if (stop_all){break}
    gates.mhcii$gate.val <- gates.mhcii$gate.val - r
    MHCII.stat <- NULL
    for (s in unique(cells.sel$sampleID)){
      s.sel <- cells.sel %>% filter(sampleID == {s})
      s.gate <- gates.mhcii[gates.mhcii$sample == s,"gate.val"]
      aux <- ifelse(s.sel$HLA.DPB1 >= s.gate,"MHCII.pos","MHCII.neg")
      MHCII.stat <- c(MHCII.stat, aux)
    }
    cells.sel$MHCII.stat <- MHCII.stat
    
    MHCII.prop.pat <- cells.sel %>% group_by(Patient) %>%
                        summarise(Prop.MHCII.pos=sum(MHCII.stat=="MHCII.pos") * 100/n()) %>%
                        filter(!Patient %in% c("S057","S123"))
    MHCII.prop.pat <- merge(cancer.prop, MHCII.prop.pat)
    pat.clin.MHCII <- merge(MHCII.prop.pat, clinical.data, by.x="Patient", by.y="study_id")
    #pat.clin.MHCII <- pat.clin.MHCII[which(pat.clin.MHCII$surgery_type == "IDS"),]
    #pat.clin.MHCII <- pat.clin.MHCII[which(pat.clin.MHCII$therapy_type == "adjuvant"),]
    
    mhcii.cuts <- quantile(pat.clin.MHCII$Prop.MHCII.pos, seq(0.5,0.85, by=0.01))

    for (cut.pfs in cut.pfs.vals){
      if (stop_all){break}
      #Stratifying according to PFS time into short and long PFS
      PFS.status <- ifelse(pat.clin.MHCII$PFS_progression > cut.pfs,"PFS.long","PFS.short")
      
      #In case if not progression after the PFS.cut and follow up time is longer, then name this as PFI.long
      for(i in which(is.na(PFS.status))){
        if (pat.clin.MHCII[i,"PFS_noProgression"] > cut.pfs){
          PFS.status[i] <- "PFS.long"
        }
      }
      pat.clin.MHCII$PFS.status <- PFS.status
      
      #Filtering those with low follow up time
      pat.clin.MHCII <- pat.clin.MHCII[which(!is.na(pat.clin.MHCII$PFS.status)),]
      
      u.test1 <- wilcox.test(pat.clin.MHCII[pat.clin.MHCII$PFS.status == "PFS.long","Prop.MHCII.pos"],
                  pat.clin.MHCII[pat.clin.MHCII$PFS.status == "PFS.short","Prop.MHCII.pos"],
                  alternative = "greater")
      if (u.test1$p.value < 0.1){
        print(paste0("Value r: ", r, " PFS.cut: ", cut.pfs, " U.test p-val: ", round(u.test1$p.value, 3)))
      }
      
      
      for (cut.mhcii in mhcii.cuts){
          numbers.pfs.mhciipos  <- pat.clin.MHCII %>%
                        mutate(MHCII.bin = ifelse(Prop.MHCII.pos >= cut.mhcii, "MHCII.high","MHCII.low"))
          
          #Stratification to perform fisher exact test  
          val1 <- sum(numbers.pfs.mhciipos$MHCII.bin == "MHCII.high" &  numbers.pfs.mhciipos$PFS.status == "PFS.long")
          val2 <- sum(numbers.pfs.mhciipos$MHCII.bin == "MHCII.high" &  numbers.pfs.mhciipos$PFS.status == "PFS.short")
          val3 <- sum(numbers.pfs.mhciipos$MHCII.bin == "MHCII.low" &  numbers.pfs.mhciipos$PFS.status == "PFS.long")
          val4 <- sum(numbers.pfs.mhciipos$MHCII.bin == "MHCII.low" &  numbers.pfs.mhciipos$PFS.status == "PFS.short")
          my.matrix <- matrix(c(val1, val2, val3, val4), nrow = 2)
          ftest <- fisher.test(my.matrix, alternative="greater")
          
          if (ftest$p.value < 0.1){
              print(paste0("Value r: ", r, " PFS.cut: ", cut.pfs, " U.test p-val: ", round(u.test1$p.value, 3)))
              print(paste0("MHCII cut: ", cut.mhcii, " Fisher test pval: ", round(ftest$p.value, 3)))
              print(my.matrix)
          }
          
          
          #Preparing data for Log-rank test
          numbers.pfs.mhciipos$PFS <- numbers.pfs.mhciipos$PFS_progression
          if(any(is.na(numbers.pfs.mhciipos$PFS))){
            numbers.pfs.mhciipos$PFS[which(is.na(numbers.pfs.mhciipos$PFS))] <- numbers.pfs.mhciipos$PFS_noProgression[which(is.na(numbers.pfs.mhciipos$PFS))]
          }
          numbers.pfs.mhciipos$PFS.bin <- ifelse(numbers.pfs.mhciipos$PFS.status == "PFS.long",1,0)
          numbers.pfs.mhciipos$MHCII.bin2 <- ifelse(numbers.pfs.mhciipos$MHCII.bin == "MHCII.high",1,0)
          
          #Log-rank test
          fit <- survfit(as.formula(formulas.kapplan), data = numbers.pfs.mhciipos)
          logrank = surv_pvalue(fit)
          pval=logrank$pval.txt
          if (pval < 0.1){
                print(paste0("Value r: ", r, " PFS.cut: ", cut.pfs, " U.test p-val: ", round(pval, 3)))
                print(paste0("MHCII cut: ", cut.mhcii, " Log rank pval: ", pval))
                print(my.matrix)
          }
      }
      
      #U test using PFS time difference between MHCII.high and MHCII.low patients
      u.test2 <- wilcox.test(numbers.pfs.mhciipos[numbers.pfs.mhciipos$MHCII.bin == "MHCII.high","PFS"],
                  numbers.pfs.mhciipos[numbers.pfs.mhciipos$MHCII.bin == "MHCII.low","PFS"],
                  alternative = "greater")
      if (u.test2$p.value < 0.1){
              print(paste0("Value r: ", r, " PFS.cut: ", cut.pfs, " U.test p-val: ", round(u.test2$p.value, 3)))
              print(paste0("MHCII cut: ", cut.mhcii, " Log rank pval: ", pval))
      }
          

      # p <- ggplot(pat.clin.MHCII, aes(x=PFS.status, y=Prop.MHCII.pos))  +
      #         geom_point(position= position_jitter(width= .1)) +
      #         geom_boxplot(fill=NA, outlier.shape = NA) +
      #         theme_bw() + ylab("Proportion of MHCII+ cancer cells")
      # print(p)
    }
}
)
```

#Setting cutoff values for MHCII pos, calculating the proportion of MHCII+ by Sample
```{r}
gates.mhcii <- gates.mhcii2
gates.mhcii$gate.val <- gates.mhcii$gate.val

MHCII.stat <- NULL
for (s in unique(cells.sel$sampleID)){
  s.sel <- cells.sel %>% filter(sampleID == {s})
  s.gate <- gates.mhcii[gates.mhcii$sample == s,"gate.val"]
  aux <- ifelse(s.sel$HLA.DPB1 >= s.gate,"MHCII.pos","MHCII.neg")
  MHCII.stat <- c(MHCII.stat, aux)
}
cells.sel$MHCII.stat <- MHCII.stat

cells.sel$CD4.T.cell[is.na(cells.sel$CD4.T.cell)] <- 0

MHCII.prop.pat <- cells.sel %>% group_by(Patient) %>%
                    mutate(CD4.plus.MHCII = ifelse(MHCII.stat == "MHCII.pos" & CD4.T.cell >= 1, "Pos","Neg")) %>% 
                    summarise(Prop.CD4.MHCII=sum(CD4.plus.MHCII=="Pos") * 100/n(),
                              Prop.MHCII.pos=sum(MHCII.stat=="MHCII.pos") * 100/n())


pat.clin.MHCII <- merge(MHCII.prop.pat, clinical.data, by.x="Patient", by.y="study_id")
pat.clin.MHCII <- merge(pat.clin.MHCII, cancer.prop, by="Patient", all.x = TRUE)

#S057 (got Olaparib at first line); S123 (got Olaparib, is HRP sBRCA1)
#Ignoring samples with R0, they can bias PFS analysis
pat.clin.MHCII <- pat.clin.MHCII %>% filter(!Patient %in% c("S057","S123"),
                                            residual != "R0")
pat.clin.MHCII
```




####Calculating by patient/sample the propotion of MHCII
```{r}
out.put.folder <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/05_ML_validations/Whole_slide_validation/"

pat.clin.MHCII$PFS.time <- pat.clin.MHCII$PFS_progression
missing.samples <- which(is.na(pat.clin.MHCII$PFS.time))
pat.clin.MHCII$PFS.time[missing.samples] <- pat.clin.MHCII$PFS_noProgression[missing.samples]

df <- pat.clin.MHCII

ggplot(df, aes(x=PFS.time, y=Prop.MHCII.pos)) +
  geom_point() + theme_bw()  + geom_smooth(method="lm") + stat_cor(method="spearman")


ggplot(df, aes(x=Prop.CD4.MHCII, y=Prop.MHCII.pos)) +
  geom_point() + theme_bw() + geom_smooth(method="lm")


df <- df %>% 
  mutate(Pathologist.cancer.prop = case_when(Tumor.percentaje.Anni.mean < 10 ~ "<10",
                                      Tumor.percentaje.Anni.mean >= 10 & Tumor.percentaje.Anni.mean < 50 ~ "10-50",
                                                  Tumor.percentaje.Anni.mean >= 50 ~ ">50"))
df$Pathologist.cancer.prop <- factor(df$Pathologist.cancer.prop, levels=c("<10","10-50", ">50"))


p <- ggplot(df, aes(x=Cancer.prop, y=Tumor.percentaje.Anni.mean)) + 
     geom_point() + theme_bw() +
     geom_smooth(method="lm") +
     geom_abline (intercept=0, lty=2, col="grey") +
     stat_cor(method = "pearson") + xlim(0,80) + ylim(0,80) +
     xlab("tCycif cancer cell proportion") +
     ylab ("Cancer cells proportion by pathologist") + 
     theme(axis.text = element_text(size=rel(1.1)),
           axis.title = element_text(size=rel(1.1)))
print(p)
ggsave(p, file=paste0(out.put.folder, "/Correlation_Pathologist_tCycif_cancer_prop.svg"),
       width = 9, height = 9, units="cm" )

my_comparisons <- list( c("<10", "10-50"), c("10-50", ">50"), c("<10", ">50") )

p <- ggboxplot(df, x="Pathologist.cancer.prop", y="Cancer.prop", add = "jitter") + 
     stat_compare_means(comparisons = my_comparisons) +
     theme_bw() +
     xlab("tCycif cancer cell proportion") +
     ylab ("Cancer cells proportion by pathologist") + 
     theme(axis.text = element_text(size=rel(1.1)),
           axis.title = element_text(size=rel(1.1)))
print(p)
ggsave(p, file=paste0(out.put.folder, "/Correlation_Pathologist_tCycif_cancer_Boxplot.svg"),
       width = 9, height = 9, units="cm" )

```

#Defining function to calculate Fold.changes in proportion of neighboors
```{r}
#The input should have a column for each cell.type neighboor of interest
get.fold.changes <- function(input.dat, cells.interest, comparison.name="All.cells"){
      fold.changes <- sapply(cells.interest, function(x){
        df <- input.dat[!is.na(input.dat[,x]),]
        fold.change <- mean(df[df$MHCII.stat=="MHCII.pos",x]) /
                                  mean(df[df$MHCII.stat=="MHCII.neg",x])
        #fold.change <- fold.change -1
        
        utest <- wilcox.test(df[df$MHCII.stat=="MHCII.pos",x],
                                  df[df$MHCII.stat=="MHCII.neg",x])
        pval1 <- utest$p.value
        return(c(fold.change, pval1))
      })
      counts <- table(input.dat$MHCII.stat)
      fold.changes <- as.data.frame(t(fold.changes))
      fold.changes <- rownames_to_column(fold.changes)
      colnames(fold.changes) <- c("Neighbors","Fold.chage","Pval")
      fold.changes$MHCII.high.prop = round(counts[1]/(counts[2] + counts[1]),2)
      fold.changes$class = comparison.name
      return(fold.changes)
}
```


#Calculating if MHCII+ have higher amount of immune 
```{r}
neighbors.interest <- c("CD8.T.cell", "CD4.T.cell", "CD11c.Myeloid", "IBA1.Myeloid" ,"Stromal")

fold.neighbors <- get.fold.changes(cells.sel, neighbors.interest, comparison.name="HRP")
```


```{r}
fold.neighbors$log2FC <- log2(fold.neighbors$Fold.chage)
fold.neighbors
```



```{r}
out.put.folder <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/05_ML_validations/Whole_slide_validation/"

df <- fold.neighbors[fold.neighbors$Pval < 0.05,]
df$Neighbors <- factor(df$Neighbors, levels=intersect(neighbors.interest, df$Neighbors))

p <- ggplot(df, aes(x=class, y=Neighbors)) + geom_point(aes(col=log2FC), size=5) + 
        scale_colour_gradient2(low= "blue", mid="grey90", high ="red", midpoint = 0, limits=c(-1,2)) +
        scale_radius(range = c(1,6), limits = c(1, 4)) +
        # scale_radius(range = c(2,6), breaks = c(1, 2, 3, 4),
        #              labels = c("<0.05", "<1^-3", "<1^-6", "<1^-12")) +
        theme_classic() +
        theme(axis.text.x=element_text(size=rel(1.1), angle = 45, vjust = 1.01, hjust=1),
                   axis.text.y=element_text(size=rel(1.1)),
                   axis.title.x=element_blank(),
                   legend.text=element_text(size=rel(1.1)),
                   legend.title=element_text(size=rel(1.1)),
                   axis.title=element_text(size=rel(1.1)),
                   legend.position = "right") +
        ylab("Cancer cell's neighboors") +
        labs(col='log2(FC)', size='FDR', title = "MHCII+/MHCII- cancer")
print(p)
ggsave(p, file=paste0(out.put.folder,"Fold_changes_MHCII_proximity2.png"), width=8, height=7, units="cm")
ggsave(p, file=paste0(out.put.folder,"Fold_changes_MHCII_proximity2.svg"), width=8, height=7, units="cm")
```



#Stratification of patients according to the proportion of MHCII+ cancer cells
```{r}
PFS.cut <- 12


pat.clin.MHCII$PFS <- pat.clin.MHCII$PFS_progression
no.progressions <- which(is.na(pat.clin.MHCII$PFS))
pat.clin.MHCII$PFS[no.progressions] <- pat.clin.MHCII$PFS_noProgression[no.progressions]
pat.clin.MHCII$PFS.bin <- ifelse(pat.clin.MHCII$progression_yn=="yes",1,0)
pat.clin.MHCII$MHCII.bin <- factor(ifelse(pat.clin.MHCII$Prop.MHCII.pos>7.5, 1,0))

#Selecting only patients with progression, or a follow up time longer that PFS.cut
pat.clin.MHCII2 <- pat.clin.MHCII %>%
                                  filter(PFS_noProgression >= PFS.cut | progression_yn=="yes")
pat.clin.MHCII2$PFS.status <- ifelse(pat.clin.MHCII2$PFS >= PFS.cut,"Long","Short")

ggplot(pat.clin.MHCII2, aes(x=PFS, y=Prop.MHCII.pos, col=PFS.status)) +
  geom_point() + theme_bw()


pat.clin.MHCII2 %>% select(PFS, PFS_noProgression, PFS.bin)

```

```{r}
pat.clin.MHCII2 %>% group_by(MHCII.bin, PFS.status) %>% 
                  summarise(n=n())


pat.clin.MHCII2 %>% group_by(MHCII.bin, progression_yn) %>% 
                  summarise(n=n())
```
#Visualization of data
```{r}
p <- ggplot(pat.clin.MHCII2, aes(x=reorder(Patient, Prop.MHCII.pos), y=Prop.MHCII.pos)) +
        geom_col(aes(fill=PFS.status)) + theme_bw() + xlab("Patient") +
        ylab("% of MHCII+ cancer cells") +
        scale_fill_manual(values=c("green4","gold")) +
        theme(axis.text.x = element_text(angle=75, hjust = 1)) #+
        #geom_hline(yintercept = 8, lty=2)
print(p)
ggsave(p, file=paste0(out.put.folder,"Barplots_PFS_MHCIIprop2.svg"), width = 10.5, height = 7, units = "cm")
ggsave(p, file=paste0(out.put.folder,"Barplots_PFS_MHCIIprop2.png"), width = 10.5, height = 7, units = "cm")
```
##Kapplan-meir 
```{r}
df <- pat.clin.MHCII
formulas.kapplan <- c('Surv(PFS, PFS.bin) ~ MHCII.bin')

fit <- survfit(as.formula(formulas.kapplan), data = df)
logrank = surv_pvalue(fit, df)
pval=logrank$pval.txt

p <- ggsurvplot(fit, data = df, risk.table = TRUE, p.val=TRUE, break.time.by=5,
                  conf.int = TRUE, palette = c("#4d4dff", "green3"),
                  legend.labs=c("MHCII.low.prop","MHCII.high.prop"), tables.y.text = FALSE, legend.title="",
                xlab="Time (months)", ylab="PFS")
p$plot <- p$plot + ggplot2::annotate("text", x = 2, y = 0.25, label =pval, size = 5)
print(p)
p$table <- p$table + theme(plot.margin = margin(t = 0,  r = 0.1, b = 2, l = 0.4, unit = "cm"))
plot.to.save <- list(p$plot, p$table)
ggsave(file=paste0(out.put.folder,"/Kapplan_PFS_MHCII.png"),
       arrangeGrob(grobs = plot.to.save, ncol = 1),
     width = 9, height = 12, units = "cm")

ggsave(file=paste0(out.put.folder,"/Kapplan_PFS_MHCII.svg"),
       arrangeGrob(grobs = plot.to.save, ncol = 1),
     width = 9, height = 12, units = "cm")

```
```{r}
df  <- pat.clin.MHCII2
df$MHCII.bin <- factor(df$MHCII.bin, levels=c(0,1), labels=c("MHCII.Low","MHCII.High"))

my_comparisons = list(levels(df$MHCII.bin))


p <- ggboxplot(df, x="MHCII.bin", y="PFS", add.params = list(fill = "white"), outlier.shape = NA) +
      stat_compare_means(comparisons = my_comparisons, 
                         method.args = list(alternative = "greater")) +
      geom_point(position= position_jitter(width= .08)) +
      theme_bw()
print(p)
ggsave(p, file=paste0(out.put.folder,"Boxplots_MHCIIbin_PFS.png"), width = 7.5, height = 8, units = "cm")
ggsave(p, file=paste0(out.put.folder,"Boxplots_MHCIIbin_PFS.svg"), width = 7.5, height = 8, units = "cm")


my_comparisons = list(unique(df$PFS.status))
p <- ggboxplot(df, x="PFS.status", y="Prop.MHCII.pos", outlier.shape = NA) +
      stat_compare_means(comparisons = my_comparisons, 
                         method.args = list(alternative = "greater")) +
      geom_point(position= position_jitter(width= .08)) +
      ylim(0,16) +
      theme_bw()
print(p)
ggsave(p, file=paste0(out.put.folder,"Boxplots_PFS_MHCIIbin.png"), width = 5, height = 8, units = "cm")
ggsave(p, file=paste0(out.put.folder,"Boxplots_PFS_MHCIIbin.svg"), width = 5, height = 8, units = "cm")
```

#Analysis of CD8.T.cells and association with PFS
```{r}
#Calculating cell proportions by Sample
Cell.props <- cells.n %>% group_by(sampleID, Global) %>% 
                          summarise(n=n()) %>% 
                          mutate(Prop=n * 100/sum(n))

Cell.props <- as.data.frame(Cell.props)
Cell.props <- Cell.props[,-3]

mt <- Cell.props %>% pivot_wider(names_from = Global, values_from = Prop)
mt$Patient <- sapply(mt$sampleID,function(x){ strsplit(x,"_")[[1]][1]})

#Ingoring patients that are R=0 and those that got PARPi
mt.clin <- merge(mt, clinical.data, by.x="Patient", by.y="study_id")
mt.clin <- mt.clin %>% filter(!Patient %in% c("S057","S123"),
                                            residual != "R0")

#Selecting only those samples from the MHCII comparison
mt.clin.sel <- mt.clin[mt.clin$Patient %in% pat.clin.MHCII2$Patient,]


PFS.cut <- 12

mt.clin.sel$PFS <- mt.clin.sel$PFS_progression
#Selecting only patients with progression, or a follow up time longer that PFS.cut
mt.clin.sel$PFS.status <- ifelse(mt.clin.sel$PFS >= PFS.cut,"Long","Short")


my_comparisons = list(unique(mt.clin.sel$PFS.status))
p <- ggboxplot(mt.clin.sel, x="PFS.status", y="CD8.T.cell", outlier.shape = NA) +
      stat_compare_means(comparisons = my_comparisons, 
                         method.args = list(alternative = "greater")) +
      geom_point(position= position_jitter(width= .08)) +
      theme_bw()
print(p)
ggsave(p, file=paste0(out.put.folder,"Boxplots_PFS_CD8bin.png"), width = 5, height = 8, units = "cm")
ggsave(p, file=paste0(out.put.folder,"Boxplots_PFS_CD8bin.svg"), width = 5, height = 8, units = "cm")


my_comparisons = list(unique(mt.clin.sel$PFS.status))
p <- ggboxplot(mt.clin.sel, x="PFS.status", y="CD4.T.cell", outlier.shape = NA) +
      stat_compare_means(comparisons = my_comparisons, 
                         method.args = list(alternative = "greater")) +
      geom_point(position= position_jitter(width= .08)) +
      theme_bw()
print(p)
ggsave(p, file=paste0(out.put.folder,"Boxplots_PFS_CD4bin.png"), width = 5, height = 8, units = "cm")
ggsave(p, file=paste0(out.put.folder,"Boxplots_PFS_CD4bin.svg"), width = 5, height = 8, units = "cm")
```


