---
title: "R Notebook"
output: html_notebook
---


```{r}
library(ggplot2)
library(tidyverse)
library(survival)
library(survminer)
```
#Defining input variables
```{r}
clinical.data <- read.table(file="L:/ltdk_farkkila/Projects/4_CellCycle/Samples_clinical_data.csv",
                            sep=",", header = TRUE)

input.annotation.folder <- "P:/h345/afarkkilab/Projects/NKI/Whole_slide_validation/tribus_analysis/output_data/"
annotation.suffix <- "_annotated.csv"

#Listing all input files
input.annotation.files <- list.files(input.annotation.folder, pattern = annotation.suffix, full.names =TRUE)


#MHCII gating values
gates.mhcii <- data.frame(sample=c("S057_iOme","S081_iOme","S073_iOme1","S072_iOme","S118_iOme", "S083_iOme1",
                                   "S098_iOme","S112_iOme","S091_iOme1","S100_iOme","S107_iOme",
                                   "S110_iTubR","S113_iOme","S121_iOme","S123_iOme","S130_iOme","S131_iOme",
                                   "S188_iOme", "S197_iOme","S268_iOme","S050_iAdn","S065_iOme"),
                         gate.val=c(3.0,3.2,3.4,3.2,2.8,3.1,2.9,2.7,
                                    3.1,3.2,3.1,3.0,3.2,3.3,3.5,3.0,3.2,
                                    3.2,3.1,3.3,3.0,3.5))
gates.mhcii$gate.val <- gates.mhcii$gate.val - 0.1
gates.mhcii
input.annotation.files
```

##Reading spatial data
```{r}
scimap.output.folder <- "P:/h345/afarkkilab/Projects/NKI/Whole_slide_validation/scimap_analysis/output_data/"
scimap.file <- "Spatial-cell-proportions_radious30px_new_fuction.csv"

neighboors <- read.table(file=paste0(scimap.output.folder, scimap.file), sep="," , header = TRUE)

```


```{r}
dim(neighboors)
```


#Reading input annotation files
```{r}
cells <- NULL
for (f in input.annotation.files){
  file.name <- basename(f)
  sampleid <- paste(strsplit(file.name, "_")[[1]][1:2], collapse = "_")
  aux <- read.table(file=f, header = TRUE, sep=",")
  aux <- aux[,-1]
  aux[,1] <- sampleid
  colnames(aux)[1] <- "sampleID"
  if(any(colnames(aux) == "Cancer")){
    aux <- aux %>% select(-Cancer)
  }
  cells <- rbind(cells, aux)
}
rm(aux)
head(cells)
```

```{r}
cells %>% group_by(sampleID) %>% 
          summarise(max(HLA.DPB1))

```


#Reading raw data
```{r}
# input.annotation.folder <- "P:/h345/afarkkilab/Projects/NKI/Whole_slide_validation/tribus_analysis/input_data/"
# annotation.suffix <- ".csv"
# 
# #Listing all input files
# input.annotation.files <- list.files(input.annotation.folder, pattern = annotation.suffix, full.names =TRUE)
# 
# raw.data <- NULL
# for (f in input.annotation.files){
#   file.name <- basename(f)
#   sampleid <- paste(strsplit(file.name, "_")[[1]][1:2], collapse = "_")
#   aux <- read.table(file=f, header = TRUE, sep=",")
#   aux <- aux[,-1]
#   aux[,1] <- sampleid
#   colnames(aux)[1] <- "sampleID"
#   raw.data <- rbind(raw.data, aux)
# }
# rm(aux)
# head(raw.data)
```

#Cancer cell proportion by patient
```{r}
cancer.prop <- cells %>% group_by(Sample, Global) %>% 
            summarise(n = n()) %>%
            mutate(proportion = n * 100 / sum(n)) %>% 
            filter(Global == "Cancer") %>% 
            separate(Sample, c("Patient", "Tissue"), sep = "_")
cancer.prop
```


#Merging cell info with neighboor information, selecting cancer cells
```{r}
cells.n <- cbind(cells, neighboors)


cells.sel <- cells.n %>%
                  filter(Sample %in% gates.mhcii$sample) %>%
                  filter(Global=="Cancer") %>% 
                  select(-Sample)
                  #mutate(MHCII.bin = ifelse(HLA.DPB1 > 2.5, "MHCII.pos","MHCII.neg"))

cells.sel$Patient <- sapply(cells.sel$sampleID, function(x){strsplit(x,"_")[[1]][1]})

MHCII.stat <- NULL
for (s in unique(cells.sel$sampleID)){
  s.sel <- cells.sel %>% filter(sampleID == {s})
  s.gate <- gates.mhcii[gates.mhcii$sample == s,"gate.val"]
  aux <- ifelse(s.sel$HLA.DPB1 >= s.gate,"MHCII.pos","MHCII.neg")
  MHCII.stat <- c(MHCII.stat, aux)
}
cells.sel$MHCII.stat <- MHCII.stat
```


####Calculating by patient/sample the propotion of MHCII
```{r}
MHCII.prop.pat <- cells.sel %>% group_by(Patient) %>%
                        summarise(Prop.MHCII.pos=sum(MHCII.stat=="MHCII.pos")/n())
MHCII.prop.pat <- merge(cancer.prop, MHCII.prop.pat)
MHCII.prop.pat

ggplot(MHCII.prop.pat, aes(x=proportion, y=Prop.MHCII.pos, col=Tissue)) +
  geom_point() + xlab("Cancer cell %") + theme_bw() 
```

#Defining function to calculate Fold.changes in proportion of neighboors
```{r}
#The input should have a column for each cell.type neighboor of interest
get.fold.changes <- function(input.dat, cells.interest, comparison.name="All.cells"){
      fold.changes <- sapply(cells.interest, function(x){
        df <- input.dat[!is.na(input.dat[,x]),]
        fold.change <- mean(df[df$MHCII.stat=="MHCII.pos",x]) /
                                  mean(df[df$MHCII.stat=="MHCII.neg",x])
        #fold.change <- fold.change -1
        
        utest <- wilcox.test(df[df$MHCII.stat=="MHCII.pos",x],
                                  df[df$MHCII.stat=="MHCII.neg",x])
        pval1 <- utest$p.value
        return(c(fold.change, pval1))
      })
      counts <- table(input.dat$MHCII.stat)
      fold.changes <- as.data.frame(t(fold.changes))
      fold.changes <- rownames_to_column(fold.changes)
      colnames(fold.changes) <- c("Neighbors","Fold.chage","Pval")
      fold.changes$MHCII.high.prop = round(counts[1]/(counts[2] + counts[1]),2)
      fold.changes$class = comparison.name
      return(fold.changes)
}
```


#Calculating if MHCII+ have higher amount of immune 
```{r}
neighbors.interest <- c("CD8.T.cell", "CD4.T.cell", "CD11c.Myeloid", "IBA1.Myeloid" ,"Stromal")

fold.neighbors <- get.fold.changes(cells.sel, neighbors.interest, comparison.name="HRP")
```


```{r}
fold.neighbors$log2FC <- log2(fold.neighbors$Fold.chage)
fold.neighbors$Fold.chage2 <- fold.neighbors$Fold.chage -1
fold.neighbors
```



```{r}
out.put.folder <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/05_ML_validations/Whole_slide_validation/"
fold.neighbors$Neighbors <- factor(fold.neighbors$Neighbors, levels=neighbors.interest)
  
p <- ggplot(fold.neighbors, aes(x=class, y=Neighbors)) + geom_point(aes(col=log2FC), size=5) + 
        scale_colour_gradient2(low= "blue", mid="grey90", high ="red", midpoint = 0, limits=c(-1,2.3)) +
        #scale_radius(range = c(1,6), limits = c(0.3, 0.8), breaks = c(0.40, 0.50, 0.60, 0.7, 0.8)) +
        # scale_radius(range = c(2,6), breaks = c(1, 2, 3, 4),
        #              labels = c("<0.05", "<1^-3", "<1^-6", "<1^-12")) +
        theme_classic() +
        theme(axis.text.x=element_text(size=rel(1.1), angle = 45, vjust = 1.01, hjust=1),
                   axis.text.y=element_text(size=rel(1.1)),
                   axis.title.x=element_blank(),
                   legend.text=element_text(size=rel(1.1)),
                   legend.title=element_text(size=rel(1.1)),
                   axis.title=element_text(size=rel(1.1)),
                   legend.position = "right") +
        ylab("Cancer cell's neighboors") +
        labs(col='log2(FC)', size='FDR', title = "MHCII+/MHCII- cancer")
print(p)
ggsave(p, file=paste0(out.put.folder,"Fold_changes_MHCII_proximity2.png"), width=8, height=8, units="cm")
```
#Assessing clinical information associaton with MHCII+
```{r}
#Months of PFS to devide them as long or short PFS
cut.pfs <- 12 

pat.clin.MHCII <- merge(MHCII.prop.pat, clinical.data, by.x="Patient", by.y="study_id")

#Stratifiying according to MHCII.status
PFS.status <- ifelse(pat.clin.MHCII$PFS_progression > cut.pfs,"PFS.long","PFS.short")

#In case if not progression after 12months, then name this as PFI.long
for(i in which(is.na(PFS.status))){
  if (pat.clin.MHCII[i,"PFS_noProgression"] > cut.pfs){
    PFS.status[i] <- "PFS.long"
  }
}
pat.clin.MHCII$PFS.status <- PFS.status

#Filtering those with low follow up time and no residual disease
pat.clin.MHCII <- pat.clin.MHCII[-which(is.na(pat.clin.MHCII$PFS.status)),]

wilcox.test(pat.clin.MHCII[pat.clin.MHCII$PFS.status == "PFS.long","Prop.MHCII.pos"],
            pat.clin.MHCII[pat.clin.MHCII$PFS.status == "PFS.short","Prop.MHCII.pos"],
            alternative = "greater")

p <- ggplot(pat.clin.MHCII, aes(x=PFS.status, y=Prop.MHCII.pos)) +
        geom_point(position= position_jitter(width= .1)) +
        geom_boxplot(fill=NA, outlier.shape = NA) +
        theme_bw() + ylab("Proportion of MHCII+ cancer cells")
print(p)
ggsave(p, file=paste0(out.put.folder,"Boxplot_MHCII_prop.svg"), width=8, height=10, units="cm")
```
#Stratification of patients according to the proportion of MHCII+ cancer cells
```{r}
pat.clin.MHCII$PFS <- pat.clin.MHCII$PFS_progression
no.progressions <- which(is.na(pat.clin.MHCII$PFS))
pat.clin.MHCII$PFS[no.progressions] <- pat.clin.MHCII$PFS_noProgression[no.progressions]

pat.clin.MHCII$PFS.bin <- ifelse(pat.clin.MHCII$progression_yn=="yes",1,0)


pat.clin.MHCII$MHCII.bin <- ifelse(pat.clin.MHCII$Prop.MHCII.pos>0.05, 1,0)
```

```{r}
pat.clin.MHCII %>% group_by(MHCII.bin, PFS.status) %>% 
                  summarise(n=n())


pat.clin.MHCII %>% group_by(MHCII.bin, progression_yn) %>% 
                  summarise(n=n())
```
#Visualization of data
```{r}
p <- ggplot(pat.clin.MHCII, aes(x=reorder(Patient, Prop.MHCII.pos), y=Prop.MHCII.pos * 100)) +
        geom_col(aes(fill=PFS.status)) + theme_bw() + xlab("Patient") +
        ylab("% of MHCII+ cancer cells") +
        theme(axis.text.x = element_blank()) + scale_fill_manual(values=c("green4","gold")) +
        geom_hline(yintercept = 22, lty=2)
print(p)
ggsave(p, file=paste0(out.put.folder,"Barplots_PFS_MHCIIprop2.svg"), width = 10.5, height = 7, units = "cm")

```


```{r}
my.matrix <- matrix(c(7, 7, 2, 5), nrow = 2)
fisher.test(my.matrix, alternative="greater")
```


##Kapplan-meir 
```{r}
df <- pat.clin.MHCII
formulas.kapplan <- c('Surv(PFS, PFS.bin) ~ MHCII.bin')

fit <- survfit(as.formula(formulas.kapplan), data = df)
logrank = surv_pvalue(fit, df)
pval=logrank$pval.txt

p <- ggsurvplot(fit, data = df, risk.table = TRUE, p.val=TRUE,
                  conf.int = TRUE, palette = c("#4d4dff", "#ff5a36"),
                  legend.labs=c("MHCII.low.prop","MHCII.high.prop"), tables.y.text = FALSE, legend.title="")
print(p)

```


