---
title: "R Notebook"
output: html_notebook
---

```{r}
library(ggplot2)
library(gridExtra)
library("dplyr")
library(ComplexHeatmap)
library(tidyr)
library(paletteer)
library(vegan)
library(circlize)

source("D:/users/fperez/NKI_TMAs_AF/devNKI-scripts/5_Spatial-interactions/RCN_inspection_functions.R")
```

#Reading clinical info
```{r}
rcn.folder <- "Analysis_results/03_RCN/"
input.prefix <- "Cluster_ids_metaclustersmerged_"
project.folder <- "D:/users/fperez/NKI_TMAs_AF/"
output.napari <- "RCN_Napary2023_2"
out.folder.name <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/03_RCN/"
method.rcn <- "radious46_k30"


cycif2samples <- read.table(file="D:/users/fperez/NKI_TMAs_AF/devNKI-scripts/3_Cell-typeClassification/tCicyf_patients-cores_pass.csv", sep=",", header = TRUE)
clinical.info <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Clinical_data/T-CycIF.csv", sep=",", header = TRUE)

clin.tcycif <- merge(cycif2samples, clinical.info, by.x="patient", by.y="TnumberAVL")

#Merging clinical data with shallowHRD
shallowHRDresults <- read.table(file="D:/users/fperez/NKI_TMAs_AF/shallowHRDinspection/shallowHRD_summary_table.csv", header=TRUE, sep=",")

IDs <- read.table(file="D:/users/fperez/NKI_TMAs_AF/shallowHRDinspection/sample_names_sWGS.csv", sep=",", header=TRUE)

#List of cores to get RNC data in Napari format
cores.interest <- c(24,28,31,109,110,112,113,114,116,118,120,122,10,11,111,113,115,117,119,121,141)
cores.interest <- sort(cores.interest)
cores.interest <- paste0("core",cores.interest)
```

#Reading input RCN table file
```{r}
RCN.clusters <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Analysis_results/Spatial_analysis/Cluster_ids_radious46_k30.csv", header = TRUE, sep=",")

#Reading names for merged RCNs
RCN.annotations <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Analysis_results/03_RCN/Merge_RCNs.csv",
                              sep=",",header=TRUE)

RCN.c.ids <- paste0("RCN", sprintf('%0.2d',1:length(unique(RCN.annotations$RCN_name))),"c")
names(RCN.c.ids) <- unique(RCN.annotations$RCN_name)

RCN.c.ids.all <- sapply(RCN.annotations$RCN_name, function(x){
  RCN.c.ids[which(x == names(RCN.c.ids))]
})

RCN.annotations$RCN.id.collapsed <- RCN.c.ids.all

RCN.annotations$RCN_id <- as.integer(gsub("RCN","",RCN.annotations$RCN_id))
RCN.annotations %>% arrange(desc(RCN_id))
```


#Reading input cell dataset
```{r}
cells <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Tables/All_cells_scimap-format.csv", header = TRUE, sep=",")
```


#Adding the core.id and slide.id
```{r}
cells.aux <- cells

cycif.slide.id <- sapply(cells$imageid, function(x){
                      splitted <- strsplit(x,"_")[[1]]
                      concat <- paste(splitted[1],splitted[2],splitted[3], sep="_")
                      return(concat)
                    })

cycif.core <- sapply(cells$imageid, function(x){
                      splitted <- strsplit(x,"_")[[1]]
                      return(splitted[4])
                    })

cells.aux$cycif.slide   <-  cycif.slide.id
cells.aux$cycif.core.id <- cycif.core
```

#Merging cells ids with patient.id 
```{r}
cells.pat <- merge(cells.aux, cycif2samples, by=c("cycif.slide","cycif.core.id"))
```

#Merging cells ids with clinical info
```{r}
cells.clin <- merge(cells.pat, clinical.info, by.x="patient", by.y="TnumberAVL")
```


#Merging clinical info with shallowHRD
```{r}
colnames(shallowHRDresults)[3] <- "ShallowHRDcall"

IDs$CF.ID.number <- as.numeric(sub("X","", IDs$CF.ID.number))
clindata.swgs <- merge(clin.tcycif, IDs, by.x="CFsample", by.y="Description", all.x = TRUE)
clindata.shrd <- merge(clindata.swgs, shallowHRDresults, by.x="CF.ID.number", by.y="Sample", all.x = TRUE)


BRCAsamples <- clindata.shrd[clindata.shrd$BRCA_methylation == "BRCA hypermethylation" | grepl("BRCA", clindata.shrd$BRCA_mutation),"patient"]

CCNE1amp <- clindata.shrd[clindata.shrd$CCNE1_revised == "Amplification", "patient"]

HRDshallow <- clindata.shrd[which(clindata.shrd$ShallowHRDcall == "positive"),"patient"]
HRPshallow <- clindata.shrd[which(clindata.shrd$ShallowHRDcall == "borderline" | clindata.shrd$ShallowHRDcall == "negative"),"patient"]

BRCAnessPos <- clindata.shrd[which(clindata.shrd$BRCAness_ovarium == "BRCAness mutation"),"patient"]
BRCAnessNeg <- clindata.shrd[which(clindata.shrd$BRCAness_ovarium == "No BRCAness mutation"),"patient"]


clindata.shrd$Molecular.profile2 <- rep("Other",  nrow(clindata.shrd))

clindata.shrd[which(clindata.shrd$patient %in% HRDshallow & clindata.shrd$patient %in% BRCAnessPos), "Molecular.profile2"] <- "HRD"
clindata.shrd[which(clindata.shrd$patient %in% HRPshallow & clindata.shrd$patient %in% BRCAnessNeg), "Molecular.profile2"] <- "HRP"
#clindata.shrd[which(clindata.shrd$patient %in% HRPshallow & clindata.shrd$patient %in% BRCAnessPos), "Molecular.profile2"] <- "BRCAness+sHRD-"
#clindata.shrd[which(clindata.shrd$patient %in% HRDshallow & clindata.shrd$patient %in% BRCAnessNeg), "Molecular.profile2"] <- "BRCAness-sHRD+"

clindata.shrd[which(clindata.shrd$patient %in% HRPshallow & clindata.shrd$patient %in% BRCAnessPos), "Molecular.profile2"] <- "Other"
clindata.shrd[which(clindata.shrd$patient %in% HRDshallow & clindata.shrd$patient %in% BRCAnessNeg), "Molecular.profile2"] <- "Other"



clindata.shrd[which(clindata.shrd$patient %in% CCNE1amp), "Molecular.profile2"] <- "CCNE1amp"
clindata.shrd[which(clindata.shrd$patient %in% BRCAsamples), "Molecular.profile2"] <- "BRCAmut/met"

clindata.shrd[which(clindata.shrd$QC_BRCA_CCNE == 1), "Molecular.profile2"] <- "Other"


clindata.shrd$imageid <- paste0(clindata.shrd$cycif.slide, "_", clindata.shrd$cycif.core.id)
```



#Loop for barplots of RNC using different criteria
```{r}
method.rcn <- "radious46_k30"

cells.aux$rcn_id <- paste0("RCN",RCN.clusters$spatial_kmeans)
n.clust <- length(unique(cells.aux$rcn_id))
n.clust <- n.clust-1
cells.aux$rcn_id <- factor(cells.aux$rcn_id,
                       levels=paste0("RCN",0:n.clust))
################################################
#Exporting data for visualization in Napari
for (slide in unique(cells.aux$cycif.slide)){
  sel.cell <- cells.aux %>%
              filter(cycif.slide == slide, cycif.core.id %in% cores.interest) %>%
              arrange(cycif.core.id, CellId)

  outout.folder.name <- paste0(project.folder, "/", slide, "/", output.napari, "_",method.rcn)
  dir.create(outout.folder.name)

  for (core in unique(sel.cell$cycif.core.id)){
    sel.cell.core <- sel.cell[which(sel.cell$cycif.core.id == core),]
    df <- data.frame(Core_Names=paste0(core,"_Probabilities_cytomask2"),
                   Cellid=sel.cell.core$CellId,
                   GlobalCellType=paste0("RCN",sel.cell.core$rcn_id))
    cell.id <- sapply(df$Cellid, function(x){
                      splitted <- strsplit(x,"_")[[1]]
                      cid <- splitted[5]
                      cid <- as.integer(gsub("c","",cid))
                      return(cid)
                    })
    df$Cellid <- cell.id
    df <- df %>% arrange(Cellid)
    write.table(df, file=paste0(outout.folder.name,"/",core,"_RCN_Napari.csv"),
                sep=",", row.names = FALSE)
  }
}
# ###########################################
# #Plotting barplot for abundance of clusters
# barplot.rcn(cells.aux, method=method.rcn)
# 
# ###########################################
# #Plotting heatmap annotations with diversity and abundance of clusters
# heatmap.diversity(cells.aux, method=method.rcn)
# 
# #cells.aux2 <- cells.aux[cells.aux$patient %in% pt.sel.prop.cancer.pt,] #Selecting cores with at least 10% of cancer cells
# 
# ###########################################
# #Plotting heatmap annotations with diversity and abundance of clusters
# heatmaps(cells.aux, method=method.rcn)

```


```{r}
cells.aux %>% group_by(imageid,rcn_id) %>%
    summarise(N=n()) %>% mutate(Proportion= N * 100 /sum(N)) %>% dplyr::filter(rcn_id=="RCN16") %>%
    arrange(desc(Proportion))
```

#Identifying those patients with really low amount of cancer cells or super high amount of cancer cells
```{r}

cores.cancer.prop <- cells.aux %>% group_by(imageid, GlobalCellType) %>% 
  summarise(N=n()) %>% mutate(Proportion= N * 100 /sum(N)) %>% dplyr::filter(GlobalCellType=="Cancer") %>%
  arrange(desc(Proportion))

cores.stromal.prop <- cells.aux %>% group_by(imageid, GlobalCellType) %>% 
  summarise(N=n()) %>% mutate(Proportion= N * 100 /sum(N)) %>%
  dplyr::filter(GlobalCellType %in% "Stromal") %>%
  arrange(desc(Proportion))

high.cancer <- cores.cancer.prop$imageid[which(cores.cancer.prop$Proportion >= 85)]

high.stromal <- cores.stromal.prop$imageid[which(cores.stromal.prop$Proportion >= 85)]
```




#Loop for barplots of RNC after merging  using different criteria
```{r}
cells.aux$rcn_id <- RCN.annotations[c(RCN.clusters$spatial_kmeans + 1),"RCN.id.collapsed"]
cells.aux$rcn_id <- factor(cells.aux$rcn_id)

rcn.folder <- "Analysis_results/03_RCN/"
input.prefix <- "Cluster_ids_metaclustersmerged_"
project.folder <- "D:/users/fperez/NKI_TMAs_AF/"
output.napari <- "RCN_Napary2023"
out.folder.name <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/03_RCN/"
method.rcn <- "radious46_k30_merged"


################################################
#Exporting data for visualization in Napari
for (slide in unique(cells.aux$cycif.slide)[c(1:4,10)]){
  # sel.cell <- cells.aux %>%
  #             filter(cycif.slide == slide, cycif.core.id %in% cores.interest) %>%
  #             arrange(cycif.core.id, CellId)
  sel.cell <- cells.aux %>%
               filter(cycif.slide == slide) %>% arrange(cycif.core.id, CellId)
              #filter(cycif.slide == slide, cycif.core.id %in% cores.interest) %>%
              #

  outout.folder.name <- paste0(project.folder, "/", slide, "/", output.napari, "_",method.rcn)
  dir.create(outout.folder.name)

  for (core in unique(sel.cell$cycif.core.id)){
    sel.cell.core <- sel.cell[which(sel.cell$cycif.core.id == core),]
    df <- data.frame(Core_Names=paste0(core,"_Probabilities_cytomask2"),
                   Cellid=sel.cell.core$CellId,
                   GlobalCellType=paste0("RCN",sel.cell.core$rcn_id))
    cell.id <- sapply(df$Cellid, function(x){
                      splitted <- strsplit(x,"_")[[1]]
                      cid <- splitted[5]
                      cid <- as.integer(gsub("c","",cid))
                      return(cid)
                    })
    df$Cellid <- cell.id
    df <- df %>% arrange(Cellid)
    write.table(df, file=paste0(outout.folder.name,"/",core,"_RCN_Napari.csv"),
                sep=",", row.names = FALSE)
  }
}
# ###########################################
#Plotting barplot for abundance of clusters
barplot.rcn(cells.aux, method=method.rcn)

###########################################
#Plotting heatmap annotations with diversity and abundance of clusters
heatmap.diversity(cells.aux, method=method.rcn)

cells.aux.clin <- merge(cells.aux, clindata.shrd, by="imageid", all.x = TRUE)

#Ignoring cores with high amount of cancer or stromal
cells.aux.clin2 <- cells.aux.clin[!(cells.aux.clin$imageid %in% high.cancer |
                                      cells.aux.clin$imageid %in% high.stromal),]

#Ignoring patients with only one core
patient.cores.nums <- cells.aux.clin2 %>% group_by(patient, imageid) %>%
  summarise(N.cores=length(unique(imageid))) %>%
  summarise(N.cores2=sum(N.cores))

patients.low.num.cores <- patient.cores.nums$patient[which(patient.cores.nums$N.cores2 < 2)]
cells.aux.clin2 <- cells.aux.clin2[!cells.aux.clin2$patient %in% patients.low.num.cores,]

#Additionally ignore those patients with total amount of cancer cells very low
tot.prop.cancer <- cells.aux.clin2 %>% group_by(patient, GlobalCellType) %>%
                      summarise(N=n()) %>% mutate(Proportion = N * 100 /sum(N)) %>%
                      dplyr::filter(GlobalCellType=="Cancer") %>% arrange(Proportion)

patients.low.cancer <- tot.prop.cancer$patient[which(tot.prop.cancer$Proportion < 10)]
cells.aux.clin2 <- cells.aux.clin2[!cells.aux.clin2$patient %in% patients.low.cancer,]

#Ignoring RCN01c and RCN02c
cells.aux.clin2 <- cells.aux.clin2[!cells.aux.clin2$rcn_id %in% c("RCN01c","RCN02c"),]

###########################################
#Plotting heatmap annotations with diversity and abundance of clusters
heatmaps(cells.aux.clin2, method=method.rcn)
```



#Saving proportion of RCN_id by image
```{r}
cells.by.image <- cells.aux.clin %>% group_by(imageid, rcn_id) %>% select(patient, imageid, rcn_id) %>%
                    summarise(N=n()) %>%
                    mutate(Proportion = round(N * 100 /sum(N), 3))
write.table(cells.by.image, file=paste0(out.folder.name,"RCN_by_image.csv"),sep=",", row.names = FALSE)
```


```{r}
ggplot(cells.by.image, aes(x=rcn_id, y=Proportion)) + geom_boxplot(fill=NA) +
  geom_point(aes(y=Proportion, x=rcn_id), position= position_jitter(width= .3), color="darkgrey",
             size= 2, alpha = 0.5) +
  theme_bw()


ggplot(cells.by.image[!cells.by.image$rcn_id %in% c("RCN01c","RCN02c"),], aes(x=rcn_id, y=Proportion)) +
  geom_point(aes(y=Proportion, x=rcn_id), position= position_jitter(width= .3), color="darkgrey",
             size= 2, alpha = 0.5) +
  geom_boxplot(fill=NA, outlier.shape = NA) +
  theme_bw()
```

