---
title: "R Notebook"
output: html_notebook
---


```{r}
library(ggplot2)
library(survival)
library(dplyr)
library(tidyverse)
```


```{r}
out.folder.name <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/03_RCN/"
```


#Reading clinical info
```{r}
cycif2samples <- read.table(file="D:/users/fperez/NKI_TMAs_AF/devNKI-scripts/3_Cell-typeClassification/tCicyf_patients-cores_pass.csv", sep=",", header = TRUE)
clinical.info <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Clinical_data/T-CycIF.csv", sep=",", header = TRUE)

clin.tcycif <- merge(cycif2samples, clinical.info, by.x="patient", by.y="TnumberAVL")

#Merging clinical data with shallowHRD
shallowHRDresults <- read.table(file="D:/users/fperez/NKI_TMAs_AF/shallowHRDinspection/shallowHRD_summary_table.csv", header=TRUE, sep=",")

IDs <- read.table(file="D:/users/fperez/NKI_TMAs_AF/shallowHRDinspection/sample_names_sWGS.csv", sep=",", header=TRUE)
```


#Reading RCN proportions
```{r}
RCN.proportions.image <- read.table(file="D:/users/fperez/NKI_TMAs_AF/Analysis_results/03_RCN/RCN_by_image.csv", sep=",",header = TRUE)
```

#Merging clinical info with shallowHRD and creating new Molecular merker
```{r}
colnames(shallowHRDresults)[3] <- "ShallowHRDcall"

IDs$CF.ID.number <- as.numeric(sub("X","", IDs$CF.ID.number))
clindata.swgs <- merge(clin.tcycif, IDs, by.x="CFsample", by.y="Description", all.x = TRUE)
clindata.shrd <- merge(clindata.swgs, shallowHRDresults, by.x="CF.ID.number", by.y="Sample", all.x = TRUE)


BRCAsamples <- clindata.shrd[clindata.shrd$BRCA_methylation == "BRCA hypermethylation" | grepl("BRCA", clindata.shrd$BRCA_mutation),"patient"]

CCNE1amp <- clindata.shrd[clindata.shrd$CCNE1_revised == "Amplification", "patient"]

HRDshallow <- clindata.shrd[which(clindata.shrd$ShallowHRDcall == "positive"),"patient"]
HRPshallow <- clindata.shrd[which(clindata.shrd$ShallowHRDcall == "borderline" | clindata.shrd$ShallowHRDcall == "negative"),"patient"]

BRCAnessPos <- clindata.shrd[which(clindata.shrd$BRCAness_ovarium == "BRCAness mutation"),"patient"]
BRCAnessNeg <- clindata.shrd[which(clindata.shrd$BRCAness_ovarium == "No BRCAness mutation"),"patient"]


clindata.shrd$Molecular.profile2 <- rep("Other",  nrow(clindata.shrd))
clindata.shrd[which(clindata.shrd$patient %in% HRPshallow & clindata.shrd$patient %in% BRCAnessNeg), "Molecular.profile2"] <- "HRP"
clindata.shrd[which(clindata.shrd$patient %in% CCNE1amp), "Molecular.profile2"] <- "CCNE1amp"
clindata.shrd[which(clindata.shrd$patient %in% HRDshallow & clindata.shrd$patient %in% BRCAnessPos), "Molecular.profile2"] <- "HRD"
clindata.shrd[which(clindata.shrd$QC_BRCA_CCNE == 1), "Molecular.profile2"] <- "Other"
clindata.shrd[which(clindata.shrd$patient %in% BRCAsamples), "Molecular.profile2"] <- "BRCAmut/met"

clindata.shrd$imageid <- paste0(clindata.shrd$cycif.slide, "_", clindata.shrd$cycif.core.id)

colmns.interest <- c("Molecular.profile2", "patient","figo_stage","therapy_sequence","finalstatus",
                "timelastfu","daystoprogression","progression")
clindata.shrd.patient <- unique(clindata.shrd[,colmns.interest])

Molecular.profiles.new <- clindata.shrd.patient %>% select(patient, Molecular.profile2, therapy_sequence)
write.table(Molecular.profiles.new,
            file="D:/users/fperez/NKI_TMAs_AF/Tables/Molecular_profiles_patients_20231123.csv",
            sep=",", row.names=FALSE)
```



#Merging RCN information with clinical.data and calculating proportion by patient and cutoff points for high abundance
```{r}
RCN.proportions.image.patient <- merge(RCN.proportions.image, clindata.shrd, by="imageid")[,c(1,2,3,7)]

RCN.proportions.patient <- RCN.proportions.image.patient %>% 
                                  group_by(patient, rcn_id) %>% 
                                  summarise(Total=sum(N)) %>% 
                                  mutate(Proportion=round(Total*100/sum(Total),2))

RCN.proportions.patient <- RCN.proportions.patient[,-3] #This can be merged with Molecular.profile2 for boxplots

RCN.proportions.patient.m <- RCN.proportions.patient %>%
                              pivot_wider(names_from = rcn_id, values_from = Proportion)


RCN.proportions.patient.m <- as.data.frame(RCN.proportions.patient.m)
row.names(RCN.proportions.patient.m) <- RCN.proportions.patient.m[,1]
RCN.proportions.patient.m[is.na(RCN.proportions.patient.m)] <- 0

RCN.proportions.patient.m.clin <- merge(RCN.proportions.patient.m, clindata.shrd,
                                        by="patient")

RCN.proportions.patient.clin <- merge(RCN.proportions.patient, clindata.shrd.patient,
                                        by="patient")


cutpoints <- sapply(colnames(RCN.proportions.patient.m)[-1], function(x){
              var = x
              sd.var <- sd(RCN.proportions.patient.m[,var])
              median.var <- median(RCN.proportions.patient.m[,var])
              cutpoint <- median.var + sd.var
              return(cutpoint)
})
RCN.proportions.patient.clin
```

#Calculating cut-off point to stablish high abundance of RNC per patient
```{r}
RCN.proportions.patient.clin$Molecular.profile2 <- factor(RCN.proportions.patient.clin$Molecular.profile2,
                                              levels=c("BRCAmut/met","HRD","HRP","CCNE1amp","Other"))

rcn_set1 <- c("RCN01c","RCN02c")
ggplot(RCN.proportions.patient.clin[RCN.proportions.patient.clin$rcn_id %in% rcn_set1,],
       aes(x=Molecular.profile2,y=Proportion, fill=Molecular.profile2)) +
  geom_boxplot() + facet_wrap(~rcn_id) + theme_bw() + xlab("Molecular profile") +
  theme(axis.text.x = element_blank()) 


rcn_set1 <- c("RCN03c","RCN05c","RCN07c","RCN08c","RCN09c","RCN11c","RCN12c","RCN13c")
ggplot(RCN.proportions.patient.clin[RCN.proportions.patient.clin$rcn_id %in% rcn_set1,],
       aes(x=Molecular.profile2,y=Proportion, fill=Molecular.profile2)) +
  geom_boxplot() + facet_wrap(~rcn_id) + theme_bw() + xlab("Molecular profile") +
  theme(axis.text.x = element_blank()) + 
  ylim(0,10)
```


```{r}
types <- c(".","NACT")
suffixes <- c("ALL","NACT","PDS")
output.folder <- "D:/users/fperez/NKI_TMAs_AF/Analysis_results/03_RCN/"

for (i in 1:3){
print(suffixes[i])
RCN.sel <- RCN.proportions.patient.clin[grepl(types[i],                                      
                                               RCN.proportions.patient.clin$therapy_sequence),]

if (i == 3){
  RCN.sel <- RCN.proportions.patient.clin[!grepl("NACT",RCN.proportions.patient.clin$therapy_sequence),]
}


#Plot for RCNs with higher abundace
#Statistics of difference in abundance of RNC4 vs RNC6 by Molecular profiles
p.vals.RCN4vsRCN6<- sapply(unique(RCN.sel$Molecular.profile2), function(x){
  mol.dat <- RCN.sel[RCN.sel$Molecular.profile2 %in% x,]
  prop.rcn4 <- mol.dat$Proportion[mol.dat$rcn_id == "RCN04c"]
  prop.rcn6 <- mol.dat$Proportion[mol.dat$rcn_id == "RCN06c"]
  wtest <- wilcox.test(prop.rcn4, prop.rcn6)
  wtest$p.value
})
#Organizing p-values in data.frame for plotting using geom_text
p.vals.dataframe <- data.frame(pval=p.vals.RCN4vsRCN6,
                     Molecular.profile=unique(RCN.sel$Molecular.profile2),
                     y.val=28,
                     rcn_id="RCN04c")

rcn_set1 <- c("RCN04c","RCN06c")
p <- ggplot(RCN.sel[RCN.sel$rcn_id %in% rcn_set1,],
       aes(x=Molecular.profile2,y=Proportion, fill=rcn_id)) +
  geom_boxplot(outlier.shape = NA) + theme_bw() + xlab("Molecular profile") +
  theme(legend.title = element_blank(), axis.text.x=element_text(size=rel(1.1)),
        axis.title.x = element_text(size=rel(1.1)), axis.title.y = element_text(size=rel(1.1)),
        axis.text.y=element_text(size=rel(1.1))) +
  scale_fill_manual(values = c("grey85", "grey30")) + ylim(0,32) +
  geom_text(data = p.vals.dataframe[p.vals.dataframe$pval < 0.05,], aes(x=Molecular.profile, y=y.val),
            label="*", size=8)
ggsave(p, filename = paste0(output.folder, "/Barplot_RNC6vsRCN4_", suffixes[i],".png"), 
       width = 12, height = 10, units = "cm")


#For RCN06c
rcn_set1 <- c("RCN06c")
print(rcn_set1)
#Statistics of difference in abundance of RNC6 in BRCAmut samples vs other Molecular profiles
p.vals.RCN6vsBRCA<- sapply(unique(RCN.sel$Molecular.profile2), function(x){
  mol.dat <- RCN.sel[RCN.sel$rcn_id %in% rcn_set1,]
  prop.BRCAmut <- mol.dat$Proportion[mol.dat$Molecular.profile2 == "BRCAmut/met"]
  prop.x <- mol.dat$Proportion[mol.dat$Molecular.profile2 == x]
  wtest <- wilcox.test(prop.BRCAmut, prop.x)
  wtest$p.value
})
#Organizing p-values in data.frame for plotting using geom_text
p.vals.dataframe <- data.frame(pval=p.vals.RCN6vsBRCA,
                     Molecular.profile=unique(RCN.sel$Molecular.profile2),
                     y.val=18,
                     Molecular.profile2=unique(RCN.sel$Molecular.profile2),
                     rcn_id=rcn_set1)

p <- ggplot(RCN.sel[RCN.sel$rcn_id %in% rcn_set1,],
       aes(x=Molecular.profile2,y=Proportion, fill=Molecular.profile2)) +
  geom_boxplot(outlier.shape = NA) + theme_bw() + xlab("Molecular profile") +
  theme(axis.text.x=element_blank(), axis.title.x = element_text(size=rel(1.1)),
        axis.title.y = element_text(size=rel(1.1)), axis.text.y=element_text(size=rel(1.1))) + ylim(0,30) +
  geom_text(data = p.vals.dataframe[p.vals.dataframe$pval < 0.05,], aes(x=Molecular.profile, y=y.val),
            label="*", size=8) +
  annotate("text", x=2.5, y=30, label="P-values calculated againts BRCAmut", fontface="italic", size = 2) +
  ggtitle(rcn_set1) + labs(fill = "Molecular profile")
ggsave(p, filename = paste0(output.folder, "Barplot_RNC06vsBRCAmut_", suffixes[i], ".png"), 
       width = 10, height = 10,  units = "cm")


#For RCN05c
rcn_set1 <- c("RCN05c")
print(rcn_set1)
#Statistics of difference in abundance of RCN05c in BRCAmut samples vs other Molecular profiles
p.vals.RCN6vsBRCA<- sapply(unique(RCN.sel$Molecular.profile2), function(x){
  mol.dat <- RCN.sel[RCN.sel$rcn_id %in% rcn_set1,]
  prop.BRCAmut <- mol.dat$Proportion[mol.dat$Molecular.profile2 == "BRCAmut/met"]
  prop.x <- mol.dat$Proportion[mol.dat$Molecular.profile2 == x]
  wtest <- wilcox.test(prop.BRCAmut, prop.x)
  wtest$p.value
})
#Organizing p-values in data.frame for plotting using geom_text
p.vals.dataframe <- data.frame(pval=p.vals.RCN6vsBRCA,
                     Molecular.profile=unique(RCN.sel$Molecular.profile2),
                     y.val=8,
                     Molecular.profile2=unique(RCN.sel$Molecular.profile2),
                     rcn_id=rcn_set1)

p <- ggplot(RCN.sel[RCN.sel$rcn_id %in% rcn_set1,],
       aes(x=Molecular.profile2,y=Proportion, fill=Molecular.profile2)) +
  geom_boxplot() + theme_bw() + xlab("Molecular profile") +
  theme(axis.text.x=element_blank(), axis.title.x = element_text(size=rel(1.1)),
        axis.title.y = element_text(size=rel(1.1)), axis.text.y=element_text(size=rel(1.1))) + ylim(0,10) +
  geom_text(data = p.vals.dataframe[p.vals.dataframe$pval < 0.05,], aes(x=Molecular.profile, y=y.val),
            label="*", size=8) +
  annotate("text", x=2.5, y=10, label="P-values calculated againts BRCAmut", fontface="italic", size = 2) +
  ggtitle(rcn_set1) + labs(fill = "Molecular profile")
ggsave(p, filename = paste0(output.folder,"/Barplot_RNC05vsBRCAmut_", suffixes[i] ,".png"), 
       width = 10, height = 10,  units = "cm")



#For RCN11c
rcn_set1 <- c("RCN11c")
print(rcn_set1)
#Statistics of difference in abundance of RCN05c in BRCAmut samples vs other Molecular profiles
p.vals.RCN6vsBRCA<- sapply(unique(RCN.sel$Molecular.profile2), function(x){
  mol.dat <- RCN.sel[RCN.sel$rcn_id %in% rcn_set1,]
  prop.BRCAmut <- mol.dat$Proportion[mol.dat$Molecular.profile2 == "BRCAmut/met"]
  prop.x <- mol.dat$Proportion[mol.dat$Molecular.profile2 == x]
  wtest <- wilcox.test(prop.BRCAmut, prop.x)
  wtest$p.value
})
#Organizing p-values in data.frame for plotting using geom_text
p.vals.dataframe <- data.frame(pval=p.vals.RCN6vsBRCA,
                     Molecular.profile=unique(RCN.sel$Molecular.profile2),
                     y.val=6,
                     Molecular.profile2=unique(RCN.sel$Molecular.profile2),
                     rcn_id=rcn_set1)

p <- ggplot(RCN.sel[RCN.sel$rcn_id %in% rcn_set1,],
       aes(x=Molecular.profile2,y=Proportion, fill=Molecular.profile2)) +
  geom_boxplot() + theme_bw() + xlab("Molecular profile") +
  theme(axis.text.x=element_blank(), axis.title.x = element_text(size=rel(1.1)),
        axis.title.y = element_text(size=rel(1.1)), axis.text.y=element_text(size=rel(1.1))) + ylim(0,8) +
  geom_text(data = p.vals.dataframe[p.vals.dataframe$pval < 0.05,], aes(x=Molecular.profile, y=y.val),
            label="*", size=8) +
  annotate("text", x=2.5, y=8, label="P-values calculated againts BRCAmut", fontface="italic", size = 2) +
  ggtitle(rcn_set1) + labs(fill = "Molecular profile")
ggsave(p, filename = paste0(output.folder,"/Barplot_RNC11vsBRCAmut_", suffixes[i],".png"), 
       width = 10, height = 10,  units = "cm")


#For RCN12c
rcn_set1 <- "RCN12c"
print(rcn_set1)
#Statistics of difference in abundance of RCN05c in BRCAmut samples vs other Molecular profiles
p.vals.RCN6vsBRCA<- sapply(unique(RCN.sel$Molecular.profile2), function(x){
  mol.dat <- RCN.sel[RCN.sel$rcn_id %in% rcn_set1,]
  prop.BRCAmut <- mol.dat$Proportion[mol.dat$Molecular.profile2 == "CCNE1amp"]
  prop.x <- mol.dat$Proportion[mol.dat$Molecular.profile2 == x]
  wtest <- wilcox.test(prop.BRCAmut, prop.x)
  wtest$p.value
})

#Organizing p-values in data.frame for plotting using geom_text
p.vals.dataframe <- data.frame(pval=p.vals.RCN6vsBRCA,
                     Molecular.profile=unique(RCN.sel$Molecular.profile2),
                     y.val=6,
                     Molecular.profile2=unique(RCN.sel$Molecular.profile2),
                     rcn_id=rcn_set1)

p <- ggplot(RCN.sel[RCN.sel$rcn_id %in% rcn_set1,],
       aes(x=Molecular.profile2,y=Proportion, fill=Molecular.profile2)) +
  geom_boxplot() + theme_bw() + xlab("Molecular profile") +
  theme(axis.text.x=element_blank(), axis.title.x = element_text(size=rel(1.1)),
        axis.title.y = element_text(size=rel(1.1)), axis.text.y=element_text(size=rel(1.1))) + ylim(0,3) +
  geom_text(data = p.vals.dataframe[p.vals.dataframe$pval < 0.05,], aes(x=Molecular.profile, y=y.val),
            label="*", size=8) +
  annotate("text", x=2.5, y=3, label="P-values calculated againts CCNE1", fontface="italic", size = 2) +
  ggtitle(rcn_set1) + labs(fill = "Molecular profile")
ggsave(p, filename = paste0(output.folder, "/Barplot_RNC12vsCCNE1_" , suffixes[i], ".png"), 
       width = 10, height = 10,  units = "cm")



#For RCN02c, considering HRP
rcn_set1 <- "RCN02c"
print(rcn_set1)
#Statistics of difference in abundance of RCN05c in BRCAmut samples vs other Molecular profiles
p.vals.RCN6vsBRCA<- sapply(unique(RCN.sel$Molecular.profile2), function(x){
  mol.dat <- RCN.sel[RCN.sel$rcn_id %in% rcn_set1,]
  prop.BRCAmut <- mol.dat$Proportion[mol.dat$Molecular.profile2 == "HRP"]
  prop.x <- mol.dat$Proportion[mol.dat$Molecular.profile2 == x]
  wtest <- wilcox.test(prop.BRCAmut, prop.x, alternative="greater")
  wtest$p.value
})

#Organizing p-values in data.frame for plotting using geom_text
p.vals.dataframe <- data.frame(pval=p.vals.RCN6vsBRCA,
                     Molecular.profile=unique(RCN.sel$Molecular.profile2),
                     y.val=100,
                     Molecular.profile2=unique(RCN.sel$Molecular.profile2),
                     rcn_id=rcn_set1)

p <- ggplot(RCN.sel[RCN.sel$rcn_id %in% rcn_set1,],
       aes(x=Molecular.profile2,y=Proportion, fill=Molecular.profile2)) +
  geom_violin() + geom_boxplot(fill=NA, width=0.3) + theme_bw() + xlab("Molecular profile") +
  theme(axis.text.x=element_blank(), axis.title.x = element_text(size=rel(1.1)),
        axis.title.y = element_text(size=rel(1.1)), axis.text.y=element_text(size=rel(1.1))) + ylim(0,108) +
  geom_text(data = p.vals.dataframe[p.vals.dataframe$pval < 0.05,], aes(x=Molecular.profile, y=y.val),
            label="*", size=8) +
  annotate("text", x=2.5, y=108, label="P-values calculated againts HRP", fontface="italic", size = 2) +
  ggtitle(rcn_set1) + labs(fill = "Molecular profile")
ggsave(p, filename = paste0(output.folder, "/Barplot_RNC02vsHRP_", suffixes[i],".png"), 
       width = 10, height = 10,  units = "cm")


#For RCN02c, considering HRP
rcn_set1 <- "RCN04c"
print(rcn_set1)
#Statistics of difference in abundance of RCN05c in BRCAmut samples vs other Molecular profiles
p.vals.RCN6vsBRCA<- sapply(unique(RCN.sel$Molecular.profile2), function(x){
  mol.dat <- RCN.sel[RCN.sel$rcn_id %in% rcn_set1,]
  prop.BRCAmut <- mol.dat$Proportion[mol.dat$Molecular.profile2 == "HRP"]
  prop.x <- mol.dat$Proportion[mol.dat$Molecular.profile2 == x]
  wtest <- wilcox.test(prop.BRCAmut, prop.x, alternative="greater")
  wtest$p.value
})

#Organizing p-values in data.frame for plotting using geom_text
p.vals.dataframe <- data.frame(pval=p.vals.RCN6vsBRCA,
                     Molecular.profile=unique(RCN.sel$Molecular.profile2),
                     y.val=40,
                     Molecular.profile2=unique(RCN.sel$Molecular.profile2),
                     rcn_id=rcn_set1)

p <- ggplot(RCN.sel[RCN.sel$rcn_id %in% rcn_set1,],
       aes(x=Molecular.profile2,y=Proportion, fill=Molecular.profile2)) +
  geom_violin() + geom_boxplot(fill=NA, width=0.3) + theme_bw() + xlab("Molecular profile") +
  theme(axis.text.x=element_blank(), axis.title.x = element_text(size=rel(1.1)),
        axis.title.y = element_text(size=rel(1.1)), axis.text.y=element_text(size=rel(1.1))) + ylim(0,40) +
  geom_text(data = p.vals.dataframe[p.vals.dataframe$pval < 0.05,], aes(x=Molecular.profile, y=y.val),
            label="*", size=8) +
  annotate("text", x=2.5, y=108, label="P-values calculated againts HRP", fontface="italic", size = 2) +
  ggtitle(rcn_set1) + labs(fill = "Molecular profile")
ggsave(p, filename = paste0(output.folder,"/Barplot_RNC04vsHRP_", suffixes[i], ".png"), 
       width = 10, height = 10,  units = "cm")
}
```


```{r}
clin.tcycif %>% dplyr::filter(patient=="T14-05219") 

 clindata.shrd.patient %>% dplyr::filter(patient =="T14-05219")
```



```{r}
RCN.proportions.patient.clin %>% dplyr::filter(rcn_id == "RCN08c") %>% arrange(desc(Proportion))
```






```{r}
RCN.proportions.patient.clin.04 <- RCN.proportions.patient.clin[RCN.proportions.patient.clin$rcn_id %in% "RCN04c",]

RCN.proportions.patient.clin.06 <- RCN.proportions.patient.clin[RCN.proportions.patient.clin$rcn_id %in% "RCN06c",]

HRP.rcn06 <- RCN.proportions.patient.clin.06$Proportion[RCN.proportions.patient.clin.06$Molecular.profile2 == "HRP"]

HRP.rcn04 <- RCN.proportions.patient.clin.04$Proportion[RCN.proportions.patient.clin.04$Molecular.profile2 == "HRP"]
p.vals.HRP.rcn04 <- sapply(unique(RCN.proportions.patient.clin.04$Molecular.profile2), function(x){
  w.tst <- wilcox.test(HRP.rcn04, 
        RCN.proportions.patient.clin.04$Proportion[RCN.proportions.patient.clin.04$Molecular.profile2 == x])
  w.tst$p.value
})
names(p.vals.HRP.rcn04) <- unique(RCN.proportions.patient.clin.04$Molecular.profile2)

BRCAmut.rcn06 <- RCN.proportions.patient.clin.06$Proportion[RCN.proportions.patient.clin.06$Molecular.profile2 == "BRCAmut/met"]
p.vals.BRCAmut.rcn06 <- sapply(unique(RCN.proportions.patient.clin.06$Molecular.profile2), function(x){
  w.tst <- wilcox.test(BRCAmut.rcn06, 
        RCN.proportions.patient.clin.06$Proportion[RCN.proportions.patient.clin.06$Molecular.profile2 == x])
  w.tst$p.value
})
names(p.vals.BRCAmut.rcn06) <- unique(RCN.proportions.patient.clin.06$Molecular.profile2)

print(p.vals.HRP.rcn04)
print(p.vals.BRCAmut.rcn06)
wilcox.test(HRP.rcn06, HRP.rcn04)
```




#Performing Kapplan-Meir plots for the abundance of each cell.type, according to therapy sequence
```{r}
suffix <- c("PDS","NACT")
list.class <- list(c("Only debulking","Primairy debulking"),
                   c("NACT","PDS followed by NACT", "NACT followed by re-debulking"))

for (j in 1:length(suffix)){
  data.in <- cells_percentages.mc[cells_percentages.mc$therapy_sequence %in% list.class[[j]],]
  lastcol <- ncol(data.in)
  colnames(data.in)[2] <- "Bcells"
  
  plots.OS <- lapply(vars.all, function(x){
  vars = x
  cut.off.value <- cutpoints[names(cutpoints) == vars]
  cat.var <- as.factor(ifelse(data.in[,vars] >=  cut.off.value, "high","low"))
  data.in[,lastcol] <- cat.var
  colnames(data.in)[lastcol] <- "aux"
  fit <- survfit(Surv(timelastfu, finalstatus) ~ aux, data = data.in)
  logrank = surv_pvalue(fit, data.in)
  pval=logrank$pval.txt
  p <- ggsurvplot(fit, data = data.in, risk.table = TRUE, p.val=TRUE, conf.int = TRUE,
                  font.legend = c(18, "plain"), font.tickslab = c(16, "plain"),
                  font.x = c(18, "plain"), font.y = c(18, "plain"),
                  tables.y.text = FALSE,  break.time.by=500, xlim = c(0,2500),
                  legend.labs=c(paste0(vars, "+"),paste0(vars, "-")),
                  palette = c("#34A0D3" ,"#FA5A41")) + 
        guides(colour = guide_legend(nrow = 2))
  p <- p + ylab("OS probability") + xlab("Time (days)")
  p$plot <- p$plot + theme(legend.key.width = unit(3, "line"), plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"), legend.key.height = unit(1.5, "line"))
  p$plot <- p$plot + ggplot2::annotate("text", x = 2200, y = 0.85, label =pval, size = 5)
  p$table <- p$table + theme(plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"))
  plots <- list(plot=p$plot, table=p$table)
  return(plots)
  })
  fig.OS <- lapply(1:length(plots.OS) , function(x){plots.OS[[x]]$plot})
  tables.OS <- lapply(1:length(plots.OS) , function(x){plots.OS[[x]]$table})
  
  
  plots.PFI <- lapply(vars.all, function(x){
  vars = x
  cut.off.value <- cutpoints[names(cutpoints) == vars]
  cat.var <- as.factor(ifelse(data.in[,vars] >=  cut.off.value, "high","low"))
  data.in[,lastcol] <- cat.var
  colnames(data.in)[lastcol] <- "aux"
  fit <- survfit(Surv(daystoprogression, progression) ~ aux, data = data.in)
  logrank = surv_pvalue(fit, data.in)
  pval=logrank$pval.txt
  p <- ggsurvplot(fit, data = data.in, risk.table = TRUE, p.val=TRUE, conf.int = TRUE,
                  font.legend = c(18, "plain"), font.tickslab = c(16, "plain"),
                  font.x = c(18, "plain"), font.y = c(18, "plain"),
                  tables.y.text = FALSE,  break.time.by=500, xlim = c(0,2500),
                  legend.labs=c(paste0(vars, "+"),paste0(vars, "-")),
                  palette = c("#34A0D3" ,"#FA5A41")) + 
        guides(colour = guide_legend(nrow = 2))
  p <- p + ylab("PFI probability") + xlab("Time (days)")
  p$plot <- p$plot + theme(legend.key.width = unit(3, "line"), plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"), legend.key.height = unit(1.5, "line"))
  p$plot <- p$plot + ggplot2::annotate("text", x = 2200, y = 0.85, label =pval, size = 5)
  p$table <- p$table + theme(plot.margin = margin(t = 0,  r = 2, b = 0, l = 2, unit = "cm"))
  plots <- list(plot=p$plot, table=p$table)
  return(plots)
  })
  fig.PFI <- lapply(1:length(plots.PFI) , function(x){plots.PFI[[x]]$plot})
  tables.PFI <- lapply(1:length(plots.PFI) , function(x){plots.PFI[[x]]$table})
  
  ggsave(file=paste0("D:/users/fperez/NKI_TMAs_AF/KM_OS_tCycif_cellproportions",suffix[j], ".png"), arrangeGrob(grobs = fig.OS, ncol = 2), width = 28, height = 40, units = "cm")
  
  ggsave(file=paste0("D:/users/fperez/NKI_TMAs_AF/KM_OS_tCycif_cellproportions",suffix[j], "-table.png"), arrangeGrob(grobs = tables.OS, ncol = 2), width = 28, height = 40, units = "cm")
  
  ggsave(file=paste0("D:/users/fperez/NKI_TMAs_AF/KM_PFI_tCycif_cellproportions",suffix[j],".png"), arrangeGrob(grobs = fig.PFI, ncol = 2), width = 28, height = 40, units = "cm")
  
  ggsave(file=paste0("D:/users/fperez/NKI_TMAs_AF/KM_PFI_tCycif_cellproportions",suffix[j],"-table.png"), arrangeGrob(grobs = tables.PFI, ncol = 2), width = 28, height = 40, units = "cm")
}
```